---
title: "AKT3 expression covariate analysis"
author: "Phoebe Parrish"
date: "2023-09-20"
output: html_document
---

# Setup

```{r setup, include=FALSE}

library(biomaRt)
library(depmap)
library(ExperimentHub)
library(tidyverse)
library(ggrepel)
library(ggforce)
library(ggsignif)
library(RColorBrewer)
library(ggpubr) # another way to do significance bars
library(ggprism) # another significance bar package...
library(scales) # scientific notation
library(ggtext) # for using superscript in labels
library(DescTools) # for robust scaling
library(kableExtra)

```


## Assign variables
```{r}

date <- Sys.Date()

in_dir <- file.path("..", "input_data")

df_dir <- file.path("..", "results", "integrated_dfs")

expr_dir <- file.path("..", "results/expression_output")

options(dplyr.summarise.inform = FALSE)


```


## Import files

```{r, message = FALSE}

d.DeKegel_top5_pairs_drug_sens_expr_annot <- read_rds(file.path(expr_dir, "d.top_outlier_pairs"))

d.DeKegel_top5_pairs_drug_sens_cn_annot <- read_rds(file.path(df_dir, "d.DeKegel_top5_pairs_drug_sens_cn_annot"))

d.PI3K_pathway <- read_tsv(file.path(in_dir, "PI3K_Akt_pathway.txt"))

```

### Import functions

```{r}

source("shared_functions.R")

```

# AKT3 expr covariates

## Expression vs. copy number

The purpose of this analysis is to determine whether the loss of expression of *AKT3* in certain cell lines is due to an underlying copy number loss across cell lines. 

```{r}

d.AKT3_expr_outlier_top_hit <- d.DeKegel_top5_pairs_drug_sens_expr_annot %>%
  filter(sorted_gene_pair == "AKT1_AKT3" & compound_name == "GSK2110183")

d.AKT3_expr_outlier_top_hit_withCN <- d.DeKegel_top5_pairs_drug_sens_cn_annot %>%
  filter(sorted_gene_pair == "AKT1_AKT3" & compound_name == "GSK2110183") %>%
  dplyr::select(depmap_id, A2_log_cn) %>%
  rename(AKT3_log_cn = A2_log_cn) %>%
  right_join(d.AKT3_expr_outlier_top_hit, by = "depmap_id")

```

```{r, warning = FALSE, message = FALSE}

ggplot(d.AKT3_expr_outlier_top_hit_withCN, aes(x = A2_log_tpm, y = AKT3_log_cn, color = A2_expr_outlier_flag)) +
  geom_hline(yintercept = 1) +
  geom_point() +
  scale_color_manual(values = c("low" = "#0080C6", "neither" = "#7F7F7F", "high" = "#B84250")) +
  theme_bw() +
  theme(aspect.ratio = 1,
            axis.text = element_text(color = "black"),
            axis.ticks = element_line(color = "black"))

```

It doesn't look like there is a correlation between low copy number and low expression for *AKT3*. 

## Mutations in PI3K/Akt pathway genes

The purpose of this analysis is to determine whether the loss of expression of *AKT3* in certain cell lines is due to mutations in a subset of cell lines in either *AKT3* itself or in other PI3K/Akt pathway genes. 

```{r, message=FALSE}

## get DepMap mutation call information
eh <- ExperimentHub()
eh.depmap <- query(eh, "depmap")

mut_calls <- eh[["EH6121"]]

```

```{r}

PI3K_pathway_genes <- d.PI3K_pathway %>%
  pull(gene_name)

d.mut_calls_PI3K_pathway <- mut_calls %>%
  filter(gene_name %in% PI3K_pathway_genes) 

d.mut_calls_PI3K_pathway %>%
  distinct(genome_change, .keep_all = TRUE) %>%
  group_by(is_deleterious) %>%
  summarize(n = n()) %>%
  print_kbl()

## keep only deleterious mutations
d.mut_calls_PI3K_pathway_deleterious <- d.mut_calls_PI3K_pathway %>%
  filter(is_deleterious == TRUE)

```

```{r}

mut_calls_PI3K_pathway_tcga_hotspot <- d.mut_calls_PI3K_pathway %>%
  filter(is_tcga_hotspot == TRUE) %>%
  distinct(depmap_id, gene_name) %>%
  unite(id) %>%
  pull(id)

## get just cell line and expr info
d.tmp <- d.AKT3_expr_outlier_top_hit %>%
  dplyr::select(depmap_id, A2_log_tpm, A2_expr_outlier_flag) %>%
  rename(AKT3_log_tpm = A2_log_tpm, AKT3_expr_outlier_flag = A2_expr_outlier_flag)

## full join using by = character to copy all cell lines into each pathway member
d.AKT3_expr_outlier_top_hit_AKT_tcga_hotspot <- full_join(d.PI3K_pathway, d.tmp, by = character()) %>%
  unite(id, c(depmap_id, gene_name), remove = FALSE) %>%
  mutate(tcga_hotspot = case_when(
    id %in% mut_calls_PI3K_pathway_tcga_hotspot ~ TRUE,
    TRUE ~ FALSE)) %>%
  mutate(tcga_hotspot = factor(tcga_hotspot, levels = c(TRUE, FALSE)))

```


```{r}

## merge PI3K deleterious mutation calls with expression DF
## get just cell line and expr info
d.tmp <- d.AKT3_expr_outlier_top_hit %>%
  dplyr::select(depmap_id, A2_log_tpm, A2_expr_outlier_flag) %>%
  rename(AKT3_log_tpm = A2_log_tpm, AKT3_expr_outlier_flag = A2_expr_outlier_flag)

mut_calls_PI3K_pathway_deleterious <- d.mut_calls_PI3K_pathway_deleterious %>%
  distinct(depmap_id, gene_name) %>%
  unite(id) %>%
  pull(id)

## full join using by = character to copy all cell lines into each pathway member
d.AKT3_expr_outlier_top_hit_AKT_del_mut <- full_join(d.PI3K_pathway, d.tmp, by = character()) %>%
  unite(id, c(depmap_id, gene_name), remove = FALSE) %>%
  mutate(deleterious_mut = case_when(
    id %in% mut_calls_PI3K_pathway_deleterious ~ TRUE,
    TRUE ~ FALSE)) %>%
  mutate(deleterious_mut = factor(deleterious_mut, levels = c(TRUE, FALSE)))

```

```{r}

d.AKT3_expr_outlier_top_hit_AKT_tcga_hotspot_fisher_pval <- d.AKT3_expr_outlier_top_hit_AKT_tcga_hotspot %>%
  filter(AKT3_expr_outlier_flag != "high") %>%
  mutate(AKT3_expr_outlier_flag = factor(AKT3_expr_outlier_flag, levels = c("low", "neither"))) %>%
  group_by(gene_name) %>%
  summarize(p_val = fisher.test(AKT3_expr_outlier_flag, tcga_hotspot)$p.value)

d.AKT3_expr_outlier_top_hit_AKT_del_mut_fisher_pval <- d.AKT3_expr_outlier_top_hit_AKT_del_mut %>%
  filter(AKT3_expr_outlier_flag != "high") %>%
  mutate(AKT3_expr_outlier_flag = factor(AKT3_expr_outlier_flag, levels = c("low", "neither"))) %>%
  group_by(gene_name) %>%
  summarize(p_val = fisher.test(AKT3_expr_outlier_flag, deleterious_mut)$p.value)

```


```{r}

d.AKT3_expr_outlier_top_hit_AKT_tcga_hotspot <- d.AKT3_expr_outlier_top_hit_AKT_tcga_hotspot %>%
  left_join(d.AKT3_expr_outlier_top_hit_AKT_tcga_hotspot_fisher_pval, by = "gene_name") %>%
  mutate(tcga_hotspot = factor(tcga_hotspot, levels = c(TRUE, FALSE)))

d.AKT3_expr_outlier_top_hit_AKT_del_mut <- d.AKT3_expr_outlier_top_hit_AKT_del_mut %>%
  left_join(d.AKT3_expr_outlier_top_hit_AKT_del_mut_fisher_pval, by = "gene_name")

```


```{r, fig.height=10, fig.width=10}

d.label <- d.AKT3_expr_outlier_top_hit_AKT_del_mut %>%
  filter(AKT3_expr_outlier_flag == "low" | AKT3_expr_outlier_flag == "neither") %>%
  mutate(AKT3_expr_outlier_flag = factor(AKT3_expr_outlier_flag, levels = c("low", "neither"))) %>%
  group_by(gene_name) %>%
  mutate(y.position = (max(AKT3_log_tpm) + 0.1*(max(AKT3_log_tpm) - min(AKT3_log_tpm)))) %>%
  dplyr::select(gene_name, p_val, y.position) %>%
  distinct(gene_name, .keep_all = TRUE) %>%
  mutate(group1 = TRUE,
         group2 = FALSE) %>%
  mutate(p_val_label = case_when(
    p_val < 0.05 ~ scientific(p_val, 3),
    p_val >= 0.05 ~ "ns")) %>%
  ungroup()

ggplot(d.AKT3_expr_outlier_top_hit_AKT_del_mut, aes(x = deleterious_mut, y = AKT3_log_tpm, color = AKT3_expr_outlier_flag)) +
  geom_jitter(size = 1) +
  stat_pvalue_manual(d.label, label = "{p_val_label}", hide.ns = TRUE,## "P = {p_val}"
                     tip.length = 0.03, size = 3, vjust = -0.5) +
  scale_color_manual(values = c("low" = "#0080C6", "neither" = "#7F7F7F", "high" = "#B84250")) +
  scale_y_continuous(expand = expansion(mult = c(0.1, 0.25))) +
  labs(x = "Deleterious_mutation_present", y = "AKT3_log2(TPM)") +
  theme_bw() +
  theme(axis.text = element_text(color = "black"),
        axis.ticks = element_line(color = "black")) +
  facet_wrap(~gene_name, ncol = 6, scales = "free_y")

```


```{r, fig.height=10, fig.width=10}

d.label <- d.AKT3_expr_outlier_top_hit_AKT_tcga_hotspot %>%
  filter(AKT3_expr_outlier_flag == "low" | AKT3_expr_outlier_flag == "neither") %>%
  mutate(AKT3_expr_outlier_flag = factor(AKT3_expr_outlier_flag, levels = c("low", "neither"))) %>%
  group_by(gene_name) %>%
  mutate(y.position = (max(AKT3_log_tpm) + 0.1*(max(AKT3_log_tpm) - min(AKT3_log_tpm)))) %>%
  dplyr::select(gene_name, p_val, y.position) %>%
  distinct(gene_name, .keep_all = TRUE) %>%
  mutate(group1 = TRUE,
         group2 = FALSE) %>%
  mutate(p_val_label = case_when(
    p_val < 0.05 ~ scientific(p_val, 3),
    p_val >= 0.05 ~ "ns")) %>%
  ungroup()

ggplot(d.AKT3_expr_outlier_top_hit_AKT_tcga_hotspot, aes(x = tcga_hotspot, y = AKT3_log_tpm, color = AKT3_expr_outlier_flag)) +
  geom_jitter(size = 1) +
  stat_pvalue_manual(d.label, label = "{p_val_label}", hide.ns = TRUE,## "P = {p_val}"
                     tip.length = 0.03, size = 3, vjust = -0.5) +
  scale_color_manual(values = c("low" = "#0080C6", "neither" = "#7F7F7F", "high" = "#B84250")) +
  scale_y_continuous(expand = expansion(mult = c(0.1, 0.25))) +
  labs(x = "TCGA_hotspot_mutation_present", y = "AKT3_log2(TPM)") +
  theme_bw() +
  theme(axis.text = element_text(color = "black"),
        axis.ticks = element_line(color = "black")) +
  facet_wrap(~gene_name, ncol = 6, scales = "free_y")

```

There may be a correlation between low *AKT3* expression and TCGA hotspot mutations in *PIK3CA*, but there are still plenty of *AKT3*-low outlier cell lines that do not have *PIK3CA* hotspot mutations, so this likely isn't the only contributor to this effect. 
