---
title: "Integrate drug sensitivity, copy number, and expression datasets"
author: "Phoebe Parrish"
date: "2023-09-10"
output: html_document
---

# Setup

```{r setup, include=FALSE, results='hide', message=FALSE, warning=FALSE}

library(biomaRt)
library(depmap)
library(ExperimentHub)
library(tidyverse)
# library(tidylog)
library(kableExtra)

```


## Assign variables
```{r}

date <- Sys.Date()

in_dir <- file.path("..", "input_data")

out_dir <- file.path("..", "results/integrated_dfs")

```


## Import files

```{r, message=FALSE}

d.DeKegel <- read_csv(file.path(in_dir, "DeKegel_Table_S8.csv"))

d.Broad_drug_info <- read_csv(file.path(in_dir, "primary-screen-replicate-collapsed-treatment-info.csv"))

# d.Broad_secondary_auc <- read_csv(file.path(in_dir, "secondary-screen-dose-response-curve-parameters.csv"))
# d.Broad_secondary_doses <- read_csv(file.path(in_dir, "secondary-screen-dose-response-curve-parameters.csv"))

```

```{r}

source("shared_functions.R")

```

# Integrate datasets

## DepMap drug data
Also see: https://depmap.org/repurposing/
```{r message=FALSE, warning=FALSE}

eh <- ExperimentHub()
eh.depmap <- query(eh, "depmap")

```

```{r message=FALSE, warning=FALSE}

metadata <- eh[["EH6122"]]
cn <- eh[["EH6119"]]
expr <- eh[["EH6120"]]
drug_sensitivity <- eh[["EH3087"]]
drug_metadata <- eh[["EH3086"]]
## note: drug info is from 19Q4, these are 19Q3

```



### Get gene list

#### De Kegel top 5%
```{r}

d.DeKegel <- d.DeKegel %>%
  unite(ensembl_pair, c(A1_ensembl, A2_ensembl), sep = "_", remove = FALSE) %>%
  unite(entrez_pair, c(A1_entrez, A2_entrez), sep = "_", remove = FALSE)
print_kbl(head(d.DeKegel))

DeKegel_top5_genes_entrez <- d.DeKegel %>%
  filter(prediction_percentile < 5) %>%
  dplyr::select(A1_entrez, A2_entrez) %>%
  gather() %>%
  distinct(value) %>%
  pull(value)

DeKegel_top5_genes_hgnc <- d.DeKegel %>%
  filter(prediction_percentile < 5) %>%
  dplyr::select(A1, A2) %>%
  gather() %>%
  distinct(value) %>%
  pull(value)
  
```

```{r}

d.expr_DeKegel_top5 <- expr %>%
  dplyr::filter(entrez_id %in% DeKegel_top5_genes_entrez)
# d.expr_DeKegel_top5

d.cn_DeKegel_top5 <- cn %>%
  dplyr::filter(entrez_id %in% DeKegel_top5_genes_entrez)
# d.cn_DeKegel_top5

```


## Paralog + drug info
### Get top 5% predicted synthetic lethal genes

```{r}

### All top 5% genes (no druggability filter)
d.DeKegel_top5 <- d.DeKegel %>%
  filter(prediction_percentile < 5) %>%
  rename(A1_hgnc = A1, A2_hgnc = A2) %>%
  dplyr::select(prediction_rank:family_size)
print_kbl(head(d.DeKegel_top5))

## pivot to get each gene on its own line in "long" format
d.DeKegel_top5_genes <- d.DeKegel_top5 %>%
  pivot_longer(cols = starts_with("A"),
               names_to = c("gene_pos", ".value"),
               names_pattern = "A(.)_(.*)")
print_kbl(head(d.DeKegel_top5_genes))

```

#### Add drug info
```{r}

# drug_metadata
# drug_sensitivity

## get annotations for drugs that target De Kegel top 5% genes
d.Broad_drug_info_DeKegel_top5 <- d.Broad_drug_info %>%
  dplyr::filter(grepl(str_c(DeKegel_top5_genes_hgnc, collapse = "|"), target))
# d.Broad_drug_info_DeKegel_top5

## filtering join to get drug sensitivity data only for drugs targeting top 5% genes
## also adds drug annotations
d.drug_sens_DeKegel_top5 <- drug_sensitivity %>%
  inner_join(d.Broad_drug_info_DeKegel_top5, by = c("compound" = "column_name"))
# head(d.drug_sens_DeKegel_top5)

## reformat cell line metadata and add to filtered drug sensitivity DF
d.drug_sens_DeKegel_top5 <- metadata %>%
  dplyr::select(depmap_id, stripped_cell_line_name, primary_disease,
                subtype_disease, lineage, lineage_subtype) %>%
  right_join(d.drug_sens_DeKegel_top5, by = "depmap_id") %>%
  dplyr::select(-c(cell_line, screen_id, disease.area:phase)) %>%
  rename(cell_line = stripped_cell_line_name, 
         compound_name = name,
         compound_id = compound) 
# d.drug_sens_DeKegel_top5

## separate out DF so each target is on its own line
d.drug_sens_DeKegel_top5_target <- d.drug_sens_DeKegel_top5 %>%
  separate_rows(target)
print_kbl(head(d.drug_sens_DeKegel_top5_target))

```

```{r}

## join DeKegel top 5% DF with filtered drug sensitivity DF
d.DeKegel_top5_genes_drug_sens <- d.drug_sens_DeKegel_top5_target <- d.DeKegel_top5_genes %>%
  inner_join(d.drug_sens_DeKegel_top5_target, by = c("hgnc" = "target"))
# d.DeKegel_top5_genes_drug_sens

```

#### Rm double-targeting drugs
```{r}

## remove 2nd hit for drugs that target both paralogs

## check for drugs that target both members of a pair
d.DeKegel_top5_genes_drug_sens %>%
  distinct(sorted_gene_pair, compound_id, gene_pos, .keep_all = TRUE) %>%
  group_by(sorted_gene_pair, compound_id) %>%
  summarize(n = n()) %>%
  filter(n >1)

## get list of double-targeting gene pair and compound names
double_targeting_drugs <- d.DeKegel_top5_genes_drug_sens %>%
  distinct(sorted_gene_pair, compound_id, gene_pos, .keep_all = TRUE) %>%
  group_by(sorted_gene_pair, compound_id) %>%
  summarize(n = n()) %>%
  filter(n >1) %>%
  unite("double_targeting_compounds", c("sorted_gene_pair", "compound_id")) %>%
  distinct(double_targeting_compounds) %>%
  pull()

## how many double-targetings will be removed?
d.DeKegel_top5_genes_drug_sens %>%
  unite("gene_pair_compound", c("sorted_gene_pair", "compound_id"), remove = FALSE) %>%
  mutate(double_targeting_drug = ifelse(gene_pair_compound %in% double_targeting_drugs, TRUE, FALSE)) %>%
  group_by(double_targeting_drug) %>%
  summarize(n = n())

# 833476 = 721 double-targeting drug/pair combos * 578 cell lines

## remove the second entry for the double-targeting drugs (the first entry will be retained)
d.DeKegel_top5_genes_drug_sens <- d.DeKegel_top5_genes_drug_sens %>%
  unite("gene_pair_compound", c("sorted_gene_pair", "compound_id"), remove = FALSE) %>%
  mutate(double_targeting_drug = ifelse(gene_pair_compound %in% double_targeting_drugs, TRUE, FALSE)) %>%
  mutate(rm_flag = case_when(
    double_targeting_drug == TRUE & gene_pos == "2" ~ TRUE,
    TRUE ~ FALSE)) %>% # if the above is not true, flag = FALSE
  filter(rm_flag == FALSE) %>%
  dplyr::select(-c("gene_pair_compound", "rm_flag"))
# d.DeKegel_top5_genes_drug_sens
# 1793534-(833476/2)

```

### Reformat DF
```{r}

d.DeKegel_top5_pairs_drug_sens <- d.DeKegel_top5_genes_drug_sens %>%
  ## add new column to flag which gene each drug targets
  mutate(drug_targets = case_when(
    double_targeting_drug == TRUE ~ "both",
    double_targeting_drug == FALSE & gene_pos == "1" ~ "A1",
    double_targeting_drug == FALSE & gene_pos == "2" ~ "A2")) %>%
  ## separate pair-based columns to get individual gene IDs
  separate(col = sorted_gene_pair, c("A1_hgnc", "A2_hgnc"), remove = FALSE) %>%
  separate(col = entrez_pair, c("A1_entrez", "A2_entrez"), remove = FALSE) %>%
  separate(col = ensembl_pair, c("A1_ensembl", "A2_ensembl"), remove = FALSE) %>%
  ## drop old gene ID columns
  dplyr::select(-c(gene_pos:ensembl))
print_kbl(head(d.DeKegel_top5_pairs_drug_sens))

```

## Add CN info
```{r}


## keep only relevant columns from DepMap CN dataframe
d.cn_tmp <- d.cn_DeKegel_top5 %>%
  dplyr::select(depmap_id, entrez_id, log_copy_number) 

## convert entrez ID type to character
d.cn_tmp$entrez_id <- as.character(d.cn_tmp$entrez_id)
# d.cn_tmp

## add CN info to drug sensitivity DF, remove rows containing NA
d.DeKegel_top5_pairs_drug_sens_cn <- d.DeKegel_top5_pairs_drug_sens %>%
  ## add A1 CN info
  left_join(d.cn_tmp, by = c("depmap_id", "A1_entrez" = "entrez_id")) %>%
  rename(A1_log_cn = log_copy_number) %>%
  ## add A2 CN info
  left_join(d.cn_tmp, by = c("depmap_id", "A2_entrez" = "entrez_id")) %>%
  rename(A2_log_cn = log_copy_number) %>%
  ## remove rows that contain NAs
  filter(!is.na(A1_log_cn) | !is.na(A2_log_cn)) %>%
  filter(!is.na(dependency))
print_kbl(head(d.DeKegel_top5_pairs_drug_sens_cn))

```

### Reformat DF
```{r}

## calculate combined CN for each pair
d.DeKegel_top5_pairs_drug_sens_cn <- d.DeKegel_top5_pairs_drug_sens_cn %>%
  mutate(combined_log_cn = log2((2^A1_log_cn) + (2^A2_log_cn)))
# d.DeKegel_top5_pairs_drug_sens_cn

## reorganize columns
d.DeKegel_top5_pairs_drug_sens_cn_annot <- d.DeKegel_top5_pairs_drug_sens_cn %>%
  dplyr::select(depmap_id, compound_id, sorted_gene_pair, dependency,
                A1_log_cn:A2_log_cn, combined_log_cn, A1_hgnc:A2_ensembl, prediction_rank,
                prediction_percentile, prediction_score:family_size,
                broad_id:drug_targets, cell_line:lineage_subtype)
# d.DeKegel_top5_pairs_drug_sens_cn_annot

```

```{r}

## save CN output
write_rds(d.DeKegel_top5_pairs_drug_sens_cn_annot,
          file.path(out_dir, "d.DeKegel_top5_pairs_drug_sens_cn_annot"))
write_tsv(d.DeKegel_top5_pairs_drug_sens_cn_annot,
          file.path(out_dir, "d.DeKegel_top5_pairs_drug_sens_cn_annot.txt"))

```

## Add expression info
```{r}

## keep only relevant columns from DepMap expression dataframe
d.expr_tmp <- d.expr_DeKegel_top5 %>%
  dplyr::select(depmap_id, entrez_id, rna_expression)

## convert entrez ID type to character
d.expr_tmp$entrez_id <- as.character(d.expr_tmp$entrez_id)
# d.expr_tmp

## add expr info to drug sensitivity DF, remove rows containing NA
d.DeKegel_top5_pairs_drug_sens_expr <- d.DeKegel_top5_pairs_drug_sens %>%
  ## add A1 expr info
  left_join(d.expr_tmp, by = c("depmap_id", "A1_entrez" = "entrez_id")) %>%
  rename(A1_log_tpm = rna_expression) %>%
  ## add A2 expr info
  left_join(d.expr_tmp, by = c("depmap_id", "A2_entrez" = "entrez_id")) %>%
  rename(A2_log_tpm = rna_expression) %>%
  ## remove rows that contain NAs
  filter(!is.na(A1_log_tpm) | !is.na(A2_log_tpm)) %>%
  filter(!is.na(dependency))
print_kbl(head(d.DeKegel_top5_pairs_drug_sens_expr))

```

### Reformat DF
```{r}

## calculate combined expr for each pair
d.DeKegel_top5_pairs_drug_sens_expr <- d.DeKegel_top5_pairs_drug_sens_expr %>%
  mutate(combined_log_tpm = log2((2^A1_log_tpm) + (2^A2_log_tpm)))
# d.DeKegel_top5_pairs_drug_sens_expr

## reorganize columns 
d.DeKegel_top5_pairs_drug_sens_expr_annot <- d.DeKegel_top5_pairs_drug_sens_expr %>%
  dplyr::select(depmap_id, compound_id, sorted_gene_pair, dependency,
                A1_log_tpm:A2_log_tpm, combined_log_tpm, A1_hgnc:A2_ensembl, 
                prediction_rank, prediction_percentile, prediction_score:family_size,
                broad_id:drug_targets, cell_line:lineage_subtype)
# d.DeKegel_top5_pairs_drug_sens_expr_annot

```

```{r}

## save expr output
write_rds(d.DeKegel_top5_pairs_drug_sens_expr_annot,
          file.path(out_dir, "d.DeKegel_top5_pairs_drug_sens_expr_annot"))
write_tsv(d.DeKegel_top5_pairs_drug_sens_expr_annot,
          file.path(out_dir, "d.DeKegel_top5_pairs_drug_sens_expr_annot.txt"))

```


## Single gene DFs
To compare the results from our paralog analyses to those seen for known drug target genes, I also got the information for single drug target genes with their expression and copy number. 

```{r}

## get druggability, expression, CN for all drug target genes (not just SL paralogs)
# d.Broad_drug_info
# drug_sensitivity

d.drug_sens_all_genes <- drug_sensitivity %>%
  inner_join(d.Broad_drug_info, by = c("compound" = "column_name"))
# d.drug_sens_all_genes

## reformat cell line metadata and add to filtered drug sensitivity DF
d.drug_sens_all_genes <- metadata %>%
  dplyr::select(depmap_id, stripped_cell_line_name, primary_disease,
                subtype_disease, lineage, lineage_subtype) %>%
  right_join(d.drug_sens_all_genes, by = "depmap_id") %>%
  dplyr::select(-c(cell_line, screen_id, disease.area:phase)) %>%
  rename(cell_line = stripped_cell_line_name,
         compound_name = name,
         compound_id = compound)
# d.drug_sens_all_genes

## separate out DF so each target is on its own line
d.drug_sens_all_genes <- d.drug_sens_all_genes %>%
  separate_rows(target) 
print_kbl(head(d.drug_sens_all_genes))

```

```{r}

d.cn_tmp <- cn %>%
  dplyr::select(-c(gene, cell_line))

d.expr_tmp <- expr %>%
  dplyr::select(-c(gene, cell_line))

```

```{r}

d.drug_sens_all_genes <- d.drug_sens_all_genes %>%
  filter(!is.na(target))

## copy number dataframe
d.drug_sens_all_genes_cn <- inner_join(d.drug_sens_all_genes, d.cn_tmp, by = c("depmap_id", "target" = "gene_name"))
print_kbl(head(d.drug_sens_all_genes_cn))

## expression dataframe
d.drug_sens_all_genes_expr <- inner_join(d.drug_sens_all_genes, d.expr_tmp, by = c("depmap_id", "target" = "gene_name"))
print_kbl(head(d.drug_sens_all_genes_expr))

```

```{r}

write_rds(d.drug_sens_all_genes_cn, file.path(out_dir, "d.drug_sens_all_genes_cn"))
write_rds(d.drug_sens_all_genes_expr, file.path(out_dir, "d.drug_sens_all_genes_expr"))

```





