---
title: "pgPEN_DeKegel_comparison"
author: "Phoebe Parrish"
date: "2022-01-07"
output: html_document
---

# To Do
* update depmap version - use more recent data (I am currently using 19Q1)
* CN as continuous variable, look at outliers (not top 2 quartiles)

# Setup

```{r setup, include=FALSE}

# library(biomaRt)
# library(depmap)
# library(ExperimentHub)
library(tidyverse)
library(ggrepel)
library(ggforce)
library(ggsignif)
library(RColorBrewer)
library(ggpubr) # another way to do significance bars
library(ggprism) # another significance bar package...
library(scales) # scientific notation
library(ggtext) # for using superscript in labels
library(DescTools) # for robust scaling

```


## Assign variables
```{r}

date <- Sys.Date()

## work laptop
base_dir <- file.path("/Users", "pparrish", "Google_Drive", "01_Research", "01_Projects", "01_Paralog_pgRNA",
                      "01_Data", "Paralog_SL_meta_analysis")

in_dir <- file.path(base_dir, "01_input_files")

out_dir <- file.path(base_dir, "02_output_files")

```


## Import files

```{r, message = FALSE}

d.DeKegel_top5_pairs_drug_sens_cn_annot <- read_rds(file.path(out_dir, "d.DeKegel_top5_pairs_drug_sens_cn_annot"))

d.Broad_drug_info <- read_csv(file.path(in_dir, "primary-screen-replicate-collapsed-treatment-info.csv"))

d.Broad_secondary_auc <- read_csv(file.path(in_dir, "secondary-screen-dose-response-curve-parameters.csv"))

d.Broad_secondary_doses <- read_csv(file.path(in_dir, "secondary-screen-dose-response-curve-parameters.csv"))

d.drug_sens_all_genes_cn <- read_rds(file.path(out_dir, "d.drug_sens_all_genes_cn"))

```

### Reformat CN DF
```{r}

## keep only relevant columns
d.DeKegel_top5_pairs_drug_sens_cn <- d.DeKegel_top5_pairs_drug_sens_cn_annot %>%
  dplyr::select(-c(entrez_pair:prediction_rank, prediction_score:family_size,
                   broad_id, primary_disease, subtype_disease))

```


### Useful functions
```{r}

## adjust p-values
BH_adjust <- function(df){
  
  d.p_vals <- df %>%
  arrange(p_val) %>%
  dplyr::select(target, compound_id, p_val) 

  p_vals <- d.p_vals %>%
    arrange(p_val) %>%
    dplyr::select(p_val) %>%
    pull()
  
  fdr_vals <- p.adjust(p_vals, method = "BH")

  d.fdr_vals <- tibble(fdr = fdr_vals)
  
  d.stats <- d.p_vals %>%
    add_column(d.fdr_vals)
  d.stats
  
  return(d.stats)
  
}


```


```{r}

## summarize stats output
summarize_stats <- function(df){
  df_new <- df %>%
    distinct(target, compound_id, .keep_all = TRUE)
  
  n_p_L0.05 <- df_new %>%
    filter(p_val < 0.05) %>%
    nrow()
  
  n_p_L0.01 <- df_new %>%
    filter(p_val < 0.01) %>%
    nrow()
  
  n_fdr_L0.05 <- df_new %>%
    filter(fdr < 0.05) %>%
    nrow()
  
  n_fdr_L0.1 <- df_new %>%
    filter(fdr < 0.1) %>%
    nrow()
  
  cat(paste("df:", deparse(substitute(df)), "\n", sep = " "))
  cat("# target/compound pairs with\n")
  cat(paste("p < 0.05:", n_p_L0.05, "\n", sep = " "))
  cat(paste("p < 0.01:", n_p_L0.01, "\n", sep = " "))
  cat(paste("FDR < 0.05:", n_fdr_L0.05, "\n", sep = " "))
  cat(paste("FDR < 0.1:", n_fdr_L0.1, "\n\n", sep = " "))
  
}

```

# Analysis

## Combined CN

### Phase 1 drug sensitivity

#### Quartiles/WRST
```{r}

## bin combined CN for each pair into 4 quartiles
d.DeKegel_top5_pairs_drug_sens_combined_cn_bin <- d.DeKegel_top5_pairs_drug_sens_cn %>%
  group_by(sorted_gene_pair) %>%
  mutate(combined_cn_bin = ntile(combined_log_cn, n = 4)) %>%
  ## factorized binned variable for easy plotting
  mutate(combined_cn_bin = factor(combined_cn_bin, levels = c(1, 2, 3, 4)))

## calculate the WRST p-value for bin 1 (lowest CN) vs. bin 4 (highest CN)
d.DeKegel_top5_pairs_drug_sens_combined_cn_bin_pval_summary <- d.DeKegel_top5_pairs_drug_sens_combined_cn_bin %>%
  filter(combined_cn_bin %in% c(1, 4)) %>%
  group_by(sorted_gene_pair, compound_id) %>%
  summarize("pval" = wilcox.test(dependency ~ combined_cn_bin, alternative = "less")$p.value)
d.DeKegel_top5_pairs_drug_sens_combined_cn_bin_pval_summary

## add p-values to main DF
d.DeKegel_top5_pairs_drug_sens_combined_cn_bin_wrst <- left_join(d.DeKegel_top5_pairs_drug_sens_combined_cn_bin,
                                                       d.DeKegel_top5_pairs_drug_sens_combined_cn_bin_pval_summary,
                                            by = c("sorted_gene_pair", "compound_id")) %>%
  ungroup()
d.DeKegel_top5_pairs_drug_sens_combined_cn_bin_wrst



```

```{r}

## FDR-adjust p-values
pvals <- d.DeKegel_top5_pairs_drug_sens_combined_cn_bin_wrst %>%
  distinct(sorted_gene_pair, compound_id, .keep_all = TRUE) %>%
  arrange(pval) %>%
  dplyr::select(pval) %>%
  pull()

fdr_vals <- p.adjust(pvals, method = "BH")
head(fdr_vals)

d.fdr_vals <- tibble(fdr = fdr_vals)

d.fdr_vals %>%
  filter(fdr < 0.05) %>%
  nrow()

d.fdr_vals %>%
  filter(fdr < 0.1) %>%
  nrow()

d.pvals_labels <- d.DeKegel_top5_pairs_drug_sens_combined_cn_bin_wrst %>%
  distinct(sorted_gene_pair, compound_id, .keep_all = TRUE) %>%
  arrange(pval) %>%
  dplyr::select(sorted_gene_pair, compound_id)
d.pvals_labels

d.fdr_vals_labeled <- d.pvals_labels %>%
  add_column(d.fdr_vals)
d.fdr_vals_labeled

d.DeKegel_top5_pairs_drug_sens_combined_cn_bin_wrst <- d.DeKegel_top5_pairs_drug_sens_combined_cn_bin_wrst %>%
  left_join(d.fdr_vals_labeled, by = c("sorted_gene_pair", "compound_id"))
d.DeKegel_top5_pairs_drug_sens_combined_cn_bin_wrst

```

##### Plot
Plotting FDR values from DF directly onto ggsignif(): 
https://github.com/const-ae/ggsignif/issues/29
https://cran.r-project.org/web/packages/ggsignif/ggsignif.pdf

```{r}

## plot all FDR < 0.1 pairs (sorted by FDR)
## add my own labels!!

d.plot <- d.DeKegel_top5_pairs_drug_sens_combined_cn_bin_wrst %>% 
  ungroup() %>%
  filter(fdr < 0.1) %>%
  unite("gene_pair_compound_label", c("sorted_gene_pair", "compound_name"), sep = "\n", remove = FALSE) %>%
  mutate("gene_pair_compound_label" = reorder(gene_pair_compound_label, fdr),
         "sorted_gene_pair" = reorder(sorted_gene_pair, fdr))

length <- d.plot %>%
  ungroup() %>%
  distinct(gene_pair_compound_label) %>%
  nrow()

# d.fdr_labels <- d.DeKegel_top5_combined_cn_drug_pair_fdr_signif %>%
#   ungroup() %>%
#   dplyr::select(gene_pair_compound_label, combined_cn_bin, fdr)
  
aspect_ratio <- 0.8
n_rows <- 1
n_cols <- 1
n_pages <- ceiling(length/(n_rows*n_cols))

# pdf(file.path(out_dir, paste(date, "DeKegel_top5_pairs_drug_sens_combined_cn_bin_wrst_fdr_L10.pdf", sep = "_")),
#     width = 4, height = 2.5)
# for(i in 1:n_pages){
#   p <-  ggplot(d.plot, aes(x = combined_cn_bin, y = dependency, fill = combined_cn_bin)) +
#     geom_hline(yintercept = 0) +
#     geom_boxplot() +
#     geom_signif(comparisons = list(c(1, 4)),
#                 map_signif_level = FALSE,
#                 test = "wilcox.test",
#                 test.args = list(alternative = "less", p.adjust.method = "BH"),
#                 # annotations = fdr_labels, # fix this to get FDR plotted direclty
#                 textsize = 3.2) +
#     # add_pvalue(d.fdr_labels) +
#     ## adds 10% of space on top & bottom so p-value label will fit
#     scale_y_continuous(expand = expansion(mult = 0.15)) +
#     scale_fill_brewer(palette = "YlGnBu") +
#     labs(x = "Combined_CN_bin",
#          y = "LFC_treated_vs_DMSO") +
#     theme_bw() +
#     theme(aspect.ratio = aspect_ratio,
#           axis.text = element_text(color = "black"),
#           axis.ticks = element_line(color = "black"),
#           legend.position = "none") +
#     facet_wrap_paginate(~gene_pair_compound_label, scales = "free_y",
#                         nrow = n_rows, ncol = n_cols, page = i)
# 
#   q <- ggplot(d.plot, aes(x = combined_cn_bin, y = (2^combined_log_cn), fill = combined_cn_bin)) +
#     geom_hline(yintercept = 4) +
#     geom_hline(yintercept = 3, linetype = "dashed") +
#     geom_hline(yintercept = 5, linetype = "dashed") +
#     geom_boxplot() +
#     ## adds 10% of space on top & bottom so p-value label will fit
#     # scale_y_continuous(expand = expansion(mult = 0.15)) +
#     scale_fill_brewer(palette = "YlGnBu") +
#     labs(x = "Combined_CN_bin",
#          y = "Combined_CN") +
#     coord_trans(y = "log2") +
#     theme_bw() +
#     theme(aspect.ratio = aspect_ratio,
#           axis.text = element_text(color = "black"),
#           axis.ticks = element_line(color = "black"),
#           legend.position = "none") +
#     facet_wrap_paginate(~gene_pair_compound_label, scales = "free_y",
#                         nrow = n_rows, ncol = n_cols, page = i)
#   print(ggarrange(p, q, ncol = 2, align = "hv"))
# }
# dev.off()

```


#### Continuous/linear regression

Per Joe's suggestion from my 2021-12-14 lab meeting, I will try analyzing the data using CN as a continuous variable rather than semi-arbitrarily binning the values, since there is often not a big difference in actual CN between different bins. 

I will also try to look at the CN of each individual pair-member to see if the "outliers" seem to be driven by the loss of a single paralog (although this may be more relevant for expression than CN).

```{r}

## test linear regression analysis and plots

d.tmp <- d.DeKegel_top5_pairs_drug_sens_cn %>%
  filter(sorted_gene_pair == "IMPDH1_IMPDH2" & compound_name == "mycophenolic-acid") %>%
  mutate(combined_cn = 2^combined_log_cn)
d.tmp

test_lm <- lm(dependency ~ combined_log_cn, data = d.tmp)
test_summary <- broom::glance(test_lm)
test_summary

summary(test_lm)
min(d.tmp$combined_cn)
max(d.tmp$dependency)

d.label <- d.tmp %>%
  group_modify(~ broom::glance(lm(dependency ~ combined_log_cn, data = .x))) %>%
  dplyr::ungroup() %>%
  mutate(r.squared = round(r.squared, 3),
         p.value = scientific(p.value, 3))
d.label

## plots
ggplot(d.tmp, aes(x = combined_cn, y = dependency)) + 
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "gray60") +
  geom_richtext(data = d.label, 
                aes(x = min(d.tmp$combined_cn), y = max(d.tmp$dependency),
                ## this package can do HTML-formatting for text
                label = paste0("r<sup>2</sup>=", r.squared, "<br>p=", p.value)),
                ## adjust label alignment and remove borders
                hjust = 0, vjust = 1, label.color = NA, size = 3.5,
                ## label background = transparent, text is not
                fill = alpha(c("white"), 0.75)) + 
  coord_trans(x = "log2") +
  labs(x = "Combined_copy_number", y = "LFC_drug_vs_DMSO") +
  theme_bw() +
  theme(aspect.ratio = 1,
        axis.text = element_text(color = "black"),
        axis.ticks = element_line(color = "black"))

## testing out a plot where dots are colored by cell line lineage
ggplot(d.tmp, aes(x = combined_cn, y = dependency)) + 
  geom_point(aes(color = lineage)) +
  geom_smooth(method = "lm", se = FALSE, color = "black") +
  geom_richtext(data = d.label,
            aes(x = min(d.tmp$combined_cn), y = max(d.tmp$dependency),
                label = paste0("r<sup>2</sup>=", r.squared, "<br>p=", p.value)),
            hjust = 0, vjust = 1, label.color = NA, size = 3.5,
            fill = alpha(c("white"), 0.6)) +
  coord_trans(x = "log2") +
  labs(x = "Combined_copy_number", y = "LFC_drug_vs_DMSO") +
  theme_bw() +
  theme(aspect.ratio = 1,
        axis.text = element_text(color = "black"),
        axis.ticks = element_line(color = "black"))

## plot each paralog CN independently
d.tmp %>%
  dplyr::select(A1_log_cn, A2_log_cn) %>%
  pivot_longer(cols = c(A1_log_cn, A2_log_cn),
               names_to = "pos",
               values_to = "log_cn") %>%
  ggplot(aes(x = pos, y = (2^log_cn))) +
  # geom_point() +
  geom_boxplot(fill = "gray") +
  coord_trans(y = "log2") +
  labs(x = "Paralog", y = "Copy_number") +
  scale_x_discrete(labels = c("A1", "A2")) +
  theme_bw() +
  theme(aspect.ratio = 1,
        axis.text = element_text(color = "black"),
        axis.ticks = element_line(color = "black"))

## plot combined CN - retire
ggplot(d.tmp, aes(x = sorted_gene_pair, y = (2^combined_log_cn))) +
  # geom_point() +
  geom_hline(yintercept = 4) +
  geom_hline(yintercept = 3, linetype = "dashed") +
  geom_hline(yintercept = 5, linetype = "dashed") +
  geom_boxplot(fill = "gray") +
  coord_trans(y = "log2") +
  labs(y = "Combined_copy_number") +
  theme_bw() +
  theme(aspect.ratio = 1.25,
        axis.text = element_text(color = "black"),
        axis.ticks = element_line(color = "black"))
  
```

Note! This is how you can set the background of a label to be transparent while the text is not in one command (see 2nd answer on this StackOverflow post): https://stackoverflow.com/questions/48557889/ggrepel-label-with-transparent-background-but-visible-font

```{r}

d.tmp <- d.DeKegel_top5_pairs_drug_sens_cn %>%
  filter((sorted_gene_pair == "IMPDH1_IMPDH2" & compound_id == "BRD-A11132253-001-08-2::2.5::HTS") |
           sorted_gene_pair == "AURKA_AURKC" & compound_name == "ENMD-2076") %>%
  mutate(combined_cn = 2^combined_log_cn) %>%
  group_by(sorted_gene_pair, compound_id) %>%
  mutate(group_label_x = min(combined_cn),
         group_label_y = max(dependency)) %>%
  ungroup()
d.tmp


d.label <- d.tmp %>%
  group_by(sorted_gene_pair, compound_id) %>%
  group_modify(~ broom::glance(lm(dependency ~ combined_log_cn, data = .x))) %>%
  dplyr::ungroup() %>%
  mutate(r.squared = round(r.squared, 3),
         p.value = scientific(p.value, 3)) 
d.label

d.label_new <- d.tmp %>%
  dplyr::select(sorted_gene_pair, compound_id, group_label_x, group_label_y) %>%
  distinct(sorted_gene_pair, .keep_all = TRUE) %>%
  right_join(d.label, by = c("sorted_gene_pair", "compound_id"))

## plots
ggplot(d.tmp, aes(x = combined_cn, y = dependency)) + 
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "gray60") +
  geom_richtext(data = d.label, 
                aes(x = min(d.tmp$combined_cn), y = max(d.tmp$dependency),
                ## this package can do HTML-formatting for text
                label = paste0("r<sup>2</sup>=", r.squared, "<br>p=", p.value)),
                ## adjust label alignment and remove borders
                hjust = 0, vjust = 1, label.color = NA, size = 3.5,
                ## label background = transparent, text is not
                fill = alpha(c("white"), 0.75)) + 
  coord_trans(x = "log2") +
  labs(x = "Combined_copy_number", y = "LFC_drug_vs_DMSO") +
  theme_bw() +
  theme(aspect.ratio = 1,
        axis.text = element_text(color = "black"),
        axis.ticks = element_line(color = "black")) +
  facet_wrap(~sorted_gene_pair+compound_id, scales = "free")

ggplot(d.tmp, aes(x = combined_cn, y = dependency)) + 
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "gray60") +
  geom_richtext(data = d.label_new, 
                aes(x = group_label_x, y = group_label_y,
                ## this package can do HTML-formatting for text
                label = paste0("r<sup>2</sup>=", r.squared, "<br>p=", p.value)),
                ## adjust label alignment and remove borders
                hjust = 0, vjust = 1, label.color = NA, size = 3.5,
                ## label background = transparent, text is not
                fill = alpha(c("white"), 0.75)) + 
  coord_trans(x = "log2") +
  labs(x = "Combined_copy_number", y = "LFC_drug_vs_DMSO") +
  theme_bw() +
  theme(aspect.ratio = 1,
        axis.text = element_text(color = "black"),
        axis.ticks = element_line(color = "black")) +
  facet_wrap(~sorted_gene_pair+compound_id, scales = "free")

```


See: https://dplyr.tidyverse.org/reference/group_map.html for more info on group_map() function
Also: https://stackoverflow.com/questions/56634033/get-summary-of-the-model-using-purrrmap-within-dplyr-piping

```{r}

## running lm for each sorted_gene_pair/compound_id group

d.DeKegel_top5_pairs_drug_sens_combined_cn_lm_summary <- d.DeKegel_top5_pairs_drug_sens_cn %>%
  filter(!is.na(dependency)) %>%
  filter(!is.na(combined_log_cn)) %>%
  group_by(sorted_gene_pair, compound_id) %>%
  group_modify(~ broom::glance(lm(dependency ~ combined_log_cn, data = .x))) %>%
  dplyr::ungroup() %>%
  rename(r_squared = r.squared, p_val = p.value)
d.DeKegel_top5_pairs_drug_sens_combined_cn_lm_summary

## add rsquared and p-value info to main DF
d.DeKegel_top5_pairs_drug_sens_combined_cn_lm <- d.DeKegel_top5_pairs_drug_sens_combined_cn_lm_summary %>% 
  dplyr::select(sorted_gene_pair:r_squared, p_val) %>%
  right_join(d.DeKegel_top5_pairs_drug_sens_cn, by = c("sorted_gene_pair", "compound_id"))

```

```{r}

## FDR adjust p-values and add back to DF
pvals <- d.DeKegel_top5_pairs_drug_sens_combined_cn_lm %>%
  distinct(sorted_gene_pair, compound_id, .keep_all = TRUE) %>%
  arrange(p_val) %>%
  dplyr::select(p_val) %>%
  pull()

fdr_vals <- p.adjust(pvals, method = "BH")
head(fdr_vals)

d.fdr_vals <- tibble(fdr = fdr_vals)

d.fdr_vals %>%
  filter(fdr < 0.05) %>%
  nrow()

d.fdr_vals %>%
  filter(fdr < 0.1) %>%
  nrow()

d.pvals_labels <- d.DeKegel_top5_pairs_drug_sens_combined_cn_lm %>%
  distinct(sorted_gene_pair, compound_id, .keep_all = TRUE) %>%
  arrange(p_val) %>%
  dplyr::select(sorted_gene_pair, compound_id)
d.pvals_labels

d.fdr_vals_labeled <- d.pvals_labels %>%
  add_column(d.fdr_vals)
d.fdr_vals_labeled

d.DeKegel_top5_pairs_drug_sens_combined_cn_lm <- d.DeKegel_top5_pairs_drug_sens_combined_cn_lm %>%
  left_join(d.fdr_vals_labeled, by = c("sorted_gene_pair", "compound_id"))
d.DeKegel_top5_pairs_drug_sens_combined_cn_lm

```

##### Plot
```{r}

# d.DeKegel_top5_pairs_drug_sens_combined_cn_lm

## filter DF for only pairs that I want to plot
d.drug_plot <- d.DeKegel_top5_pairs_drug_sens_combined_cn_lm %>% 
  ungroup() %>%
  filter(fdr < 0.1) %>%
  mutate(combined_cn = 2^combined_log_cn) %>%
  unite("gene_pair_compound_label", c("sorted_gene_pair", "compound_name"), sep = "\n", remove = FALSE) %>%
  mutate("gene_pair_compound_label" = reorder(gene_pair_compound_label, fdr),
         "sorted_gene_pair" = reorder(sorted_gene_pair, fdr))

## set length of DF to plot
length <- d.drug_plot %>%
  ungroup() %>%
  distinct(gene_pair_compound_label) %>%
  nrow()

## make df to label rsquared and pvalue for each plot
d.label <- d.drug_plot %>%
  group_by(sorted_gene_pair, compound_id) %>%
  mutate(group_label_x = min(combined_cn),
         group_label_y = max(dependency)) %>%
  dplyr::ungroup() %>%
  dplyr::select(gene_pair_compound_label, r_squared, fdr, group_label_x, group_label_y) %>%
  distinct(gene_pair_compound_label, .keep_all = TRUE) %>%
  mutate(r_squared = round(r_squared, 3),
         fdr = scientific(fdr, 3)) 
d.label

## make df to plot CN for each paralog
d.cn_plot <- d.drug_plot %>%
  dplyr::select(gene_pair_compound_label, A1_log_cn, A2_log_cn) %>%
  pivot_longer(cols = c(A1_log_cn, A2_log_cn),
               names_to = "pos",
               values_to = "log_cn") 

aspect_ratio <- 1
n_rows <- 1
n_cols <- 1
n_pages <- ceiling(length/(n_rows*n_cols))

# pdf(file.path(out_dir, paste(date, "d.DeKegel_top5_pairs_drug_sens_combined_cn_lm_fdr_L10.pdf", sep = "_")),
#     width = 6, height = 3)
# for(i in 1:n_pages){
#   p <-  ggplot(d.drug_plot, aes(x = combined_cn, y = dependency)) +
#     geom_hline(yintercept = 0) +
#     geom_point(size = 1) +
#     geom_smooth(method = lm, se = FALSE, color = "gray60") +
#     geom_richtext(data = d.label, 
#                   aes(x = group_label_x, y = group_label_y,
#                   ## this package can do HTML-formatting for text
#                   label = paste0("r<sup>2</sup>=", r_squared, "<br>FDR=", fdr)),
#                   ## adjust label alignment and remove borders
#                   hjust = 0, vjust = 1, label.color = NA, size = 3,
#                   ## label background = transparent, text is not
#                   fill = alpha(c("white"), 0.75)) + 
#     labs(x = "Combined_CN",
#          y = "LFC_treated_vs_DMSO") +
#     coord_trans(x = "log2") +
#     theme_bw() +
#     theme(aspect.ratio = aspect_ratio,
#           axis.text = element_text(color = "black"),
#           axis.ticks = element_line(color = "black")) +
#     facet_wrap_paginate(~gene_pair_compound_label, scales = "free",
#                         nrow = n_rows, ncol = n_cols, page = i)
# 
#   q <- ggplot(d.cn_plot, aes(x = pos, y = (2^log_cn))) +
#     geom_boxplot(fill = "gray", width = 0.7) +
#     labs(x = "Paralog",
#          y = "Copy_number") +
#     coord_trans(y = "log2") +
#     scale_x_discrete(labels = c("A1", "A2")) +
#     theme_bw() +
#     theme(aspect.ratio = aspect_ratio,
#           axis.text = element_text(color = "black"),
#           axis.ticks = element_line(color = "black")) +
#     facet_wrap_paginate(~gene_pair_compound_label, scales = "free_y",
#                         nrow = n_rows, ncol = n_cols, page = i)
#   print(ggarrange(p, q, ncol = 2, align = "hv"))
# }
# dev.off()

```

## Single-gene
```{r}

d.drug_sens_all_genes_cn

```
### Phase 1 drug sensitivity

#### Quartiles/WRST
```{r}

## bin combined CN for each pair into 4 quartiles
d.drug_sens_all_genes_cn_bin <- d.drug_sens_all_genes_cn %>%
  filter(!is.na(log_copy_number)) %>%
  filter(!is.na(dependency)) %>%
  group_by(target, compound_id) %>% # is this correct?
  mutate(log_cn_bin = ntile(log_copy_number, n = 4)) %>%
  ## factorized binned variable for easy plotting
  mutate(log_cn_bin = factor(log_cn_bin, levels = c(1, 2, 3, 4)))

## calculate the WRST p-value for bin 1 (lowest CN) vs. bin 4 (highest CN)
d.drug_sens_all_genes_cn_bin_pval_summary <- d.drug_sens_all_genes_cn_bin %>%
  filter(log_cn_bin %in% c(1, 4)) %>%
  group_by(target, compound_id) %>%
  summarize("p_val" = wilcox.test(dependency ~ log_cn_bin)$p.value)
d.drug_sens_all_genes_cn_bin_pval_summary

## add p-values to main DF
d.drug_sens_all_genes_cn_bin_wrst <- left_join(d.drug_sens_all_genes_cn_bin,
                                               d.drug_sens_all_genes_cn_bin_pval_summary,
                                               by = c("target", "compound_id")) %>%
  ungroup()
d.drug_sens_all_genes_cn_bin_wrst

```

```{r}

## FDR-adjust p-values
p_vals <- d.drug_sens_all_genes_cn_bin_wrst %>%
  distinct(target, compound_id, .keep_all = TRUE) %>%
  arrange(p_val) %>%
  dplyr::select(p_val) %>%
  pull()

fdr_vals <- p.adjust(p_vals, method = "BH")
head(fdr_vals)

d.fdr_vals <- tibble(fdr = fdr_vals)

d.fdr_vals %>%
  filter(fdr < 0.05) %>%
  nrow()

d.fdr_vals %>%
  filter(fdr < 0.1) %>%
  nrow()

d.p_vals_labels <- d.drug_sens_all_genes_cn_bin_wrst %>%
  distinct(target, compound_id, .keep_all = TRUE) %>%
  arrange(p_val) %>%
  dplyr::select(target, compound_id)
d.p_vals_labels

d.fdr_vals_labeled <- d.p_vals_labels %>%
  add_column(d.fdr_vals)
d.fdr_vals_labeled

d.drug_sens_all_genes_cn_bin_wrst <- d.drug_sens_all_genes_cn_bin_wrst %>%
  left_join(d.fdr_vals_labeled, by = c("target", "compound_id"))
d.drug_sens_all_genes_cn_bin_wrst

```

##### Plot
Plotting FDR values from DF directly onto ggsignif(): 
https://github.com/const-ae/ggsignif/issues/29
https://cran.r-project.org/web/packages/ggsignif/ggsignif.pdf

```{r}

## plot all FDR < 0.1 pairs (sorted by FDR)
## add my own labels!!

d.plot <- d.drug_sens_all_genes_cn_bin_wrst %>% 
  ungroup() %>%
  filter(fdr < 0.1) %>%
  unite("target_compound_label", c("target", "compound_name"), sep = "\n", remove = FALSE) %>%
  mutate("target_compound_label" = reorder(target_compound_label, fdr),
         "target" = reorder(target, fdr))

length <- d.plot %>%
  ungroup() %>%
  distinct(target_compound_label) %>%
  nrow()

# d.fdr_labels <- d.DeKegel_top5_combined_cn_drug_pair_fdr_signif %>%
#   ungroup() %>%
#   dplyr::select(gene_pair_compound_label, combined_cn_bin, fdr)
  
aspect_ratio <- 0.8
n_rows <- 1
n_cols <- 1
n_pages <- ceiling(length/(n_rows*n_cols))

# pdf(file.path(out_dir, paste(date, "drug_sens_all_genes_cn_bin_wrst_fdr_L10.pdf", sep = "_")),
#     width = 4, height = 2.5)
# for(i in 1:n_pages){
#   p <-  ggplot(d.plot, aes(x = log_cn_bin, y = dependency, fill = log_cn_bin)) +
#     geom_hline(yintercept = 0) +
#     geom_boxplot() +
#     geom_signif(comparisons = list(c(1, 4)),
#                 map_signif_level = FALSE,
#                 test = "wilcox.test",
#                 # test.args = list(alternative = "less", p.adjust.method = "BH"),
#                 # annotations = fdr_labels, # fix this to get FDR plotted direclty
#                 textsize = 3.2) +
#     # add_pvalue(d.fdr_labels) +
#     ## adds 10% of space on top & bottom so p-value label will fit
#     scale_y_continuous(expand = expansion(mult = 0.15)) +
#     scale_fill_brewer(palette = "YlGnBu") +
#     labs(x = "CN_bin",
#          y = "LFC_treated_vs_DMSO") +
#     theme_bw() +
#     theme(aspect.ratio = aspect_ratio,
#           axis.text = element_text(color = "black"),
#           axis.ticks = element_line(color = "black"),
#           legend.position = "none") +
#     facet_wrap_paginate(~target_compound_label, scales = "free_y",
#                         nrow = n_rows, ncol = n_cols, page = i)
# 
#   q <- ggplot(d.plot, aes(x = log_cn_bin, y = (2^log_copy_number), fill = log_cn_bin)) +
#     geom_hline(yintercept = 2) +
#     geom_hline(yintercept = 1, linetype = "dashed") +
#     geom_hline(yintercept = 3, linetype = "dashed") +
#     geom_boxplot() +
#     ## adds 10% of space on top & bottom so p-value label will fit
#     # scale_y_continuous(expand = expansion(mult = 0.15)) +
#     scale_fill_brewer(palette = "YlGnBu") +
#     labs(x = "CN_bin",
#          y = "CN") +
#     coord_trans(y = "log2") +
#     theme_bw() +
#     theme(aspect.ratio = aspect_ratio,
#           axis.text = element_text(color = "black"),
#           axis.ticks = element_line(color = "black"),
#           legend.position = "none") +
#     facet_wrap_paginate(~target_compound_label, scales = "free_y",
#                         nrow = n_rows, ncol = n_cols, page = i)
#   print(ggarrange(p, q, ncol = 2, align = "hv"))
# }
# dev.off()

```

#### Continuous/linear regression

```{r}

## running lm for each sorted_gene_pair/compound_id group

d.drug_sens_all_genes_cn_lm_summary <- d.drug_sens_all_genes_cn %>%
  filter(!is.na(dependency)) %>%
  filter(!is.na(log_copy_number)) %>%
  group_by(target, compound_id) %>%
  group_modify(~ broom::glance(lm(dependency ~ log_copy_number, data = .x))) %>%
  dplyr::ungroup() %>%
  rename(r_squared = r.squared, p_val = p.value)
d.drug_sens_all_genes_cn_lm_summary

## add rsquared and p-value info to main DF
d.drug_sens_all_genes_cn_lm <- d.drug_sens_all_genes_cn_lm_summary %>%
  dplyr::select(target:r_squared, p_val) %>%
  right_join(d.drug_sens_all_genes_cn, by = c("target", "compound_id"))

```

```{r}

## FDR adjust p-values and add back to DF
p_vals <- d.drug_sens_all_genes_cn_lm %>%
  distinct(target, compound_id, .keep_all = TRUE) %>%
  arrange(p_val) %>%
  dplyr::select(p_val) %>%
  pull()

fdr_vals <- p.adjust(p_vals, method = "BH")
head(fdr_vals)

d.fdr_vals <- tibble(fdr = fdr_vals)

d.fdr_vals %>%
  filter(fdr < 0.05) %>%
  nrow()

d.fdr_vals %>%
  filter(fdr < 0.1) %>%
  nrow()

d.p_vals_labels <- d.drug_sens_all_genes_cn_lm %>%
  distinct(target, compound_id, .keep_all = TRUE) %>%
  arrange(p_val) %>%
  dplyr::select(target, compound_id)
d.p_vals_labels

d.fdr_vals_labeled <- d.p_vals_labels %>%
  add_column(d.fdr_vals)
d.fdr_vals_labeled

d.drug_sens_all_genes_cn_lm <- d.drug_sens_all_genes_cn_lm %>%
  left_join(d.fdr_vals_labeled, by = c("target", "compound_id"))
d.drug_sens_all_genes_cn_lm

```

##### Plot
```{r}

# d.DeKegel_top5_pairs_drug_sens_combined_cn_lm

## filter DF for only pairs that I want to plot
d.drug_plot <- d.drug_sens_all_genes_cn_lm %>% 
  ungroup() %>%
  filter(fdr < 0.1) %>%
  # mutate(cn = 2^log_copy_number) %>%
  filter(!is.na(dependency)) %>%
  filter(!is.na(log_copy_number)) %>%
  unite("target_compound_label", c("target", "compound_name"), sep = "\n", remove = FALSE) %>%
  mutate("target_compound_label" = reorder(target_compound_label, fdr),
         "target" = reorder(target, fdr))

## set length of DF to plot
length <- d.drug_plot %>%
  ungroup() %>%
  distinct(target_compound_label) %>%
  nrow()

## make df to label rsquared and pvalue for each plot
d.label <- d.drug_plot %>%
  group_by(target, compound_id) %>%
  mutate(group_label_x = min(log_copy_number),
         group_label_y = max(dependency)) %>%
  dplyr::ungroup() %>%
  dplyr::select(target_compound_label, r_squared, fdr, group_label_x, group_label_y) %>%
  distinct(target_compound_label, .keep_all = TRUE) %>%
  mutate(r_squared = round(r_squared, 3),
         fdr = scientific(fdr, 3)) 
d.label

## make df to plot CN for each paralog
# d.cn_plot <- d.drug_plot %>%
#   dplyr::select(target_compound_label, log_copy_number) 

aspect_ratio <- 1
n_rows <- 1
n_cols <- 1
n_pages <- ceiling(length/(n_rows*n_cols))

# pdf(file.path(out_dir, paste(date, "drug_sens_all_genes_cn_lm_fdr_L10.pdf", sep = "_")),
#     width = 4, height = 3)
# for(i in 1:n_pages){
#   p <-  ggplot(d.drug_plot, aes(x = log_copy_number, y = dependency)) +
#     geom_hline(yintercept = 0) +
#     geom_point(size = 1) +
#     geom_smooth(method = lm, se = FALSE, color = "gray60") +
#     geom_richtext(data = d.label,
#                   aes(x = group_label_x, y = group_label_y,
#                   ## this package can do HTML-formatting for text
#                   label = paste0("r<sup>2</sup>=", r_squared, "<br>FDR=", fdr)),
#                   ## adjust label alignment and remove borders
#                   hjust = 0, vjust = 1, label.color = NA, size = 3,
#                   ## label background = transparent, text is not
#                   fill = alpha(c("white"), 0.75)) +
#     labs(x = "Log_copy_number",
#          y = "LFC_treated_vs_DMSO") +
#     # coord_trans(x = "log2") +
#     theme_bw() +
#     theme(aspect.ratio = aspect_ratio,
#           axis.text = element_text(color = "black"),
#           axis.ticks = element_line(color = "black")) +
#     facet_wrap_paginate(~target_compound_label, scales = "free",
#                         nrow = n_rows, ncol = n_cols, page = i)
# 
#   # q <- ggplot(d.cn_plot, aes(x = pos, y = (2^log_cn))) +
#   #   geom_boxplot(fill = "gray", width = 0.7) +
#   #   labs(x = "Paralog",
#   #        y = "Copy_number") +
#   #   coord_trans(y = "log2") +
#   #   scale_x_discrete(labels = c("A1", "A2")) +
#   #   theme_bw() +
#   #   theme(aspect.ratio = aspect_ratio,
#   #         axis.text = element_text(color = "black"),
#   #         axis.ticks = element_line(color = "black")) +
#   #   facet_wrap_paginate(~gene_pair_compound_label, scales = "free_y",
#   #                       nrow = n_rows, ncol = n_cols, page = i)
#   print(ggarrange(p, ncol = 1, align = "hv"))
# }
# dev.off()

```

```{r}
# 
# ## make volcano plot of r^2 vs. FDR...or something (how to indicate neg vs pos?)
# d.volcano_plot <- d.drug_sens_all_genes_cn_lm %>% 
#   ungroup() %>%
#   # mutate(cn = 2^log_copy_number) %>%
#   filter(!is.na(dependency)) %>%
#   filter(!is.na(log_copy_number)) %>%
#   unite("target_compound_label", c("target", "compound_name"), sep = "\n", remove = FALSE) %>%
#   mutate("target_compound_label" = reorder(target_compound_label, fdr),
#          "target" = reorder(target, fdr))
# 
# ggplot(d.volcano_plot, aes(x = r_squared, y = -log10(fdr))) +
#   geom_point()

```

```{r}

# phase1_cn_labels <- d.phase1_cn_volcano_plot %>%
#   filter(fdr < 0.1) %>%
#   pull(new_label)
# 
# ggplot(d.phase1_cn_volcano_plot, aes(x = lfc_dependency_Q1_v_Q4, y = -log10(fdr))) +
#   geom_hline(yintercept = 1, linetype = "dashed") +
#   geom_vline(xintercept = 0, linetype = "dashed") +
#   geom_point(color = ifelse(d.phase1_cn_volcano_plot$fdr < 0.1, "black", "gray50")) +
#   geom_label_repel(data = subset(d.phase1_cn_volcano_plot, fdr < 0.1),
#                    aes(label = phase1_cn_labels), 
#                    segment.size = 0.2,
#                    force = 0.5,
#                    size = 2.5, 
#                    alpha = 0.8, 
#                    max.overlaps = Inf) +
#   theme_bw() +
#   theme(aspect.ratio = 0.75,
#         axis.text = element_text(color = "black"),
#         axis.ticks = element_line(color = "black"))
# # ggsave(file.path(out_dir, "d.DeKegel_top5_combined_cn_drug_pair_median_dependency_lfc_fdr_volcano_plot.pdf"),
# #        width = 6, height = 4, useDingbats = FALSE)

```


### Z-scaled outliers

```{r}

## bin combined CN for each pair into 4 quartiles
d.drug_sens_all_genes_cn_outlier <- d.drug_sens_all_genes_cn %>%
  filter(!is.na(log_copy_number)) %>%
  filter(!is.na(dependency)) %>%
  group_by(target, compound_id) %>% # is this correct?
  mutate(scaled_log_cn = RobScale(log_copy_number, center = TRUE, scale = TRUE)) %>%
  mutate(cn_outlier_flag = case_when(
    scaled_log_cn < -2 ~ "low",
    scaled_log_cn > 2 ~ "high", 
    TRUE ~ "neither")) %>%
  mutate(cn_low_flag = case_when(
    scaled_log_cn < -2 ~ TRUE,
    TRUE ~ FALSE)) %>% ## if CN ≥ -2, classify as "not low"
  ungroup()

d.drug_sens_all_genes_cn_outlier %>%
  group_by(cn_outlier_flag) %>%
  summarize(n = n())

d.drug_sens_all_genes_cn_outlier %>%
  group_by(cn_low_flag) %>%
  summarize(n = n())

d.drug_sens_all_genes_cn_outlier %>%
  group_by(target, compound_id) %>%
  distinct(cn_low_flag) %>%
  summarize(n = n()) %>%
  filter(n < 2)

## calculate the WRST p-value for low-CN vs. not low-CN cell lines
d.drug_sens_all_genes_cn_outlier_wrst_summary <- d.drug_sens_all_genes_cn_outlier %>%
  group_by(target, compound_id) %>%
  filter(any(cn_low_flag == TRUE)) %>% ## keep only groups where >1 cell line has low CN
  summarize("p_val" = wilcox.test(dependency ~ cn_low_flag)$p.value)
d.drug_sens_all_genes_cn_outlier_wrst_summary

## add p-values to main DF
d.drug_sens_all_genes_cn_outlier_wrst <- left_join(d.drug_sens_all_genes_cn_outlier,
                                                   d.drug_sens_all_genes_cn_outlier_wrst_summary,
                                                   by = c("target", "compound_id")) %>%
  ungroup()
d.drug_sens_all_genes_cn_outlier_wrst


## from script: CRISPR_screens/02_iCas9-HeLa/04_NGS_analysis/mageck/00_scripts/06_SL_paralog_druggability_and_expression.Rmd

# ## calculate z-score for expression using scale() function
# d.paralog.tpm.scaled <- d.paralog.tpm %>%
#   group_by(gene_name, primary_disease) %>%
#   mutate(tissue_expr_scaled=RobScale(expression, center=TRUE, scale=TRUE)) %>% # calculate z-score using scale()
#   mutate(tissue_expr_outlier_broad=ifelse(abs(tissue_expr_scaled) > 2, TRUE, FALSE)) %>%
#   mutate(tissue_expr_outlier_narrow=ifelse(is.na(tissue_expr_scaled), "NA", 
#                                            ifelse(tissue_expr_scaled < -2, "negative",
#                                                   ifelse(tissue_expr_scaled > 2, "positive", "neither")))) %>%
#   ungroup()

```

```{r}

## FDR adjust p-values and add back to DF
p_vals <- d.drug_sens_all_genes_cn_outlier_wrst %>%
  distinct(target, compound_id, .keep_all = TRUE) %>%
  arrange(p_val) %>%
  dplyr::select(p_val) %>%
  pull()

fdr_vals <- p.adjust(p_vals, method = "BH")
head(fdr_vals)

d.fdr_vals <- tibble(fdr = fdr_vals)

d.fdr_vals %>%
  filter(fdr < 0.05) %>%
  nrow()

d.fdr_vals %>%
  filter(fdr < 0.1) %>%
  nrow()

d.p_vals_labels <- d.drug_sens_all_genes_cn_outlier_wrst %>%
  distinct(target, compound_id, .keep_all = TRUE) %>%
  arrange(p_val) %>%
  dplyr::select(target, compound_id)
d.p_vals_labels

d.fdr_vals_labeled <- d.p_vals_labels %>%
  add_column(d.fdr_vals)
d.fdr_vals_labeled

d.drug_sens_all_genes_cn_outlier_wrst <- d.drug_sens_all_genes_cn_outlier_wrst %>%
  left_join(d.fdr_vals_labeled, by = c("target", "compound_id"))
d.drug_sens_all_genes_cn_outlier_wrst

```

#### Plot
```{r}

## plot all p < 0.001 pairs (sorted by p-val)
## add my own labels!!

d.plot <- d.drug_sens_all_genes_cn_outlier_wrst %>% 
  ungroup() %>%
  filter(p_val < 0.001) %>%
  unite("target_compound_label", c("target", "compound_name"), sep = "\n", remove = FALSE) %>%
  mutate("target_compound_label" = reorder(target_compound_label, fdr),
         "target" = reorder(target, fdr)) %>%
  mutate(cn_low_flag = factor(cn_low_flag, levels = c(TRUE, FALSE)))

length <- d.plot %>%
  ungroup() %>%
  distinct(target_compound_label) %>%
  nrow()

# d.cn_plot <- d.plot %>%
#   dplyr::select(target_compound_label, target, log_copy_number, cn_low_flag) 
  
aspect_ratio <- 0.8
n_rows <- 1
n_cols <- 1
n_pages <- ceiling(length/(n_rows*n_cols))

# pdf(file.path(out_dir, paste(date, "drug_sens_all_genes_cn_outlier_wrst_p_L0.001.pdf", sep = "_")),
#     width = 6, height = 3)
# for(i in 1:n_pages){
#   p <-  ggplot(d.plot, aes(x = cn_low_flag, y = dependency)) +
#     geom_hline(yintercept = 0) +
#     geom_jitter(aes(color = cn_low_flag), width = 0.3) +
#     geom_boxplot(alpha = 0) +
#     geom_signif(comparisons = list(c(1, 2)),
#                 map_signif_level = FALSE,
#                 test = "wilcox.test",
#                 # test.args = list(alternative = "less", p.adjust.method = "BH"),
#                 # annotations = fdr_labels, # fix this to get FDR plotted direclty
#                 textsize = 3.2) +
#     # add_pvalue(d.fdr_labels) +
#     ## adds 10% of space on top & bottom so p-value label will fit
#     scale_y_continuous(expand = expansion(mult = 0.15)) +
#     scale_color_manual(values = c("blue", "gray50")) +
#     labs(x = "CN_outlier_low",
#          y = "LFC_treated_vs_DMSO") +
#     theme_bw() +
#     theme(aspect.ratio = aspect_ratio,
#           axis.text = element_text(color = "black"),
#           axis.ticks = element_line(color = "black"),
#           legend.position = "none") +
#     facet_wrap_paginate(~target_compound_label, scales = "free_y",
#                         nrow = n_rows, ncol = n_cols, page = i)
# 
#   q <- ggplot(d.plot, aes(x = target, y = (2^log_copy_number))) +
#     geom_hline(yintercept = 2) +
#     geom_hline(yintercept = 1, linetype = "dashed") +
#     geom_hline(yintercept = 3, linetype = "dashed") +
#     geom_jitter(aes(color = cn_low_flag), width = 0.3) +
#     geom_boxplot(alpha = 0) +
#     ## adds 10% of space on top & bottom so p-value label will fit
#     # scale_y_continuous(expand = expansion(mult = 0.15)) +
#     scale_color_manual(values = c("blue", "gray50")) +
#     labs(x = "target", y = "CN") +
#     coord_trans(y = "log2") +
#     theme_bw() +
#     theme(aspect.ratio = aspect_ratio,
#           axis.text = element_text(color = "black"),
#           axis.ticks = element_line(color = "black")) +
#     facet_wrap_paginate(~target_compound_label, scale = "free",
#                         nrow = n_rows, ncol = n_cols, page = i)
#   print(ggarrange(p, q, ncol = 2, align = "hv", common.legend = TRUE, legend = "right"))
# }
# dev.off()

```

```{r}

d.plot <- d.drug_sens_all_genes_cn_outlier_wrst %>% 
  ungroup() %>%
  filter(p_val < 0.001) %>%
  unite("target_compound_label", c("target", "compound_name"), sep = "\n", remove = FALSE) %>%
  mutate("target_compound_label" = reorder(target_compound_label, fdr),
         "target" = reorder(target, fdr)) %>%
  filter(target_compound_label == "GBF1\ngolgicide-a" | target_compound_label == "ASPH\nsuccinic-acid" |
           target_compound_label == "PDE4A\ndipyridamole")
d.plot

ggplot(d.plot, aes(x = target, y = (2^log_copy_number))) +
    geom_hline(yintercept = 2) +
    geom_hline(yintercept = 1, linetype = "dashed") +
    geom_hline(yintercept = 3, linetype = "dashed") +
    geom_jitter(aes(color = cn_low_flag), width = 0.3) +
    geom_boxplot(alpha = 0) +
    ## adds 10% of space on top & bottom so p-value label will fit
    # scale_y_continuous(expand = expansion(mult = 0.15)) +
    scale_color_manual(values = c("darkgray", "blue")) +
    labs(x = "target",
         y = "CN") +
    coord_trans(y = "log2") +
    theme_bw() +
    theme(aspect.ratio = 1.25,
          axis.text = element_text(color = "black"),
          axis.ticks = element_line(color = "black")) +
    facet_wrap(~target, scale = "free")

```

#### + and - outliers
```{r}

d.drug_sens_all_genes_cn_outlier <- d.drug_sens_all_genes_cn_outlier %>%
  mutate(cn_outlier_flag = factor(cn_outlier_flag, levels = c("low", "neither", "high"))) 

```


```{r}


d.drug_sens_all_genes_cn_outlier %>%
  group_by(target, compound_id, cn_outlier_flag) %>%
  summarize(n = n()) %>%
  group_by(cn_outlier_flag) %>%
  summarize(mean = mean(n),
            median = median(n))

## how many compounds/targets have 0 high or low outlier cell lines?
d.drug_sens_all_genes_cn_outlier %>%
  group_by(target, compound_id, cn_outlier_flag) %>%
  summarize(n = n()) 

d.drug_sens_all_genes_cn_outlier %>%
  group_by(target, compound_id, cn_outlier_flag) %>%
  summarize(n = n()) %>%
  ungroup() %>%
  group_by(target, compound_id) %>%
  summarize(n = n())
  

```

```{r}

## low vs. neither (with high included but ignoring it) => 
## calculate diff in median LFC between relevant groups

d.drug_sens_all_genes_cn_outlier_low <- d.drug_sens_all_genes_cn_outlier %>%
  group_by(target, compound_id) %>%
  ## get rid of drug/targets that don't have any "low" cell lines
  filter(any(cn_outlier_flag == "low")) %>%
  ungroup() 

d.drug_sens_all_genes_cn_outlier_low_grp_med_diff <- d.drug_sens_all_genes_cn_outlier_low %>%
  group_by(target, compound_id, cn_outlier_flag) %>%
  ## calculate median dependency score for each outlier group
  summarize(grp_med_dependency = median(dependency)) %>%
  pivot_wider(names_from = cn_outlier_flag, values_from = grp_med_dependency) %>%
  ## calculate difference between low and neither groups
  mutate(grp_med_diff = low - neither) %>%
  dplyr::select(target, compound_id, grp_med_diff) 

d.drug_sens_all_genes_cn_outlier_low <- d.drug_sens_all_genes_cn_outlier_low %>%
  left_join(d.drug_sens_all_genes_cn_outlier_low_grp_med_diff, by = c("target", "compound_id"))

## high vs. neither (with low included but ignoring it) => 
## calculate diff in median LFC between relevant groups =>
d.drug_sens_all_genes_cn_outlier_high <- d.drug_sens_all_genes_cn_outlier %>%
  group_by(target, compound_id) %>%
  ## get rid of drug/targets that don't have any "high" cell lines
  filter(any(cn_outlier_flag == "high")) %>%
  ungroup() 

d.drug_sens_all_genes_cn_outlier_high_grp_med_diff <- d.drug_sens_all_genes_cn_outlier_high %>%
  group_by(target, compound_id, cn_outlier_flag) %>%
  ## calculate median dependency score for each outlier group
  summarize(grp_med_dependency = median(dependency)) %>%
  pivot_wider(names_from = cn_outlier_flag, values_from = grp_med_dependency) %>%
  ## calculate difference between low and neither groups
  mutate(grp_med_diff = high - neither) %>%
  dplyr::select(target, compound_id, grp_med_diff) 

d.drug_sens_all_genes_cn_outlier_high <- d.drug_sens_all_genes_cn_outlier_high %>%
  left_join(d.drug_sens_all_genes_cn_outlier_high_grp_med_diff, by = c("target", "compound_id"))

```

```{r}

## low outliers and drug sensitivity
## calculate p-value (adjusted for multiple within-group tests?)
d.drug_sens_all_genes_cn_outlier_low_wrst_sens_pvals <- d.drug_sens_all_genes_cn_outlier_low %>%
  ## remove cell lines that are high outliers (I think this makes sense... but I'm not sure)
  filter(cn_outlier_flag != "high" & grp_med_diff < 0) %>%
  group_by(target, compound_id) %>%
  summarize(p_val = wilcox.test(dependency ~ cn_outlier_flag, alternative = "less")$p.value)
# d.drug_sens_all_genes_cn_outlier_low_wrst_sens_pvals

## adjust for multiple testing overall
d.stats <- BH_adjust(d.drug_sens_all_genes_cn_outlier_low_wrst_sens_pvals)
# d.stats

d.drug_sens_all_genes_cn_outlier_low_wrst_sens <- d.drug_sens_all_genes_cn_outlier_low %>%
  left_join(d.stats, by = c("target", "compound_id"))
# d.drug_sens_all_genes_cn_outlier_low_wrst_sens

## low outliers and drug resistance
## calculate p-value
d.drug_sens_all_genes_cn_outlier_low_wrst_resist_pvals <- d.drug_sens_all_genes_cn_outlier_low %>%
  ## remove cell lines that are high outliers (I think this makes sense... but I'm not sure)
  filter(cn_outlier_flag != "high" & grp_med_diff > 0) %>%
  group_by(target, compound_id) %>%
  summarize(p_val = wilcox.test(dependency ~ cn_outlier_flag, alternative = "greater")$p.value)
# d.drug_sens_all_genes_cn_outlier_low_wrst_resist_pvals

## adjust for multiple testing overall
d.stats <- BH_adjust(d.drug_sens_all_genes_cn_outlier_low_wrst_resist_pvals)
# d.stats

d.drug_sens_all_genes_cn_outlier_low_wrst_resist <- d.drug_sens_all_genes_cn_outlier_low %>%
  left_join(d.stats, by = c("target", "compound_id"))

# d.drug_sens_all_genes_cn_outlier_low_wrst_resist

## high outliers and drug sensitivity
d.drug_sens_all_genes_cn_outlier_high_wrst_sens_pvals <- d.drug_sens_all_genes_cn_outlier_high %>%
  ## remove cell lines that are high outliers (I think this makes sense... but I'm not sure)
  filter(cn_outlier_flag != "low" & grp_med_diff < 0) %>%
  group_by(target, compound_id) %>%
  summarize(p_val = wilcox.test(dependency ~ cn_outlier_flag, alternative = "greater")$p.value)
# d.drug_sens_all_genes_cn_outlier_high_wrst_sens_pvals

## adjust for multiple testing overall
d.stats <- BH_adjust(d.drug_sens_all_genes_cn_outlier_high_wrst_sens_pvals)
# d.stats

d.drug_sens_all_genes_cn_outlier_high_wrst_sens <- d.drug_sens_all_genes_cn_outlier_high %>%
  left_join(d.stats, by = c("target", "compound_id"))
# d.drug_sens_all_genes_cn_outlier_high_wrst_sens

## high outliers and drug resistance
d.drug_sens_all_genes_cn_outlier_high_wrst_resist_pvals <- d.drug_sens_all_genes_cn_outlier_high %>%
  ## remove cell lines that are high outliers (I think this makes sense... but I'm not sure)
  filter(cn_outlier_flag != "low" & grp_med_diff > 0) %>%
  group_by(target, compound_id) %>%
  summarize(p_val = wilcox.test(dependency ~ cn_outlier_flag, alternative = "less")$p.value)
# d.drug_sens_all_genes_cn_outlier_high_wrst_resist_pvals

## adjust for multiple testing overall
d.stats <- BH_adjust(d.drug_sens_all_genes_cn_outlier_high_wrst_resist_pvals)
# d.stats

d.drug_sens_all_genes_cn_outlier_high_wrst_resist <- d.drug_sens_all_genes_cn_outlier_high %>%
  left_join(d.stats, by = c("target", "compound_id"))
# d.drug_sens_all_genes_cn_outlier_high_wrst_resist


```

```{r}

## low outliers
summarize_stats(d.drug_sens_all_genes_cn_outlier_low_wrst_sens)
summarize_stats(d.drug_sens_all_genes_cn_outlier_low_wrst_resist)

## high outliers
summarize_stats(d.drug_sens_all_genes_cn_outlier_high_wrst_sens)
summarize_stats(d.drug_sens_all_genes_cn_outlier_high_wrst_resist)

```

For deparsing variables to use in dplyr: https://cran.r-project.org/web/packages/dplyr/vignettes/programming.html
```{r}

make_outlier_plot <- function(df, direction, col_name, stat_cutoff){
  
  ## save col_name as a string for later use
  stat_name <- deparse(substitute(col_name))

  d.plot <- df %>% 
    ungroup() %>%
    ##use {{}} to deparse variable name for use in dplyr functions
    filter({{col_name}} < stat_cutoff) %>%
    unite("target_compound_label", c("target", "compound_name"), sep = "\n", remove = FALSE) %>%
    mutate("target_compound_label" = reorder(target_compound_label, {{col_name}}),
           "target" = reorder(target, {{col_name}})) #%>%

  ## make df to label adjusted FDR/p-val for each plot
  d.label <- d.plot %>%
    group_by(target, compound_id) %>%
    mutate(y.position = (max(dependency) + 0.1*(max(dependency) - min(dependency)))) %>%
    dplyr::ungroup() %>%
    dplyr::select(target_compound_label, p_val, fdr, y.position) %>%
    distinct(target_compound_label, .keep_all = TRUE) %>%
    mutate(group1 = direction,
           group2 = "neither",
           p_val = scientific(p_val, 3),
           fdr = scientific(fdr, 3)) 
  d.label
  
  length <- d.plot %>%
    ungroup() %>%
    distinct(target_compound_label) %>%
    nrow()
    
  aspect_ratio <- 1
  n_rows <- 1
  n_cols <- 1
  n_pages <- ceiling(length/(n_rows*n_cols))
  
  ## make stat label for labeling significance brackets
  stat_label <- paste0(toupper(stat_name), " = ", "{", stat_name, "}")

  ## remove the first 2 characters of dataframe name
  df_name <- str_sub(deparse(substitute(df)), start = 3)

  ## save output file name
  file_name <- paste(date, df_name, stat_name, "L", stat_cutoff, ".pdf", sep = "_")

  pdf(file.path(out_dir, file_name), width = 6, height = 3)
  for(i in 1:n_pages){
    p <-  ggplot(d.plot, aes(x = cn_outlier_flag, y = dependency)) +
      geom_hline(yintercept = 0) +
      geom_jitter(aes(color = cn_outlier_flag), width = 0.3) +
      geom_boxplot(alpha = 0) +
      stat_pvalue_manual(d.label, label = stat_label, tip.length = 0.03, size = 3, vjust = -0.5) +
      ## adds extra space on top & bottom so p-value label will fit
      scale_y_continuous(expand = expansion(mult = 0.15)) +
      scale_color_manual(values = c("low" = "blue", "neither" = "gray50", "high" = "red")) +
      labs(x = "CN_group",
           y = "LFC_treated_vs_DMSO") +
      theme_bw() +
      theme(aspect.ratio = aspect_ratio,
            axis.text = element_text(color = "black"),
            axis.ticks = element_line(color = "black"),
            legend.position = "none") +
      facet_wrap_paginate(~target_compound_label, scales = "free_y",
                          nrow = n_rows, ncol = n_cols, page = i)

    q <- ggplot(d.plot, aes(x = target, y = (2^log_copy_number))) +
      geom_jitter(aes(color = cn_outlier_flag), width = 0.3) +
      geom_boxplot(alpha = 0) +
      scale_color_manual(values = c("low" = "blue", "neither" = "gray50", "high" = "red")) +
      labs(x = "Target", y = "Copy_number") +
      coord_trans(y = "log2") +
      theme_bw() +
      theme(aspect.ratio = aspect_ratio,
            axis.text = element_text(color = "black"),
            axis.ticks = element_line(color = "black")) +
      facet_wrap_paginate(~target_compound_label, scale = "free",
                          nrow = n_rows, ncol = n_cols, page = i)
    print(ggarrange(p, q, ncol = 2, align = "hv", common.legend = TRUE, legend = "right"))
  }
  dev.off()
  
}


```



```{r}

# d.drug_sens_all_genes_cn_outlier_low_wrst_sens

# make_outlier_plot(d.drug_sens_all_genes_cn_outlier_low_wrst_sens, "low", fdr, 0.1)

# d.drug_sens_all_genes_expr_outlier_low_wrst_resist = 0 FDR < 0.1 - skip

# make_outlier_plot(d.drug_sens_all_genes_cn_outlier_high_wrst_sens, "high", fdr, 0.1)

# make_outlier_plot(d.drug_sens_all_genes_cn_outlier_high_wrst_resist, "high", fdr, 0.1)

```






