---
title: "pgPEN_DeKegel_comparison"
author: "Phoebe Parrish"
date: "2023-09-12"
output: html_document
---

# To Do
* get rid of all_genes code for everything except EGFR inhibitors and other known examples
* update the renv lock file


# Setup

```{r setup, include=FALSE}

# library(biomaRt)
# library(depmap)
# library(ExperimentHub)
library(tidyverse)
library(ggrepel)
library(ggforce)
library(ggsignif)
library(RColorBrewer)
library(ggpubr) # another way to do significance bars
library(ggprism) # another significance bar package...
library(scales) # scientific notation
library(ggtext) # for using superscript in labels
library(DescTools) # for robust scaling
library(kableExtra)

```


### Assign variables
```{r}

date <- Sys.Date()

in_dir <- file.path("..", "results/integrated_dfs")

out_dir <- file.path("..", "results/expression_output")

```


### Import files

```{r, message = FALSE}

d.DeKegel_top5_pairs_drug_sens_expr_annot <- read_rds(file.path(in_dir, "d.DeKegel_top5_pairs_drug_sens_expr_annot"))

d.drug_sens_all_genes_expr <- read_rds(file.path(in_dir, "d.drug_sens_all_genes_expr"))

# d.Broad_drug_info <- read_csv(file.path(in_dir, "primary-screen-replicate-collapsed-treatment-info.csv"))
# d.Broad_secondary_auc <- read_csv(file.path(in_dir, "secondary-screen-dose-response-curve-parameters.csv"))
# d.Broad_secondary_doses <- read_csv(file.path(in_dir, "secondary-screen-dose-response-curve-parameters.csv"))

```

#### Reformat DFs
```{r}

## keep only relevant columns
d.DeKegel_top5_pairs_drug_sens_expr <- d.DeKegel_top5_pairs_drug_sens_expr_annot %>%
  dplyr::select(-c(entrez_pair:prediction_rank, prediction_score:family_size,
                   broad_id, primary_disease, subtype_disease))

d.drug_sens_all_genes_expr <- d.drug_sens_all_genes_expr %>%
  rename(log_tpm = rna_expression)

## how many gene pairs are included in the paralog dataset?
d.DeKegel_top5_pairs_drug_sens_expr %>%
  distinct(sorted_gene_pair) %>%
  nrow()

```

### Import functions

#### Stats
```{r}

source("shared_functions.R")

```


# Analysis

## Single-gene

This analysis is intended to demonstrate expected results for targeted inhibitors and their target (single) genes. 

### PRISM Phase 1 drug sensitivity

#### Continuous/linear regression

```{r}

## running lm for each sorted_gene_pair/compound_id group

## use broom::glance to get rsquared and pval
d.drug_sens_all_genes_expr_lm_summary <- d.drug_sens_all_genes_expr %>%
  filter(!is.na(dependency)) %>%
  filter(!is.na(log_tpm)) %>%
  group_by(target, compound_id) %>%
  group_modify(~ broom::glance(lm(dependency ~ log_tpm, data = .x))) %>%
  dplyr::ungroup() %>%
  rename(r_squared = r.squared, p_val = p.value)
# d.drug_sens_all_genes_expr_lm_summary

## FDR-adjust p-vals for multiple testing and add back to summary DF
d.stats <- BH_adjust(d.drug_sens_all_genes_expr_lm_summary)
# d.stats

d.drug_sens_all_genes_expr_lm_summary <- d.stats %>%
  dplyr::select(-p_val) %>%
  right_join(d.drug_sens_all_genes_expr_lm_summary, by = c("target", "compound_id")) %>%
  dplyr::select(target, compound_id, r_squared, p_val, fdr)
# d.drug_sens_all_genes_expr_lm_summary

## use broom::tidy to get slope
d.drug_sens_all_genes_expr_lm_slope <- d.drug_sens_all_genes_expr %>%
  filter(!is.na(dependency)) %>%
  filter(!is.na(log_tpm)) %>%
  group_by(target, compound_id) %>%
  group_modify(~ broom::tidy(lm(dependency ~ log_tpm, data = .x))) %>%
  dplyr::ungroup() %>%
  ## extract slope of lm fit line from output
  filter(term == "log_tpm") %>%
  rename(slope = estimate) %>%
  dplyr::select(target, compound_id, slope) 
# d.drug_sens_all_genes_expr_lm_slope

## add rsquared, p-val, fdr, and slope info to main DF
d.drug_sens_all_genes_expr_lm <- d.drug_sens_all_genes_expr %>%
  left_join(d.drug_sens_all_genes_expr_lm_summary, by = c("target", "compound_id")) %>%
  left_join(d.drug_sens_all_genes_expr_lm_slope, by = c("target", "compound_id")) %>%
  filter(!is.na(dependency)) %>%
  filter(!is.na(log_tpm))

summarize_stats(d.drug_sens_all_genes_expr_lm)

```



```{r}

## number of pairs with negative vs. positive slope
d.drug_sens_all_genes_expr_lm %>%
  distinct(target, compound_id, .keep_all = TRUE) %>%
  mutate(slope_flag = ifelse(slope < 0, "negative", "positive")) %>%
  group_by(slope_flag) %>%
  summarize(n = n())

## get low expression = low LFC pairs (positive slope)
d.drug_sens_all_genes_expr_lm_pos_slope <-d.drug_sens_all_genes_expr_lm %>%
  filter(slope > 0)

## get high expression = low LFC pairs (negative slope)
d.drug_sens_all_genes_expr_lm_neg_slope <- d.drug_sens_all_genes_expr_lm %>%
  filter(slope < 0)

## summarize significant drug/gene combos for each new DF
summarize_stats(d.drug_sens_all_genes_expr_lm_pos_slope)
summarize_stats(d.drug_sens_all_genes_expr_lm_neg_slope)

```

##### Plot
###### example genes
```{r}

plot_list <- d.drug_sens_all_genes_expr_lm_pos_slope %>%
  unite("target_compound_label", c("target", "compound_name"), sep = "\n", remove = FALSE) %>%
  filter(fdr < 0.05) %>%
  mutate("target_compound_label" = reorder(target_compound_label, fdr),
         "target" = reorder(target, fdr)) %>%
  arrange(fdr) %>%
  distinct(target_compound_label) %>%
  head(10) %>%
  pull()

d.plot <- d.drug_sens_all_genes_expr_lm_pos_slope %>%
  unite("target_compound_label", c("target", "compound_name"), sep = "\n", remove = FALSE) %>%
  filter(target_compound_label %in% plot_list) 

ggplot(d.plot, aes(x= log_tpm, y = dependency)) +
      geom_point(size = 1) +
      geom_smooth(method = lm, se = FALSE, color = "gray60") +
      # geom_richtext(data = d.label,
      #               aes(x = group_label_x, y = group_label_y,
      #               ## this package can do HTML-formatting for text
      #               label = paste0("r<sup>2</sup>=", r_squared, "<br>", 
      #                              toupper(stat_name), "=", {{col_name}})),
      #               ## adjust label alignment and remove borders
      #               hjust = 0, vjust = 1, label.color = NA, size = 3,
      #               ## label background = transparent, text is not
      #               fill = alpha(c("white"), 0.75),
      #               label.margin = unit(c(0, 0, 0, 0), "lines"),
      #               label.padding = unit(c(0, 0, 0, 0), "lines")) +
      # labs(x = x_title, y = y_title) +
      theme_bw() +
      theme(aspect.ratio = 1,
            axis.text = element_text(color = "black"),
            axis.ticks = element_line(color = "black")) +
      facet_wrap_paginate(~target_compound_label)
```

###### all genes
```{r}

make_lm_plot <- function(df, x_var, x_title, y_var, y_title, col_name, stat_cutoff, plot_n){
  ## save col_name as a string for later
  stat_name <- deparse(substitute(col_name))
  
  ## remove the first 2 characters of dataframe name
  df_name <- str_sub(deparse(substitute(df)), start = 3)
  
  ## filter DF for only pairs that I want to plot
  d.drug_plot <- df %>% 
    ungroup() %>%
    ##use {{}} to deparse variable name for use in dplyr functions
    filter({{col_name}} < stat_cutoff) %>%
    unite("target_compound_label", c("target", "compound_name"), sep = "\n", remove = FALSE) %>%
    mutate("target_compound_label" = reorder(target_compound_label, {{col_name}}),
           "target" = reorder(target, {{col_name}}))
  
  if(plot_n == "all") {
    ## do nothing to the df
    ## save output file name
    file_name <- paste0(date, "_", df_name, "_", stat_name, "_L", stat_cutoff, "all.pdf")
    
  } else {
    plot_pairs <- d.drug_plot %>%
      arrange({{col_name}}) %>%
      distinct(target_compound_label) %>%
      head(n = plot_n) %>%
      pull(target_compound_label)
    
    d.drug_plot <- d.drug_plot %>%
      filter(target_compound_label %in% plot_pairs)
    
    ## save output file name
    file_name <- paste0(date, "_", df_name, "_", stat_name, "_L", stat_cutoff, "_top", plot_n, ".pdf")
  }

  ## set length of DF to plot
  length <- d.drug_plot %>%
    ungroup() %>%
    distinct(target_compound_label) %>%
    nrow()
  
  ## make df to label rsquared and pvalue for each plot
  d.label <- d.drug_plot %>%
    group_by(target, compound_id) %>%
    mutate(group_label_x = min(log_tpm),
           group_label_y = max(dependency)) %>%
    dplyr::ungroup() %>%
    dplyr::select(target_compound_label, r_squared, p_val, fdr, group_label_x, group_label_y) %>%
    distinct(target_compound_label, .keep_all = TRUE) %>%
    mutate(r_squared = round(r_squared, 3),
           p_val = scientific(p_val, 3),
           fdr = scientific(fdr, 3)) 
  d.label
  
  aspect_ratio <- 1
  n_rows <- 1
  n_cols <- 1
  n_pages <- ceiling(length/(n_rows*n_cols))

  pdf(file.path(out_dir, file_name), width = 4, height = 3) 
  # label = stat_label
  for(i in 1:n_pages){
    p <-  ggplot(d.drug_plot, aes_string(x = x_var, y = y_var)) +
      geom_hline(yintercept = 0) +
      geom_point(size = 1) +
      geom_smooth(method = lm, se = FALSE, color = "gray60") +
      geom_richtext(data = d.label,
                    aes(x = group_label_x, y = group_label_y,
                    ## this package can do HTML-formatting for text
                    label = paste0("r<sup>2</sup>=", r_squared, "<br>", 
                                   toupper(stat_name), "=", {{col_name}})),
                    ## adjust label alignment and remove borders
                    hjust = 0, vjust = 1, label.color = NA, size = 3,
                    ## label background = transparent, text is not
                    fill = alpha(c("white"), 0.75),
                    label.margin = unit(c(0, 0, 0, 0), "lines"),
                    label.padding = unit(c(0, 0, 0, 0), "lines")) +
      labs(x = x_title, y = y_title) +
      theme_bw() +
      theme(aspect.ratio = aspect_ratio,
            axis.text = element_text(color = "black"),
            axis.ticks = element_line(color = "black")) +
      facet_wrap_paginate(~target_compound_label, scales = "free",
                          nrow = n_rows, ncol = n_cols, page = i)
    print(ggarrange(p, ncol = 1, align = "hv"))
  }
  dev.off()
  
}

```



```{r}

## run make_lm_plot function

# make_lm_plot(d.drug_sens_all_genes_expr_lm_pos_slope, 
#              "log_tpm", "log2(TPM)",
#              "dependency", "LFC_treated_vs_DMSO",
#              fdr, 0.1, 50)
# 
# make_lm_plot(d.drug_sens_all_genes_expr_lm_neg_slope, 
#              "log_tpm", "log2(TPM)",
#              "dependency", "LFC_treated_vs_DMSO", 
#              fdr, 0.1, 50)

```


```{r}

d.drug_sens_all_genes_expr_lm_pos_slope_fdr_L0.1 <- d.drug_sens_all_genes_expr_lm_pos_slope %>% 
  filter(fdr < 0.1) %>%
  arrange(fdr) %>%
  distinct(target, compound_id, .keep_all = TRUE) %>%
  dplyr::select(target, compound_name, slope, r_squared:fdr, moa)
print_kbl(head(d.drug_sens_all_genes_expr_lm_pos_slope_fdr_L0.1))

# write_tsv(d.drug_sens_all_genes_expr_lm_pos_slope_fdr_L0.1,
#           file.path(out_dir, "d.drug_sens_all_genes_expr_lm_pos_slope_fdr_L0.1.txt"))

d.drug_sens_all_genes_expr_lm_neg_slope_fdr_L0.1 <- d.drug_sens_all_genes_expr_lm_neg_slope %>% 
  filter(fdr < 0.1) %>%
  arrange(fdr) %>%
  distinct(target, compound_id, .keep_all = TRUE) %>%
  dplyr::select(target, compound_name, slope, r_squared:fdr, moa)
print_kbl(head(d.drug_sens_all_genes_expr_lm_neg_slope_fdr_L0.1))

# write_tsv(d.drug_sens_all_genes_expr_lm_neg_slope_fdr_L0.1, 
#           file.path(out_dir, "d.drug_sens_all_genes_expr_lm_neg_slope_fdr_L0.1.txt"))

```

```{r}

make_lm_lineage_plot <- function(df, x_var, x_title, y_var, y_title, col_name, stat_cutoff, plot_n){
  
  ## save col_name as a string for later use
  stat_name <- deparse(substitute(col_name))
  
  ## remove the first 2 characters of dataframe name
  df_name <- str_sub(deparse(substitute(df)), start = 3)

  d.plot <- df %>% 
    ungroup() %>%
    ##use {{}} to deparse variable name for use in dplyr functions
    filter({{col_name}} < stat_cutoff) %>%
    unite("target_compound_label", c("target", "compound_name"), sep = "\n", remove = FALSE) %>%
    mutate("target_compound_label" = reorder(target_compound_label, {{col_name}}),
           "target" = reorder(target, {{col_name}})) #%>%
  
    if(plot_n == "all") {
    ## do nothing to the df
    ## save output file name
    file_name <- paste0(date, "_", df_name, "_", stat_name, "_L", stat_cutoff, "_all_lineage.pdf")
    
  } else {
    plot_pairs <- d.plot %>%
      arrange({{col_name}}) %>%
      distinct(target_compound_label) %>%
      head(n = plot_n) %>%
      pull(target_compound_label)
    
    d.plot <- d.plot %>%
      filter(target_compound_label %in% plot_pairs)
    
    ## save output file name
    file_name <- paste0(date, "_", df_name, "_", stat_name, "_L", stat_cutoff, "_top", plot_n, "_lineage.pdf")
  }

  
  length <- d.plot %>%
    ungroup() %>%
    distinct(target_compound_label) %>%
    nrow()
    
  # aspect_ratio <- 1
  n_rows <- 1
  n_cols <- 1
  n_pages <- ceiling(length/(n_rows*n_cols))

  pdf(file.path(out_dir, file_name), width = 6, height = 4)
  for(i in 1:n_pages){

    q <- ggplot(d.plot, aes_string(x = x_var, y = y_var)) +
      geom_jitter(width = 0.3, size = 1) +
      geom_boxplot(alpha = 0, color = "gray55", outlier.shape = NA, coef = 0) +
      labs(x = x_title, y = y_title) +
      theme_bw() +
      theme(axis.text = element_text(color = "black"),
            axis.ticks = element_line(color = "black"),
            axis.text.x = element_text(angle = 45, hjust=1),
            legend.position = "none") +
      facet_wrap_paginate(~target_compound_label, scale = "free",
                          nrow = n_rows, ncol = n_cols, page = i)
    print(ggarrange(q, ncol = 1, align = "hv"))
  }
  dev.off()
  
}


```


```{r}

# make_lm_lineage_plot(d.drug_sens_all_genes_expr_lm_pos_slope, 
#                      "lineage", "Lineage",
#                      "log_tpm", "log2(TPM)",
#                      fdr, 0.1, 50)
# 
# make_lm_lineage_plot(d.drug_sens_all_genes_expr_lm_neg_slope,  
#                      "lineage", "Lineage",
#                      "log_tpm", "log2(TPM)",
#                      fdr, 0.1, 50)

```





#### Z-scaled outliers

```{r}

## robustly scale expression data
d.drug_sens_all_genes_expr_outlier <- d.drug_sens_all_genes_expr %>%
  filter(!is.na(log_tpm)) %>%
  filter(!is.na(dependency)) %>%
  group_by(target, compound_id) %>% # is this correct?
  mutate(scaled_log_tpm = RobScale(log_tpm, center = TRUE, scale = TRUE)) %>%
  mutate(expr_outlier_flag = case_when(
    scaled_log_tpm < -2 ~ "low",
    scaled_log_tpm > 2 ~ "high", 
    TRUE ~ "neither")) %>%
  mutate(expr_low_flag = case_when(
    scaled_log_tpm < -2 ~ TRUE,
    TRUE ~ FALSE)) %>% ## if CN ≥ -2, classify as "not low"
  ungroup()

d.drug_sens_all_genes_expr_outlier %>%
  group_by(expr_outlier_flag) %>%
  summarize(n = n())

d.drug_sens_all_genes_expr_outlier %>%
  group_by(expr_low_flag) %>%
  summarize(n = n())

d.drug_sens_all_genes_expr_outlier %>%
  group_by(target, compound_id) %>%
  distinct(expr_low_flag) %>%
  summarize(n = n()) %>%
  filter(n < 2)

```


```{r}

## calculate the WRST p-value for low-CN vs. not low-CN cell lines
d.drug_sens_all_genes_expr_outlier_wrst_summary <- d.drug_sens_all_genes_expr_outlier %>%
  group_by(target, compound_id) %>%
  filter(any(expr_low_flag == TRUE)) %>% ## keep only groups where >1 cell line has low CN
  summarize("p_val" = wilcox.test(dependency ~ expr_low_flag)$p.value)
# d.drug_sens_all_genes_expr_outlier_wrst_summary

d.stats <- BH_adjust(d.drug_sens_all_genes_expr_outlier_wrst_summary)

d.drug_sens_all_genes_expr_outlier_wrst_summary <- d.drug_sens_all_genes_expr_outlier_wrst_summary %>%
  dplyr::select(-p_val) %>%
  left_join(d.stats, by = c("target", "compound_id"))



## add p-values and FDR to main DF
d.drug_sens_all_genes_expr_outlier_wrst <- left_join(d.drug_sens_all_genes_expr_outlier,
                                                   d.drug_sens_all_genes_expr_outlier_wrst_summary,
                                                   by = c("target", "compound_id")) %>%
  ungroup()
# d.drug_sens_all_genes_expr_outlier_wrst

summarize_stats(d.drug_sens_all_genes_expr_outlier_wrst)

```


##### Plot
# TO DO
Plot just top ~5 or so genes
```{r}

d.plot <- d.drug_sens_all_genes_expr_outlier_wrst %>% 
  ungroup() %>%
  filter(fdr < 0.1) %>%
  unite("target_compound_label", c("target", "compound_name"), sep = "\n", remove = FALSE) %>%
  mutate("target_compound_label" = reorder(target_compound_label, fdr),
         "target" = reorder(target, fdr)) %>%
  mutate(expr_low_flag = factor(expr_low_flag, levels = c(TRUE, FALSE)))

length <- d.plot %>%
  ungroup() %>%
  distinct(target_compound_label) %>%
  nrow()
  
aspect_ratio <- 0.8
n_rows <- 1
n_cols <- 1
n_pages <- ceiling(length/(n_rows*n_cols))

# pdf(file.path(out_dir, paste(date, "drug_sens_all_genes_expr_outlier_wrst_fdr_L10.pdf", sep = "_")),
#     width = 6, height = 3)
# for(i in 1:n_pages){
#   p <-  ggplot(d.plot, aes(x = expr_low_flag, y = dependency)) +
#     geom_hline(yintercept = 0) +
#     geom_jitter(aes(color = expr_low_flag), width = 0.3) +
#     geom_boxplot(alpha = 0) +
#     geom_signif(comparisons = list(c(1, 2)),
#                 map_signif_level = FALSE,
#                 test = "wilcox.test",
#                 # test.args = list(alternative = "less", p.adjust.method = "BH"),
#                 # annotations = fdr_labels, # fix this to get FDR plotted direclty
#                 textsize = 3.2) +
#     # add_pvalue(d.fdr_labels) +
#     ## adds 10% of space on top & bottom so p-value label will fit
#     scale_y_continuous(expand = expansion(mult = 0.15)) +
#     scale_color_manual(values = c("blue", "gray50")) +
#     labs(x = "Expr_outlier_low",
#          y = "LFC_treated_vs_DMSO") +
#     theme_bw() +
#     theme(aspect.ratio = aspect_ratio,
#           axis.text = element_text(color = "black"),
#           axis.ticks = element_line(color = "black"),
#           legend.position = "none") +
#     facet_wrap_paginate(~target_compound_label, scales = "free_y",
#                         nrow = n_rows, ncol = n_cols, page = i)
# 
#   q <- ggplot(d.plot, aes(x = target, y = log_tpm)) +
#     geom_jitter(aes(color = expr_low_flag), width = 0.3) +
#     geom_boxplot(alpha = 0) +
#     ## adds 10% of space on top & bottom so p-value label will fit
#     # scale_y_continuous(expand = expansion(mult = 0.15)) +
#     scale_color_manual(values = c("blue", "gray50")) +
#     labs(x = "target", y = "log2(TPM)") +
#     theme_bw() +
#     theme(aspect.ratio = aspect_ratio,
#           axis.text = element_text(color = "black"),
#           axis.ticks = element_line(color = "black")) +
#     facet_wrap_paginate(~target_compound_label, scale = "free",
#                         nrow = n_rows, ncol = n_cols, page = i)
#   print(ggarrange(p, q, ncol = 2, align = "hv", common.legend = TRUE, legend = "right"))
# }
# dev.off()

```




##### Split into positive and negative outliers

```{r}

d.drug_sens_all_genes_expr_outlier <- d.drug_sens_all_genes_expr_outlier %>%
  mutate(expr_outlier_flag = factor(expr_outlier_flag, levels = c("low", "neither", "high"))) 

```


```{r}


d.drug_sens_all_genes_expr_outlier %>%
  group_by(target, compound_id, expr_outlier_flag) %>%
  summarize(n = n()) %>%
  group_by(expr_outlier_flag) %>%
  summarize(mean = mean(n),
            median = median(n))

## how many compounds/targets have 0 high or low outlier cell lines?
d.drug_sens_all_genes_expr_outlier %>%
  group_by(target, compound_id, expr_outlier_flag) %>%
  summarize(n = n()) 

d.drug_sens_all_genes_expr_outlier %>%
  group_by(target, compound_id, expr_outlier_flag) %>%
  summarize(n = n()) %>%
  ungroup() %>%
  group_by(target, compound_id) %>%
  summarize(n = n())
  

```

```{r}

## low vs. neither (with high included but ignoring it) => 
## calculate diff in median LFC between relevant groups

d.drug_sens_all_genes_expr_outlier_low <- d.drug_sens_all_genes_expr_outlier %>%
  group_by(target, compound_id) %>%
  ## get rid of drug/targets that don't have any "low" cell lines
  filter(any(expr_outlier_flag == "low")) %>%
  ungroup() 

d.drug_sens_all_genes_expr_outlier_low_grp_med_diff <- d.drug_sens_all_genes_expr_outlier_low %>%
  group_by(target, compound_id, expr_outlier_flag) %>%
  ## calculate median dependency score for each outlier group
  summarize(grp_med_dependency = median(dependency)) %>%
  pivot_wider(names_from = expr_outlier_flag, values_from = grp_med_dependency) %>%
  ## calculate difference between low and neither groups
  mutate(grp_med_diff = low - neither) %>%
  dplyr::select(target, compound_id, grp_med_diff) 

d.drug_sens_all_genes_expr_outlier_low <- d.drug_sens_all_genes_expr_outlier_low %>%
  left_join(d.drug_sens_all_genes_expr_outlier_low_grp_med_diff, by = c("target", "compound_id"))

## high vs. neither (with low included but ignoring it) => 
## calculate diff in median LFC between relevant groups =>
d.drug_sens_all_genes_expr_outlier_high <- d.drug_sens_all_genes_expr_outlier %>%
  group_by(target, compound_id) %>%
  ## get rid of drug/targets that don't have any "high" cell lines
  filter(any(expr_outlier_flag == "high")) %>%
  ungroup() 

d.drug_sens_all_genes_expr_outlier_high_grp_med_diff <- d.drug_sens_all_genes_expr_outlier_high %>%
  group_by(target, compound_id, expr_outlier_flag) %>%
  ## calculate median dependency score for each outlier group
  summarize(grp_med_dependency = median(dependency)) %>%
  pivot_wider(names_from = expr_outlier_flag, values_from = grp_med_dependency) %>%
  ## calculate difference between low and neither groups
  mutate(grp_med_diff = high - neither) %>%
  dplyr::select(target, compound_id, grp_med_diff) 

d.drug_sens_all_genes_expr_outlier_high <- d.drug_sens_all_genes_expr_outlier_high %>%
  left_join(d.drug_sens_all_genes_expr_outlier_high_grp_med_diff, by = c("target", "compound_id"))

```



```{r}

## low outliers and drug sensitivity
## calculate p-value (adjusted for multiple within-group tests?)
d.drug_sens_all_genes_expr_outlier_low_wrst_pvals <- d.drug_sens_all_genes_expr_outlier_low %>%
  ## remove cell lines that are high outliers (I think this makes sense... but I'm not sure)
  filter(expr_outlier_flag != "high") %>%
  group_by(target, compound_id) %>%
  summarize(p_val = wilcox.test(dependency ~ expr_outlier_flag)$p.value)
# d.drug_sens_all_genes_expr_outlier_low_wrst_sens_pvals

## adjust for multiple testing overall
d.stats <- BH_adjust(d.drug_sens_all_genes_expr_outlier_low_wrst_pvals)
# d.stats

d.drug_sens_all_genes_expr_outlier_low_wrst <- d.drug_sens_all_genes_expr_outlier_low %>%
  left_join(d.stats, by = c("target", "compound_id")) 

d.drug_sens_all_genes_expr_outlier_low_wrst_sens <- d.drug_sens_all_genes_expr_outlier_low_wrst %>%
  filter(grp_med_diff < 0)

d.drug_sens_all_genes_expr_outlier_low_wrst_resist <- d.drug_sens_all_genes_expr_outlier_low_wrst %>%
  filter(grp_med_diff > 0)

# d.drug_sens_all_genes_expr_outlier_low_wrst_resist

## high outliers and drug sensitivity
d.drug_sens_all_genes_expr_outlier_high_wrst_sens_pvals <- d.drug_sens_all_genes_expr_outlier_high %>%
  ## remove cell lines that are high outliers (I think this makes sense... but I'm not sure)
  filter(expr_outlier_flag != "low") %>%
  group_by(target, compound_id) %>%
  summarize(p_val = wilcox.test(dependency ~ expr_outlier_flag)$p.value)
# d.drug_sens_all_genes_expr_outlier_high_wrst_sens_pvals

## adjust for multiple testing overall
d.stats <- BH_adjust(d.drug_sens_all_genes_expr_outlier_high_wrst_sens_pvals)
# d.stats

d.drug_sens_all_genes_expr_outlier_high_wrst <- d.drug_sens_all_genes_expr_outlier_high %>%
  left_join(d.stats, by = c("target", "compound_id"))
# d.drug_sens_all_genes_expr_outlier_high_wrst_sens

## high outliers and drug sensitivity
d.drug_sens_all_genes_expr_outlier_high_wrst_sens <- d.drug_sens_all_genes_expr_outlier_high_wrst %>%
  ## remove cell lines that are high outliers (I think this makes sense... but I'm not sure)
  filter(grp_med_diff < 0)

d.drug_sens_all_genes_expr_outlier_high_wrst_resist <- d.drug_sens_all_genes_expr_outlier_high_wrst %>%
  ## remove cell lines that are high outliers (I think this makes sense... but I'm not sure)
  filter(grp_med_diff > 0)

```

```{r}

## low outliers
summarize_stats(d.drug_sens_all_genes_expr_outlier_low_wrst_sens)
summarize_stats(d.drug_sens_all_genes_expr_outlier_low_wrst_resist)

## high outliers
summarize_stats(d.drug_sens_all_genes_expr_outlier_high_wrst_sens)
summarize_stats(d.drug_sens_all_genes_expr_outlier_high_wrst_resist)

```

###### Plot outliers


## REMOVE BELOW CHUNK?
For deparsing variables to use in dplyr: https://cran.r-project.org/web/packages/dplyr/vignettes/programming.html
```{r}

# make_outlier_plot <- function(df, direction, col_name, stat_cutoff, plot_n){
#   
#   ## save col_name as a string for later use
#   stat_name <- deparse(substitute(col_name))
#   
#   ## remove the first 2 characters of dataframe name
#   df_name <- str_sub(deparse(substitute(df)), start = 3)
# 
#   d.plot <- df %>% 
#     ungroup() %>%
#     ##use {{}} to deparse variable name for use in dplyr functions
#     filter({{col_name}} < stat_cutoff) %>%
#     unite("target_compound_label", c("target", "compound_name"), sep = "\n", remove = FALSE) %>%
#     mutate("target_compound_label" = reorder(target_compound_label, {{col_name}}),
#            "target" = reorder(target, {{col_name}})) #%>%
#   
#   if(plot_n == "all") {
#     ## do nothing to the df
#     ## save output file name
#     file_name <- paste0(date, "_", df_name, "_", stat_name, "_L", stat_cutoff, "all.pdf")
#   
#   } else {
#     plot_pairs <- d.plot %>%
#       arrange({{col_name}}) %>%
#       distinct(target_compound_label) %>%
#       head(n = plot_n) %>%
#       pull(target_compound_label)
#         
#     d.plot <- d.plot %>%
#       filter(target_compound_label %in% plot_pairs)
#     
#     ## save output file name
#     file_name <- paste0(date, "_", df_name, "_", stat_name, "_L", stat_cutoff, "_top", plot_n, ".pdf")
#   }
# 
#   ## make df to label adjusted FDR/p-val for each plot
#   d.label <- d.plot %>%
#     group_by(target, compound_id) %>%
#     mutate(y.position = (max(dependency) + 0.1*(max(dependency) - min(dependency)))) %>%
#     dplyr::ungroup() %>%
#     dplyr::select(target_compound_label, p_val, fdr, y.position) %>%
#     distinct(target_compound_label, .keep_all = TRUE) %>%
#     mutate(group1 = direction,
#            group2 = "neither",
#            p_val = scientific(p_val, 3),
#            fdr = scientific(fdr, 3)) 
#   d.label
#   
#   length <- d.plot %>%
#     ungroup() %>%
#     distinct(target_compound_label) %>%
#     nrow()
#     
#   aspect_ratio <- 1
#   n_rows <- 1
#   n_cols <- 1
#   n_pages <- ceiling(length/(n_rows*n_cols))
#   
#   ## make stat label for labeling significance brackets
#   if(stat_name == "p_val"){
#     tmp_label <- str_sub(stat_name, end = 1)
#     stat_label <- paste0(toupper(tmp_label), " = ", "{", stat_name, "}")
#   } else if (stat_name == "fdr"){
#     stat_label <- paste0(toupper(stat_name), " = ", "{", stat_name, "}")
#   } else{
#     cat("\nPlease enter a valid statistic (p-value or FDR)")
#   }
#   
#   # stat_label <- paste0(toupper(stat_name), " = ", "{", stat_name, "}")
# 
#   pdf(file.path(out_dir, file_name), width = 6, height = 3)
#   for(i in 1:n_pages){
#     p <-  ggplot(d.plot, aes(x = expr_outlier_flag, y = dependency)) +
#       geom_hline(yintercept = 0) +
#       geom_jitter(aes(color = expr_outlier_flag), width = 0.3) +
#       geom_boxplot(alpha = 0) +
#       stat_pvalue_manual(d.label, label = stat_label, tip.length = 0.03, size = 3, vjust = -0.5) +
#       ## adds extra space on top & bottom so p-value label will fit
#       scale_y_continuous(expand = expansion(mult = 0.15)) +
#       scale_color_manual(values = c("low" = "blue", "neither" = "gray50", "high" = "red")) +
#       labs(x = "Expression_group",
#            y = "LFC_treated_vs_DMSO") +
#       theme_bw() +
#       theme(aspect.ratio = aspect_ratio,
#             axis.text = element_text(color = "black"),
#             axis.ticks = element_line(color = "black"),
#             legend.position = "none") +
#       facet_wrap_paginate(~target_compound_label, scales = "free_y",
#                           nrow = n_rows, ncol = n_cols, page = i)
# 
#     q <- ggplot(d.plot, aes(x = target, y = log_tpm)) +
#       geom_jitter(aes(color = expr_outlier_flag), width = 0.3) +
#       geom_boxplot(alpha = 0) +
#       scale_color_manual(values = c("low" = "blue", "neither" = "gray50", "high" = "red")) +
#       labs(x = "Target", y = "log2(TPM)") +
#       theme_bw() +
#       theme(aspect.ratio = aspect_ratio,
#             axis.text = element_text(color = "black"),
#             axis.ticks = element_line(color = "black")) +
#       facet_wrap_paginate(~target_compound_label, scale = "free",
#                           nrow = n_rows, ncol = n_cols, page = i)
#     print(ggarrange(p, q, ncol = 2, align = "hv", common.legend = TRUE, legend = "right"))
#   }
#   dev.off()
#   
# }


```

```{r}

make_outlier_plot <- function(df, x_var, x_title, y_var, y_title, direction, col_name, stat_cutoff, plot_n){
  
  ## save col_name as a string for later use
  stat_name <- deparse(substitute(col_name))
  
  ## remove the first 2 characters of dataframe name
  df_name <- str_sub(deparse(substitute(df)), start = 3)

  d.plot <- df %>% 
    ungroup() %>%
    ## use {{}} to deparse variable name for use in dplyr functions
    filter({{col_name}} < stat_cutoff) %>%
    unite("target_compound_label", c("target", "compound_name"), sep = "\n", remove = FALSE) %>%
    mutate("target_compound_label" = reorder(target_compound_label, {{col_name}}),
           "target" = reorder(target, {{col_name}})) #%>%
  
  if(plot_n == "all") {
    ## do nothing to the df
    ## save output file name
    file_name <- paste0(date, "_", df_name, "_", stat_name, "_L", stat_cutoff, "all.pdf")
  
  } else {
    plot_pairs <- d.plot %>%
      arrange({{col_name}}) %>%
      distinct(target_compound_label) %>%
      head(n = plot_n) %>%
      pull(target_compound_label)
        
    d.plot <- d.plot %>%
      filter(target_compound_label %in% plot_pairs)
    
    ## save output file name
    file_name <- paste0(date, "_", df_name, "_", stat_name, "_L", stat_cutoff, "_top", plot_n, ".pdf")
  }

  ## make df to label adjusted FDR/p-val for each plot
  d.label <- d.plot %>%
    group_by(target, compound_id) %>%
    mutate(y.position = (max(dependency) + 0.1*(max(dependency) - min(dependency)))) %>%
    dplyr::ungroup() %>%
    dplyr::select(target_compound_label, p_val, fdr, y.position) %>%
    distinct(target_compound_label, .keep_all = TRUE) %>%
    mutate(group1 = direction,
           group2 = "neither",
           p_val = scientific(p_val, 3),
           fdr = scientific(fdr, 3)) 
  d.label
  
  length <- d.plot %>%
    ungroup() %>%
    distinct(target_compound_label) %>%
    nrow()
    
  aspect_ratio <- 1
  n_rows <- 1
  n_cols <- 1
  n_pages <- ceiling(length/(n_rows*n_cols))
  
  ## make stat label for labeling significance brackets
  if(stat_name == "p_val"){
    tmp_label <- str_sub(stat_name, end = 1)
    stat_label <- paste0(toupper(tmp_label), " = ", "{", stat_name, "}")
  } else if (stat_name == "fdr"){
    stat_label <- paste0(toupper(stat_name), " = ", "{", stat_name, "}")
  } else{
    cat("\nPlease enter a valid statistic (p-value or FDR)")
  }
  
  # stat_label <- paste0(toupper(stat_name), " = ", "{", stat_name, "}")

  pdf(file.path(out_dir, file_name), width = 6, height = 3)
  for(i in 1:n_pages){
    p <-  ggplot(d.plot, aes_string(x = x_var[1], y = y_var[1])) +
      geom_hline(yintercept = 0) +
      geom_jitter(aes(color = expr_outlier_flag), width = 0.3) +
      geom_boxplot(alpha = 0) +
      stat_pvalue_manual(d.label, label = stat_label, tip.length = 0.03, size = 3, vjust = -0.5) +
      ## adds extra space on top & bottom so p-value label will fit
      scale_y_continuous(expand = expansion(mult = 0.15)) +
      # scale_color_manual(values = c("low" = "blue", "neither" = "gray50", "high" = "red")) +
      scale_color_manual(values = c("low" = "#0080C6", "neither" = "#7F7F7F", "high" = "#B84250")) +
      labs(x = x_title[1],
           y = y_title[1]) +
      theme_bw() +
      theme(aspect.ratio = aspect_ratio,
            axis.text = element_text(color = "black"),
            axis.ticks = element_line(color = "black"),
            legend.position = "none") +
      facet_wrap_paginate(~target_compound_label, scales = "free_y",
                          nrow = n_rows, ncol = n_cols, page = i)

    q <- ggplot(d.plot, aes_string(x = x_var[2], y = y_var[2])) +
      geom_jitter(aes(color = expr_outlier_flag), width = 0.3) +
      geom_boxplot(alpha = 0) +
      # scale_color_manual(values = c("low" = "blue", "neither" = "gray50", "high" = "red")) +
      scale_color_manual(values = c("low" = "#0080C6", "neither" = "#7F7F7F", "high" = "#B84250")) +
      labs(x = x_title[2], y = y_title[2]) +
      theme_bw() +
      theme(aspect.ratio = aspect_ratio,
            axis.text = element_text(color = "black"),
            axis.ticks = element_line(color = "black")) +
      facet_wrap_paginate(~target_compound_label, scale = "free",
                          nrow = n_rows, ncol = n_cols, page = i)
    print(ggarrange(p, q, ncol = 2, align = "hv", common.legend = TRUE, legend = "right"))
  }
  dev.off()
  
}


```


```{r}

# make_outlier_plot(d.drug_sens_all_genes_expr_outlier_low_wrst_sens,
#                   ## x-axis info
#                   c("expr_outlier_flag", "target"), c("Expression_group", "Target"),
#                   ## y-axis info
#                   c("dependency", "log_tpm"), c("LFC_treated_vs_DMSO", "log2(TPM)"),
#                   ## other information
#                   "low", p_val, 0.05, 50)
# 
# make_outlier_plot(d.drug_sens_all_genes_expr_outlier_low_wrst_resist,
#                   ## x-axis info
#                   c("expr_outlier_flag", "target"), c("Expression_group", "Target"),
#                   ## y-axis info
#                   c("dependency", "log_tpm"), c("LFC_treated_vs_DMSO", "log2(TPM)"),
#                   ## other information
#                   "low", fdr, 0.1, 1)
# 
# make_outlier_plot(d.drug_sens_all_genes_expr_outlier_high_wrst_sens,
#                   ## x-axis info
#                   c("expr_outlier_flag", "target"), c("Expression_group", "Target"),
#                   ## y-axis info
#                   c("dependency", "log_tpm"), c("LFC_treated_vs_DMSO", "log2(TPM)"),
#                   ## other information
#                   "high", p_val, 0.05, 50)
# 
# make_outlier_plot(d.drug_sens_all_genes_expr_outlier_high_wrst_resist,
#                   ## x-axis info
#                   c("expr_outlier_flag", "target"), c("Expression_group", "Target"),
#                   ## y-axis info
#                   c("dependency", "log_tpm"), c("LFC_treated_vs_DMSO", "log2(TPM)"),
#                   ## other information
#                   "high", p_val, 0.05, 50)


```

```{r}

# make_outlier_plot(d.drug_sens_all_genes_expr_outlier_low_wrst_sens, "low", p_val, 0.05, 50)
# 
# make_outlier_plot(d.drug_sens_all_genes_expr_outlier_low_wrst_resist, "low", p_val, 0.05, 50)
# 
# make_outlier_plot(d.drug_sens_all_genes_expr_outlier_high_wrst_sens, "high", p_val, 0.05, 50)
# 
# make_outlier_plot(d.drug_sens_all_genes_expr_outlier_high_wrst_resist, "high", p_val, 0.05, 50)

```

```{r}

write_output_df <- function(df, col_name, stat_cutoff){
  
  df_out <- df %>%
    filter({{col_name}} < stat_cutoff) %>%
    arrange({{col_name}}) %>%
    distinct(target, compound_id, .keep_all = TRUE) %>%
    dplyr::select(target, compound_name, grp_med_diff, p_val, fdr)
  
  df_name <- deparse(substitute(df))
  stat_name <- deparse(substitute(col_name))
  
  # df_name <- paste0(deparse(substitute(df)), "_",
  #                   deparse(substitute(col_name)), "_",
  #                   "L", stat_cutoff, ".txt")
  # print(df_name)
  # print(stat_name)
  
  # print(nrow(df_out))
  
  out_path <- file.path(out_dir, paste0(df_name, "_", stat_name, "_", "L", stat_cutoff, ".txt"))
  # print(out_path)
  
  write_tsv(df_out, out_path)
  
  cat("\ndf written to", out_path)
}


```


```{r}

# write_output_df(d.drug_sens_all_genes_expr_outlier_low_wrst_sens, p_val, 0.05)
# 
# write_output_df(d.drug_sens_all_genes_expr_outlier_low_wrst_resist, p_val, 0.05)
# 
# write_output_df(d.drug_sens_all_genes_expr_outlier_high_wrst_sens, p_val, 0.05)
# 
# write_output_df(d.drug_sens_all_genes_expr_outlier_high_wrst_resist, p_val, 0.05)

```


##### Plot lineage

```{r}

make_lineage_plot <- function(df, x_var, x_title, y_var, y_title, col_name, stat_cutoff, plot_n){
  
  ## save col_name as a string for later use
  stat_name <- deparse(substitute(col_name))
  
  ## remove the first 2 characters of dataframe name
  df_name <- str_sub(deparse(substitute(df)), start = 3)

  d.plot <- df %>% 
    ungroup() %>%
    ##use {{}} to deparse variable name for use in dplyr functions
    filter({{col_name}} < stat_cutoff) %>%
    unite("target_compound_label", c("target", "compound_name"), sep = "\n", remove = FALSE) %>%
    mutate("target_compound_label" = reorder(target_compound_label, {{col_name}}),
           "target" = reorder(target, {{col_name}})) #%>%
  
    if(plot_n == "all") {
      ## don't do anything to the df
      ## save output file name
      file_name <- paste0(date, "_", df_name, "_", stat_name, "_L", stat_cutoff, "_all_lineage.pdf")
    
  } else {
      plot_pairs <- d.plot %>%
        arrange({{col_name}}) %>%
        distinct(target_compound_label) %>%
        head(n = plot_n) %>%
        pull(target_compound_label)
      
      d.plot <- d.plot %>%
        filter(target_compound_label %in% plot_pairs)
      
      ## save output file name
      file_name <- paste0(date, "_", df_name, "_", stat_name, "_L", stat_cutoff, "_top", plot_n, "_lineage.pdf")
  }
  
  length <- d.plot %>%
    ungroup() %>%
    distinct(target_compound_label) %>%
    nrow()
    
  # aspect_ratio <- 1
  n_rows <- 1
  n_cols <- 1
  n_pages <- ceiling(length/(n_rows*n_cols))
  
  ## make stat label for labeling significance brackets
  stat_label <- paste0(toupper(stat_name), " = ", "{", stat_name, "}")

  ## save output file name
  file_name <- paste(date, df_name, stat_name, "L", stat_cutoff, "cell_lineage.pdf", sep = "_")

  pdf(file.path(out_dir, file_name), width = 6, height = 4)
  for(i in 1:n_pages){

    q <- ggplot(d.plot, aes_string(x = x_var, y = y_var)) +
      geom_jitter(aes(color = expr_outlier_flag), width = 0.3, size = 1) +
      geom_boxplot(alpha = 0) +
      scale_color_manual(values = c("low" = "blue", "neither" = "gray50", "high" = "red")) +
      labs(x = x_title, y = y_title) +
      theme_bw() +
      theme(#aspect.ratio = aspect_ratio,
            axis.text = element_text(color = "black"),
            axis.ticks = element_line(color = "black"),
            axis.text.x = element_text(angle = 45, hjust=1)) +
      facet_wrap_paginate(~target_compound_label, scale = "free",
                          nrow = n_rows, ncol = n_cols, page = i)
    print(ggarrange(q, ncol = 1, align = "hv", legend = "right"))
  }
  dev.off()
  
}


```

```{r}

# make_lineage_plot(d.drug_sens_all_genes_expr_outlier_low_wrst_sens, 
#                   "lineage", "Cell_lineage", 
#                   "log_tpm", "log2(TPM)",
#                   p_val, 0.05, 50)
# # 
# make_lineage_plot(d.drug_sens_all_genes_expr_outlier_low_wrst_resist, 
#                   "lineage", "Cell_lineage", 
#                   "log_tpm", "log2(TPM)",
#                   p_val, 0.05, 50)
# # 
# make_lineage_plot(d.drug_sens_all_genes_expr_outlier_high_wrst_sens, 
#                   "lineage", "Cell_lineage", 
#                   "log_tpm", "log2(TPM)",
#                   p_val, 0.05, 50)
# # 
# make_lineage_plot(d.drug_sens_all_genes_expr_outlier_high_wrst_resist, 
#                   "lineage", "Cell_lineage", 
#                   "log_tpm", "log2(TPM)",
#                   p_val, 0.05, 50)

```



## Paralogs

### PRISM Phase 1 drug sensitivity

```{r}

## add target col so single-gene functions will work
d.DeKegel_top5_pairs_drug_sens_expr <- d.DeKegel_top5_pairs_drug_sens_expr %>%
  mutate(target = sorted_gene_pair)
d.DeKegel_top5_pairs_drug_sens_expr

```

#### Continuous/linear regression

```{r}

## running lm for each sorted_gene_pair/compound_id group

## combined expr
## use broom::glance to get rsquared and pval
# d.DeKegel_top5_pairs_drug_sens_combined_expr_lm_summary <- d.DeKegel_top5_pairs_drug_sens_expr %>%
#   filter(!is.na(dependency)) %>%
#   filter(!is.na(combined_log_tpm)) %>%
#   group_by(target, compound_id) %>%
#   group_modify(~ broom::glance(lm(dependency ~ combined_log_tpm, data = .x))) %>%
#   dplyr::ungroup() %>%
#   rename(r_squared = r.squared, p_val = p.value)
# # d.DeKegel_top5_pairs_drug_sens_combined_expr_lm_summary
# 
# ## FDR-adjust p-vals for multiple testing and add back to summary DF
# d.stats <- BH_adjust(d.DeKegel_top5_pairs_drug_sens_combined_expr_lm_summary)
# # d.stats
# 
# d.DeKegel_top5_pairs_drug_sens_combined_expr_lm_summary <- d.stats %>%
#   dplyr::select(-p_val) %>%
#   right_join(d.DeKegel_top5_pairs_drug_sens_combined_expr_lm_summary, by = c("target", "compound_id")) %>%
#   dplyr::select(target, compound_id, r_squared, p_val, fdr)
# # d.DeKegel_top5_pairs_drug_sens_combined_expr_lm_summary
# 
# ## use broom::tidy to get slope
# ## note: group_modify expects a DF output
# d.DeKegel_top5_pairs_drug_sens_combined_expr_lm_slope <- d.DeKegel_top5_pairs_drug_sens_expr %>%
#   filter(!is.na(dependency)) %>%
#   filter(!is.na(combined_log_tpm)) %>%
#   group_by(target, compound_id) %>%
#   group_modify(~ broom::tidy(lm(dependency ~ combined_log_tpm, data = .x))) %>%
#   dplyr::ungroup() %>%
#   ## extract slope of lm fit line from output
#   filter(term == "combined_log_tpm") %>%
#   rename(slope = estimate) %>%
#   dplyr::select(target, compound_id, slope) 
# # d.DeKegel_top5_pairs_drug_sens_combined_expr_lm_slope
# 
# ## add rsquared, p-val, fdr, and slope info to main DF
# d.DeKegel_top5_pairs_drug_sens_combined_expr_lm <- d.DeKegel_top5_pairs_drug_sens_expr %>%
#   left_join(d.DeKegel_top5_pairs_drug_sens_combined_expr_lm_summary, by = c("target", "compound_id")) %>%
#   left_join(d.DeKegel_top5_pairs_drug_sens_combined_expr_lm_slope, by = c("target", "compound_id")) %>%
#   filter(!is.na(dependency)) %>%
#   filter(!is.na(combined_log_tpm))
# 
# summarize_stats(d.DeKegel_top5_pairs_drug_sens_combined_expr_lm)

```

```{r}

## A1 expr

## use broom::glance to get rsquared and pval
d.DeKegel_top5_pairs_drug_sens_A1_expr_lm_summary <- d.DeKegel_top5_pairs_drug_sens_expr %>%
  filter(!is.na(dependency)) %>%
  filter(!is.na(A1_log_tpm)) %>%
  group_by(target, compound_id) %>%
  group_modify(~ broom::glance(lm(dependency ~ A1_log_tpm, data = .x))) %>%
  dplyr::ungroup() %>%
  rename(r_squared = r.squared, p_val = p.value)
# d.DeKegel_top5_pairs_drug_sens_A1_expr_lm_summary

## FDR-adjust p-vals for multiple testing and add back to summary DF
d.stats <- BH_adjust(d.DeKegel_top5_pairs_drug_sens_A1_expr_lm_summary)
# d.stats

d.DeKegel_top5_pairs_drug_sens_A1_expr_lm_summary <- d.stats %>%
  dplyr::select(-p_val) %>%
  right_join(d.DeKegel_top5_pairs_drug_sens_A1_expr_lm_summary, by = c("target", "compound_id")) %>%
  dplyr::select(target, compound_id, r_squared, p_val, fdr)
# d.DeKegel_top5_pairs_drug_sens_A1_expr_lm_summary

## use broom::tidy to get slope
## note: group_modify expects a DF output
d.DeKegel_top5_pairs_drug_sens_A1_expr_lm_slope <- d.DeKegel_top5_pairs_drug_sens_expr %>%
  filter(!is.na(dependency)) %>%
  filter(!is.na(A1_log_tpm)) %>%
  group_by(target, compound_id) %>%
  group_modify(~ broom::tidy(lm(dependency ~ A1_log_tpm, data = .x))) %>%
  dplyr::ungroup() %>%
  ## extract slope of lm fit line from output
  filter(term == "A1_log_tpm") %>%
  rename(slope = estimate) %>%
  dplyr::select(target, compound_id, slope) 
# d.DeKegel_top5_pairs_drug_sens_A1_expr_lm_slope

## add rsquared, p-val, fdr, and slope info to main DF
d.DeKegel_top5_pairs_drug_sens_A1_expr_lm <- d.DeKegel_top5_pairs_drug_sens_expr %>%
  left_join(d.DeKegel_top5_pairs_drug_sens_A1_expr_lm_summary, by = c("target", "compound_id")) %>%
  left_join(d.DeKegel_top5_pairs_drug_sens_A1_expr_lm_slope, by = c("target", "compound_id")) %>%
  filter(!is.na(dependency)) %>%
  filter(!is.na(A1_log_tpm))

summarize_stats(d.DeKegel_top5_pairs_drug_sens_A1_expr_lm)

```

```{r}

## A2 expr

## use broom::glance to get rsquared and pval
d.DeKegel_top5_pairs_drug_sens_A2_expr_lm_summary <- d.DeKegel_top5_pairs_drug_sens_expr %>%
  filter(!is.na(dependency)) %>%
  filter(!is.na(A2_log_tpm)) %>%
  group_by(target, compound_id) %>%
  group_modify(~ broom::glance(lm(dependency ~ A2_log_tpm, data = .x))) %>%
  dplyr::ungroup() %>%
  rename(r_squared = r.squared, p_val = p.value)
# d.DeKegel_top5_pairs_drug_sens_A2_expr_lm_summary

## FDR-adjust p-vals for multiple testing and add back to summary DF
d.stats <- BH_adjust(d.DeKegel_top5_pairs_drug_sens_A2_expr_lm_summary)
# d.stats

d.DeKegel_top5_pairs_drug_sens_A2_expr_lm_summary <- d.stats %>%
  dplyr::select(-p_val) %>%
  right_join(d.DeKegel_top5_pairs_drug_sens_A2_expr_lm_summary, by = c("target", "compound_id")) %>%
  dplyr::select(target, compound_id, r_squared, p_val, fdr)
# d.DeKegel_top5_pairs_drug_sens_A2_expr_lm_summary

## use broom::tidy to get slope
## note: group_modify expects a DF output
d.DeKegel_top5_pairs_drug_sens_A2_expr_lm_slope <- d.DeKegel_top5_pairs_drug_sens_expr %>%
  filter(!is.na(dependency)) %>%
  filter(!is.na(A2_log_tpm)) %>%
  group_by(target, compound_id) %>%
  group_modify(~ broom::tidy(lm(dependency ~ A2_log_tpm, data = .x))) %>%
  dplyr::ungroup() %>%
  ## extract slope of lm fit line from output
  filter(term == "A2_log_tpm") %>%
  rename(slope = estimate) %>%
  dplyr::select(target, compound_id, slope) 
# d.DeKegel_top5_pairs_drug_sens_A2_expr_lm_slope

## add rsquared, p-val, fdr, and slope info to main DF
d.DeKegel_top5_pairs_drug_sens_A2_expr_lm <- d.DeKegel_top5_pairs_drug_sens_expr %>%
  left_join(d.DeKegel_top5_pairs_drug_sens_A2_expr_lm_summary, by = c("target", "compound_id")) %>%
  left_join(d.DeKegel_top5_pairs_drug_sens_A2_expr_lm_slope, by = c("target", "compound_id")) %>%
  filter(!is.na(dependency)) %>%
  filter(!is.na(A2_log_tpm))

summarize_stats(d.DeKegel_top5_pairs_drug_sens_A2_expr_lm)

```



```{r}

## number of pairs with negative vs. positive slope
# d.DeKegel_top5_pairs_drug_sens_combined_expr_lm %>%
#   distinct(target, compound_id, .keep_all = TRUE) %>%
#   mutate(slope_flag = ifelse(slope < 0, "negative", "positive")) %>%
#   group_by(slope_flag) %>%
#   summarize(n = n())

d.DeKegel_top5_pairs_drug_sens_A1_expr_lm %>%
  distinct(target, compound_id, .keep_all = TRUE) %>%
  mutate(slope_flag = ifelse(slope < 0, "negative", "positive")) %>%
  group_by(slope_flag) %>%
  summarize(n = n())

d.DeKegel_top5_pairs_drug_sens_A2_expr_lm %>%
  distinct(target, compound_id, .keep_all = TRUE) %>%
  mutate(slope_flag = ifelse(slope < 0, "negative", "positive")) %>%
  group_by(slope_flag) %>%
  summarize(n = n())

```


```{r}

# ## get low expression = low LFC pairs (positive slope)
# d.DeKegel_top5_pairs_drug_sens_combined_expr_lm_pos_slope <-d.DeKegel_top5_pairs_drug_sens_combined_expr_lm %>%
#   filter(slope > 0)
# summarize_stats(d.DeKegel_top5_pairs_drug_sens_combined_expr_lm_pos_slope)
# 
# ## get high expression = low LFC pairs (negative slope)
# d.DeKegel_top5_pairs_drug_sens_combined_expr_lm_neg_slope <- d.DeKegel_top5_pairs_drug_sens_combined_expr_lm %>%
#   filter(slope < 0)
# summarize_stats(d.DeKegel_top5_pairs_drug_sens_combined_expr_lm_neg_slope)

## get low expression = low LFC pairs (positive slope)
d.DeKegel_top5_pairs_drug_sens_A1_expr_lm_pos_slope <-d.DeKegel_top5_pairs_drug_sens_A1_expr_lm %>%
  filter(slope > 0)
summarize_stats(d.DeKegel_top5_pairs_drug_sens_A1_expr_lm_pos_slope)

## get high expression = low LFC pairs (negative slope)
d.DeKegel_top5_pairs_drug_sens_A1_expr_lm_neg_slope <- d.DeKegel_top5_pairs_drug_sens_A1_expr_lm %>%
  filter(slope < 0)
summarize_stats(d.DeKegel_top5_pairs_drug_sens_A1_expr_lm_neg_slope)

## get low expression = low LFC pairs (positive slope)
d.DeKegel_top5_pairs_drug_sens_A2_expr_lm_pos_slope <-d.DeKegel_top5_pairs_drug_sens_A2_expr_lm %>%
  filter(slope > 0)
summarize_stats(d.DeKegel_top5_pairs_drug_sens_A2_expr_lm_pos_slope)

## get high expression = low LFC pairs (negative slope)
d.DeKegel_top5_pairs_drug_sens_A2_expr_lm_neg_slope <- d.DeKegel_top5_pairs_drug_sens_A2_expr_lm %>%
  filter(slope < 0)
summarize_stats(d.DeKegel_top5_pairs_drug_sens_A2_expr_lm_neg_slope)

```

##### Plot

```{r}

make_paralog_lm_plot <- function(df, x_var, x_title, y_var, y_title, col_name, stat_cutoff, plot_n){
  ## save col_name as a string for later
  stat_name <- deparse(substitute(col_name))
  
  ## remove the first 2 characters of dataframe name
  df_name <- str_sub(deparse(substitute(df)), start = 3)
  
  ## filter DF for only pairs that I want to plot
  d.drug_plot <- df %>% 
    ungroup() %>%
    ##use {{}} to deparse variable name for use in dplyr functions
    filter({{col_name}} < stat_cutoff) %>%
    unite("target_compound_label", c("target", "compound_name"), sep = "\n", remove = FALSE) %>%
    mutate("target_compound_label" = reorder(target_compound_label, {{col_name}}),
           "target" = reorder(target, {{col_name}}))
  
  # print("hi")
  
  if(plot_n == "all") {
    ## do nothing to the df
    ## save output file name
    file_name <- paste0(date, "_", df_name, "_", stat_name, "_L", stat_cutoff, "all.pdf")
    
  } else {
    plot_pairs <- d.drug_plot %>%
      arrange({{col_name}}) %>%
      distinct(target_compound_label) %>%
      head(n = plot_n) %>%
      pull(target_compound_label)
    
    d.drug_plot <- d.drug_plot %>%
      filter(target_compound_label %in% plot_pairs)
    
    ## save output file name
    file_name <- paste0(date, "_", df_name, "_", stat_name, "_L", stat_cutoff, "_top", plot_n, ".pdf")
  }
  
  # print(d.drug_plot)
  
  # print("hello")

    ## filter DF for only pairs that I want to plot
  d.expr_plot <- d.drug_plot %>% 
    ungroup() %>%
    dplyr::select(target_compound_label, A1_log_tpm, A2_log_tpm) %>%
    pivot_longer(cols = c(A1_log_tpm, A2_log_tpm),
               names_to = "pos",
               values_to = "log_tpm") 
  # print(d.expr_plot)
  
  # print("good morning")
  
  ## make df to label rsquared and pvalue for each plot
  d.label <- d.drug_plot %>%
    group_by(target, compound_id) %>%
    mutate(group_label_x = min({{x_var}}),
           group_label_y = max({{y_var}})) %>%
    dplyr::ungroup() %>%
    dplyr::select(target_compound_label, r_squared, p_val, fdr, group_label_x, group_label_y) %>%
    distinct(target_compound_label, .keep_all = TRUE) %>%
    mutate(r_squared = round(r_squared, 3),
           p_val = scientific(p_val, 3),
           fdr = scientific(fdr, 3)) 
  # print(d.label)
  
  # print("good afternoon")
  
  ## set length of DF to plot
  length <- d.drug_plot %>%
    ungroup() %>%
    distinct(target_compound_label) %>%
    nrow()
  
  aspect_ratio <- 1
  n_rows <- 1
  n_cols <- 1
  n_pages <- ceiling(length/(n_rows*n_cols))

  pdf(file.path(out_dir, file_name), width = 6, height = 3) 
  # label = stat_label
  for(i in 1:n_pages){
    p <-  ggplot(d.drug_plot, aes(x = {{x_var}}, y = {{y_var}})) +
      geom_hline(yintercept = 0) +
      geom_point(size = 1) +
      geom_smooth(method = lm, se = FALSE, color = "gray60") +
      geom_richtext(data = d.label,
                    aes(x = group_label_x, y = group_label_y,
                    ## this package can do HTML-formatting for text
                    label = paste0("r<sup>2</sup>=", r_squared, "<br>", 
                                   toupper(stat_name), "=", {{col_name}})),
                    ## adjust label alignment and remove borders
                    hjust = 0, vjust = 1, label.color = NA, size = 3,
                    ## label background = transparent, text is not
                    fill = alpha(c("white"), 0.75),
                    label.margin = unit(c(0, 0, 0, 0), "lines"),
                    label.padding = unit(c(0, 0, 0, 0), "lines")) +
      labs(x = x_title, y = y_title) +
      theme_bw() +
      theme(aspect.ratio = aspect_ratio,
            axis.text = element_text(color = "black"),
            axis.ticks = element_line(color = "black")) +
      facet_wrap_paginate(~target_compound_label, scales = "free",
                          nrow = n_rows, ncol = n_cols, page = i)
    
    q <- ggplot(d.expr_plot, aes(x = pos, y = log_tpm)) +
      geom_boxplot(fill = "gray", width = 0.7) +
      labs(x = "Paralog",
           y = "log2(TPM)") +
      scale_x_discrete(labels = c("A1", "A2")) +
      theme_bw() +
      theme(aspect.ratio = aspect_ratio,
            axis.text = element_text(color = "black"),
            axis.ticks = element_line(color = "black")) +
      facet_wrap_paginate(~target_compound_label, scales = "free_y",
                          nrow = n_rows, ncol = n_cols, page = i)
  print(ggarrange(p, q, ncol = 2, align = "hv"))
  }
  dev.off()
  
}

```

```{r}
## run make_lm_plot function

d.DeKegel_top5_pairs_drug_sens_A2_expr_lm_pos_slope %>%
  filter(fdr < 0.1) %>%
  distinct(sorted_gene_pair) %>%
  nrow()

d.DeKegel_top5_pairs_drug_sens_A1_expr_lm_pos_slope %>%
  filter(fdr < 0.1) %>%
  distinct(sorted_gene_pair) %>%
  nrow()

# d.DeKegel_top5_pairs_drug_sens_combined_expr_lm_pos_slope

## positive slope scatterplots
# make_paralog_lm_plot(d.DeKegel_top5_pairs_drug_sens_combined_expr_lm_pos_slope,
#              combined_log_tpm, "log2(Combined_TPM)",
#              dependency, "LFC_treated_vs_DMSO",
#              fdr, 0.1, 50)
# 
# make_paralog_lm_plot(d.DeKegel_top5_pairs_drug_sens_A1_expr_lm_pos_slope,
#              A1_log_tpm, "log2(A1_TPM)",
#              dependency, "LFC_treated_vs_DMSO",
#              fdr, 0.1, "all")
# 
# make_paralog_lm_plot(d.DeKegel_top5_pairs_drug_sens_A2_expr_lm_pos_slope,
#              A2_log_tpm, "log2(A2_TPM)",
#              dependency, "LFC_treated_vs_DMSO",
#              fdr, 0.1, 50)

## negative slope scatterplots
# make_paralog_lm_plot(d.DeKegel_top5_pairs_drug_sens_combined_expr_lm_neg_slope,
#              combined_log_tpm, "log2(Combined_TPM)",
#              dependency, "LFC_treated_vs_DMSO",
#              fdr, 0.1, 50)
# 
# make_paralog_lm_plot(d.DeKegel_top5_pairs_drug_sens_A1_expr_lm_neg_slope,
#              A1_log_tpm, "log2(A1_TPM)",
#              dependency, "LFC_treated_vs_DMSO",
#              fdr, 0.1, 50)
# 
# make_paralog_lm_plot(d.DeKegel_top5_pairs_drug_sens_A2_expr_lm_neg_slope,
#              A2_log_tpm, "log2(A2_TPM)",
#              dependency, "LFC_treated_vs_DMSO",
#              fdr, 0.1, 50)

```


```{r}

make_paralog_lm_lineage_plot <- function(df, x_var, x_title, y_title, col_name, stat_cutoff, plot_n){
  
  ## save col_name as a string for later use
  stat_name <- deparse(substitute(col_name))
  
  ## remove the first 2 characters of dataframe name
  df_name <- str_sub(deparse(substitute(df)), start = 3)

  d.plot <- df %>% 
    ungroup() %>%
    ##use {{}} to deparse variable name for use in dplyr functions
    filter({{col_name}} < stat_cutoff) %>%
    unite("target_compound_label", c("target", "compound_name"), sep = " ", remove = FALSE) %>%
    mutate("target_compound_label" = reorder(target_compound_label, {{col_name}}),
           "target" = reorder(target, {{col_name}})) #%>%
  
    if(plot_n == "all") {
    ## do nothing to the df
    ## save output file name
    file_name <- paste0(date, "_", df_name, "_", stat_name, "_L", stat_cutoff, "_all_lineage.pdf")
    
  } else {
    plot_pairs <- d.plot %>%
      arrange({{col_name}}) %>%
      distinct(target_compound_label) %>%
      head(n = plot_n) %>%
      pull(target_compound_label)
    
    d.plot <- d.plot %>%
      filter(target_compound_label %in% plot_pairs)
    
    ## save output file name
    file_name <- paste0(date, "_", df_name, "_", stat_name, "_L", stat_cutoff, "_top", plot_n, "_lineage.pdf")
  }
  
  ## pivot_longer to get into tidy long format (for plotting each gene separately)
  d.plot <- d.plot %>%
    dplyr::select(c(target_compound_label, {{x_var}}, A1_hgnc, A2_hgnc, A1_log_tpm, A2_log_tpm)) %>%
    pivot_longer(cols = starts_with("A"),
                 names_to = c("pos", ".value"), 
                 names_pattern = "A(.)_(.*)") 
  
  length <- d.plot %>%
    ungroup() %>%
    distinct(target_compound_label) %>%
    nrow()

  aspect_ratio <- 0.25
  n_rows <- 2
  n_cols <- 1
  n_pages <- length
  
  pdf(file.path(out_dir, file_name), width = 6, height = 6)
  for(i in 1:n_pages){
    p <- ggplot(d.plot, aes(x = {{x_var}}, y = log_tpm)) +
      geom_jitter(width = 0.3, size = 1) +
      geom_boxplot(alpha = 0, color = "gray55", outlier.shape = NA, coef = 0) +
      labs(x = x_title, y = y_title) +
      theme_bw() +
      theme(axis.text = element_text(color = "black"),
            axis.ticks = element_line(color = "black"),
            axis.text.x = element_text(angle = 45, hjust=1),
            aspect.ratio = aspect_ratio,
            legend.position = "none") +
      facet_wrap_paginate(~ target_compound_label + hgnc,
                          nrow = n_rows, ncol = n_cols, page = i)
    
    print(ggarrange(p, nrow = 1, ncol = 1, align = "hv"))
  }
  dev.off()

  
}


```

```{r}


# make_paralog_lm_lineage_plot(d.DeKegel_top5_pairs_drug_sens_combined_expr_lm_pos_slope,
#              lineage, "Cell_lineage", "log2(TPM)",
#              fdr, 0.1, 50)
# 
# make_paralog_lm_lineage_plot(d.DeKegel_top5_pairs_drug_sens_A1_expr_lm_pos_slope,
#              lineage, "Cell_lineage", "log2(TPM)",
#              fdr, 0.1, "all")
# 
# make_paralog_lm_lineage_plot(d.DeKegel_top5_pairs_drug_sens_A2_expr_lm_pos_slope,
#              lineage, "Cell_lineage", "log2(TPM)",
#              fdr, 0.1, 50)
# 
# make_paralog_lm_lineage_plot(d.DeKegel_top5_pairs_drug_sens_combined_expr_lm_neg_slope,
#              lineage, "Cell_lineage", "log2(TPM)",
#              fdr, 0.1, 50)
# 
# make_paralog_lm_lineage_plot(d.DeKegel_top5_pairs_drug_sens_A1_expr_lm_neg_slope,
#              lineage, "Cell_lineage", "log2(TPM)",
#              fdr, 0.1, 50)
# 
# make_paralog_lm_lineage_plot(d.DeKegel_top5_pairs_drug_sens_A2_expr_lm_neg_slope,
#              lineage, "Cell_lineage", "log2(TPM)",
#              fdr, 0.1, 50)

```

```{r}

## test out getting expr diff and coloring scatterplot points by that
d.DeKegel_top5_pairs_drug_sens_combined_expr_lm_pos_slope %>%
  filter(sorted_gene_pair == "ATP1A1_ATP1A2" & compound_name == "k-strophanthidin") %>%
  mutate(lfc_tpm = A1_log_tpm)

```

##### Write output df
```{r}


write_paralog_lm_output_df <- function(df, col_name, stat_cutoff){
  
  df_out <- df %>%
    filter({{col_name}} < stat_cutoff) %>%
    arrange({{col_name}}) %>%
    distinct(target, compound_id, .keep_all = TRUE) %>%
    dplyr::select(target, compound_name, r_squared:slope, moa)
  
  df_name <- deparse(substitute(df))
  stat_name <- deparse(substitute(col_name))
  
  # df_name <- paste0(deparse(substitute(df)), "_",
  #                   deparse(substitute(col_name)), "_",
  #                   "L", stat_cutoff, ".txt")
  # print(df_name)
  # print(stat_name)
  
  # print(nrow(df_out))
  
  out_path <- file.path(out_dir, paste0(df_name, "_", stat_name, "_", "L", stat_cutoff, ".txt"))
  # print(out_path)
  
  write_tsv(df_out, out_path)
  
  cat("\ndf written to", out_path)
}

```


```{r}

# write_paralog_lm_output_df(d.DeKegel_top5_pairs_drug_sens_combined_expr_lm_pos_slope,
#                            fdr, 0.1)
# 
# write_paralog_lm_output_df(d.DeKegel_top5_pairs_drug_sens_A1_expr_lm_pos_slope,
#                            fdr, 0.1)
# 
# write_paralog_lm_output_df(d.DeKegel_top5_pairs_drug_sens_A2_expr_lm_pos_slope,
#                            fdr, 0.1)
# 
# write_paralog_lm_output_df(d.DeKegel_top5_pairs_drug_sens_combined_expr_lm_neg_slope,
#                            fdr, 0.1)
# 
# write_paralog_lm_output_df(d.DeKegel_top5_pairs_drug_sens_A1_expr_lm_neg_slope,
#                            fdr, 0.1)
# 
# write_paralog_lm_output_df(d.DeKegel_top5_pairs_drug_sens_A2_expr_lm_neg_slope,
#                            fdr, 0.1)

```


#### Outliers

##### Top hits for RR
```{r}

d.drug_sens_all_genes_expr_lm %>%
  filter(target == "PTPN11" | target == "PTPN6") %>%
  distinct(target, compound_id, .keep_all = TRUE) %>%
  dplyr::select(target, compound_name, r_squared:slope, moa)

d.drug_sens_all_genes_expr_lm %>%
  filter(target == "AKT3") %>%
  distinct(target, compound_id, .keep_all = TRUE) %>%
  dplyr::select(target, compound_name, r_squared:slope, moa)

d.drug_sens_all_genes_expr_lm %>%
  filter(target == "AKT3" | target == "AKT2" | target == "AKT1") %>%
  distinct(target, compound_id, .keep_all = TRUE) %>%
  dplyr::select(target, compound_name, r_squared:slope, moa)

d.drug_sens_all_genes_expr_lm %>%
  filter(target == "CDK1" | target == "CDK2" | target == "CDK5") %>%
  distinct(target, compound_id, .keep_all = TRUE) %>%
  dplyr::select(target, compound_name, r_squared:slope, moa)

```

```{r}

## top hits for RR only
target_list <- c("AKT1_AKT2", "AKT2_AKT3", "AKT1_AKT3", "PTPN11_PTPN6", "CDK2_CDK5", "CDK1_CDK2", "MAP2K1_MAP2K2")
compound_list <- c("GSK2110183", "MK−2206", "GDC−0068", "AZD5363", "BVT_948", "bosutinib", "NU6027", "canertinib")

# write_rds(d.DeKegel_top5_pairs_drug_sens_expr, file.path(out_dir, "d.DeKegel_top5_pairs_drug_sens_expr"))

## robustly scale expression data
d.DeKegel_top5_pairs_drug_sens_expr_A2_lm_pos_top_outliers_unfiltered <- d.DeKegel_top5_pairs_drug_sens_expr %>%
  filter(target %in% target_list) %>%
  filter(grepl(str_c(compound_list, collapse = "|"), compound_name)) %>%
  filter(!is.na(A2_log_tpm) | !is.na(dependency)) %>%
  group_by(target, compound_id) %>%
  mutate(scaled_A2_log_tpm = RobScale(A2_log_tpm, center = TRUE, scale = TRUE)) %>%
  mutate(A2_expr_outlier_flag = case_when(
    scaled_A2_log_tpm < -1.5 ~ "low",
    scaled_A2_log_tpm > 1.5 ~ "high",
    TRUE ~ "neither")) %>%
  mutate(scaled_A1_log_tpm = RobScale(A1_log_tpm, center = TRUE, scale = TRUE)) %>%
  mutate(A1_expr_outlier_flag = case_when(
    scaled_A1_log_tpm < -1.5 ~ "low",
    scaled_A1_log_tpm > 1.5 ~ "high",
    TRUE ~ "neither")) %>%
  ungroup() %>%
  mutate(A2_expr_outlier_flag = factor(A2_expr_outlier_flag, levels = c("low", "neither", "high")))%>%
  mutate(A1_expr_outlier_flag = factor(A1_expr_outlier_flag, levels = c("low", "neither", "high")))
d.DeKegel_top5_pairs_drug_sens_expr_A2_lm_pos_top_outliers_unfiltered

```

```{r}


## low vs. neither (with high included but ignoring it) => 
## calculate diff in median LFC between relevant groups

d.DeKegel_top5_pairs_drug_sens_expr_A2_lm_pos_top_outliers <- d.DeKegel_top5_pairs_drug_sens_expr_A2_lm_pos_top_outliers_unfiltered %>%
  group_by(target, compound_id) %>%
  ## get rid of drug/targets that don't have any "low" cell lines
  filter(any(A2_expr_outlier_flag == "low")) %>%
  ungroup() 

d.DeKegel_top5_pairs_drug_sens_expr_A2_lm_pos_top_outliers_grp_med_diff <- d.DeKegel_top5_pairs_drug_sens_expr_A2_lm_pos_top_outliers %>%
  group_by(target, compound_id, A2_expr_outlier_flag) %>%
  ## calculate median dependency score for each outlier group
  summarize(grp_med_dependency = median(dependency)) %>%
  pivot_wider(names_from = A2_expr_outlier_flag, values_from = grp_med_dependency) %>%
  ## calculate difference between low and neither groups
  mutate(grp_med_diff = low - neither) %>%
  dplyr::select(target, compound_id, grp_med_diff) 

d.DeKegel_top5_pairs_drug_sens_expr_A2_lm_pos_top_outliers <- d.DeKegel_top5_pairs_drug_sens_expr_A2_lm_pos_top_outliers %>%
  left_join(d.DeKegel_top5_pairs_drug_sens_expr_A2_lm_pos_top_outliers_grp_med_diff, by = c("target", "compound_id"))

## low outliers and drug sensitivity
## calculate p-value (adjusted for multiple within-group tests?)
d.DeKegel_top5_pairs_drug_sens_expr_A2_lm_pos_top_outliers_wrst_pvals <- d.DeKegel_top5_pairs_drug_sens_expr_A2_lm_pos_top_outliers %>%
  ## remove cell lines that are high outliers (I think this makes sense... but I'm not sure)
  filter(A2_expr_outlier_flag != "high") %>%
  group_by(target, compound_id) %>%
  summarize(p_val = wilcox.test(dependency ~ A2_expr_outlier_flag)$p.value)
# d.drug_sens_all_genes_expr_outlier_low_wrst_sens_pvals

## adjust for multiple testing overall
d.stats <- BH_adjust(d.DeKegel_top5_pairs_drug_sens_expr_A2_lm_pos_top_outliers_wrst_pvals)
# d.stats

d.DeKegel_top5_pairs_drug_sens_expr_A2_lm_pos_top_outliers_wrst <- d.DeKegel_top5_pairs_drug_sens_expr_A2_lm_pos_top_outliers %>%
  left_join(d.stats, by = c("target", "compound_id")) 

d.DeKegel_top5_pairs_drug_sens_expr_A2_lm_pos_top_outliers_wrst_sens <- d.DeKegel_top5_pairs_drug_sens_expr_A2_lm_pos_top_outliers_wrst %>%
  filter(grp_med_diff < 0)

d.DeKegel_top5_pairs_drug_sens_expr_A2_lm_pos_top_outliers_wrst_resist <- d.DeKegel_top5_pairs_drug_sens_expr_A2_lm_pos_top_outliers_wrst %>%
  filter(grp_med_diff > 0)

summarize_stats(d.DeKegel_top5_pairs_drug_sens_expr_A2_lm_pos_top_outliers_wrst_sens)
summarize_stats(d.DeKegel_top5_pairs_drug_sens_expr_A2_lm_pos_top_outliers_wrst_resist)

```

```{r}

# d.DeKegel_top5_pairs_drug_sens_expr_A2_lm_pos_top_outliers_unfiltered %>%
#   group_by(target, compound_id, A2_expr_outlier_flag) %>%
#   summarize(n = n())

d.DeKegel_top5_pairs_drug_sens_expr %>%
  filter(target %in% target_list) %>%
  filter(grepl(str_c(compound_list, collapse = "|"), compound_name)) %>%
  filter(!is.na(A2_log_tpm) | !is.na(dependency)) %>%
  group_by(target, compound_id) %>%
  mutate(scaled_A2_log_tpm = RobScale(A2_log_tpm, center = TRUE, scale = TRUE)) %>%
  mutate(A2_expr_outlier_flag = case_when(
    scaled_A2_log_tpm < -1.5 ~ "low",
    scaled_A2_log_tpm > 1.5 ~ "high",
    TRUE ~ "neither")) %>%
  mutate(scaled_A1_log_tpm = RobScale(A1_log_tpm, center = TRUE, scale = TRUE)) %>%
  mutate(A1_expr_outlier_flag = case_when(
    scaled_A1_log_tpm < -1.5 ~ "low",
    scaled_A1_log_tpm > 1.5 ~ "high", ## before 22-04-26 this said 2
    TRUE ~ "neither")) %>%
  ungroup() %>%
  mutate(A2_expr_outlier_flag = factor(A2_expr_outlier_flag, levels = c("low", "neither", "high")))%>%
  mutate(A1_expr_outlier_flag = factor(A1_expr_outlier_flag, levels = c("low", "neither", "high")))  %>%
  group_by(target, compound_id, A2_expr_outlier_flag) %>%
  summarize(n = n())

d.DeKegel_top5_pairs_drug_sens_expr %>%
  filter(target %in% target_list) %>%
  filter(grepl(str_c(compound_list, collapse = "|"), compound_name)) %>%
  filter(!is.na(A2_log_tpm) | !is.na(dependency)) %>%
  group_by(target, compound_id) %>%
  mutate(scaled_A2_log_tpm = RobScale(A2_log_tpm, center = TRUE, scale = TRUE)) %>%
  mutate(A2_expr_outlier_flag = case_when(
    scaled_A2_log_tpm < -2 ~ "low",
    scaled_A2_log_tpm > 2 ~ "high",
    TRUE ~ "neither")) %>%
  mutate(scaled_A1_log_tpm = RobScale(A1_log_tpm, center = TRUE, scale = TRUE)) %>%
  mutate(A1_expr_outlier_flag = case_when(
    scaled_A1_log_tpm < -2 ~ "low",
    scaled_A1_log_tpm > 2 ~ "high",
    TRUE ~ "neither")) %>%
  ungroup() %>%
  mutate(A2_expr_outlier_flag = factor(A2_expr_outlier_flag, levels = c("low", "neither", "high")))%>%
  mutate(A1_expr_outlier_flag = factor(A1_expr_outlier_flag, levels = c("low", "neither", "high")))  %>%
  group_by(target, compound_id, A2_expr_outlier_flag) %>%
  summarize(n = n())

# d.DeKegel_top5_pairs_drug_sens_expr_A2_lm_pos_top_outliers_wrst %>%
#   distinct(target, compound_id, .keep_all = TRUE) %>%
#   dplyr::select(sorted_gene_pair, compound_name, grp_med_diff:fdr)
# 
# d.DeKegel_top5_pairs_drug_sens_expr_A2_lm_pos_top_outliers_wrst_sens %>%
#   distinct(target, compound_id, .keep_all = TRUE) %>%
#   dplyr::select(sorted_gene_pair, compound_name, grp_med_diff:fdr,
#                 pgPEN_flag:drug_targets)

# d.DeKegel_top5_pairs_drug_sens_expr_A2_lm_pos_top_outliers %>%
#   distinct(target, compound_id, .keep_all = TRUE)%>%
#   group_by(target, compound_id, A2_expr_outlier_flag) %>%
#   summarize(n = n())
# 
# d.DeKegel_top5_pairs_drug_sens_expr_A2_lm_pos_top_outliers_grp_med_diff

```


```{r}

d.DeKegel_top5_pairs_drug_sens_expr_A2_lm_pos_top_outliers_wrst_sens_adjOutlierCutoff <- d.DeKegel_top5_pairs_drug_sens_expr_A2_lm_pos_top_outliers_wrst_sens

df <- d.DeKegel_top5_pairs_drug_sens_expr_A2_lm_pos_top_outliers_wrst_sens_adjOutlierCutoff

  ## remove the first 2 characters of dataframe name
  df_name <- str_sub(deparse(substitute(d.DeKegel_top5_pairs_drug_sens_expr_A2_lm_pos_top_outliers_wrst_sens_adjOutlierCutoff)), start = 3)
  
  file_name <- paste0(date, "_", df_name, "_FDR_all.pdf")
  
  ## make stat label for labeling significance brackets
  stat_label <- paste0("FDR = ", "{fdr}")


  d.plot <- df %>% 
    ungroup() %>%
    ## use {{}} to deparse variable name for use in dplyr functions
    # filter({{col_name}} < stat_cutoff) %>%
    unite("target_compound_label", c("target", "compound_name"), sep = "\n", remove = FALSE) %>%
    mutate("target_compound_label" = reorder(target_compound_label, p_val),
           "target" = reorder(target, p_val)) #%>%


  ## make df to label adjusted FDR/p-val for each plot
  d.label <- d.plot %>%
    group_by(target, compound_id) %>%
    mutate(y.position = (max(dependency) + 0.1*(max(dependency) - min(dependency)))) %>%
    dplyr::ungroup() %>%
    dplyr::select(target_compound_label, p_val, fdr, y.position) %>%
    distinct(target_compound_label, .keep_all = TRUE) %>%
    mutate(group1 = "low",
           group2 = "neither",
           p_val = scientific(p_val, 3),
           fdr = scientific(fdr, 3)) 
  d.label
  
  d.expr_plot <- d.plot %>%
    dplyr::select(c(target_compound_label, A1_log_tpm, A2_log_tpm, 
                    A1_expr_outlier_flag, A2_expr_outlier_flag)) %>%
    pivot_longer(cols = starts_with("A"),
                 names_to = c("pos", ".value"), 
                 names_pattern = "A(.)_(.*)") 

    # file_name <- paste0(date, "_", df_name, "_FDR_all.pdf")
  d.expr_plot
    
    
  length <- d.plot %>%
    ungroup() %>%
    distinct(target_compound_label) %>%
    nrow()
    
  aspect_ratio <- 1
  n_rows <- 1
  n_cols <- 1
  n_pages <- ceiling(length/(n_rows*n_cols))
  

  
  # pdf(file.path(out_dir, file_name), width = 6, height = 3)
  # for(i in 1:n_pages){
  #   p <-  ggplot(d.plot, aes(x = A2_expr_outlier_flag, y = dependency)) +
  #     geom_hline(yintercept = 0) +
  #     geom_jitter(aes(color = A2_expr_outlier_flag), width = 0.3) +
  #     geom_boxplot(alpha = 0) +
  #     stat_pvalue_manual(d.label, label = stat_label, tip.length = 0.03, size = 3, vjust = -0.5) +
  #     ## adds extra space on top & bottom so p-value label will fit
  #     scale_y_continuous(expand = expansion(mult = 0.15)) +
  #     scale_color_manual(values = c("low" = "#0080C6", "neither" = "#7F7F7F", "high" = "#B84250")) +
  #     labs(x = "A2_log_tpm",
  #          y = "Dependency") +
  #     theme_bw() +
  #     theme(aspect.ratio = aspect_ratio,
  #           axis.text = element_text(color = "black"),
  #           axis.ticks = element_line(color = "black"),
  #           legend.position = "none") +
  #     facet_wrap_paginate(~target_compound_label, scales = "free_y",
  #                         nrow = n_rows, ncol = n_cols, page = i)
  # 
  #   q <- ggplot(d.expr_plot, aes(x = pos, y = log_tpm)) +
  #     geom_jitter(aes(color = expr_outlier_flag), width = 0.3) +
  #     geom_boxplot(alpha = 0) +
  #     scale_color_manual(values = c("low" = "#0080C6", "neither" = "#7F7F7F", "high" = "#B84250")) +
  #     labs(x = "Expression_group", y = "log2(TPM)") +
  #     theme_bw() +
  #     theme(aspect.ratio = aspect_ratio,
  #           axis.text = element_text(color = "black"),
  #           axis.ticks = element_line(color = "black")) +
  #     facet_wrap_paginate(~target_compound_label, scale = "free",
  #                         nrow = n_rows, ncol = n_cols, page = i)
  #   print(ggarrange(p, q, ncol = 2, align = "hv", common.legend = TRUE, legend = "right"))
  # }
  # dev.off()


```

```{r}

# make_lineage_plot <- function(df, x_var, x_title, y_var, y_title, col_name, stat_cutoff, plot_n){
#   
#   ## save col_name as a string for later use
#   stat_name <- deparse(substitute(col_name))
#   
#   ## remove the first 2 characters of dataframe name
#   df_name <- str_sub(deparse(substitute(df)), start = 3)
# 
#   d.plot <- df %>% 
#     ungroup() %>%
#     ##use {{}} to deparse variable name for use in dplyr functions
#     filter({{col_name}} < stat_cutoff) %>%
#     unite("target_compound_label", c("target", "compound_name"), sep = "\n", remove = FALSE) %>%
#     mutate("target_compound_label" = reorder(target_compound_label, {{col_name}}),
#            "target" = reorder(target, {{col_name}})) #%>%
#   
#     if(plot_n == "all") {
#       ## don't do anything to the df
#       ## save output file name
#       file_name <- paste0(date, "_", df_name, "_", stat_name, "_L", stat_cutoff, "_all_lineage.pdf")
#     
#   } else {
#       plot_pairs <- d.plot %>%
#         arrange({{col_name}}) %>%
#         distinct(target_compound_label) %>%
#         head(n = plot_n) %>%
#         pull(target_compound_label)
#       
#       d.plot <- d.plot %>%
#         filter(target_compound_label %in% plot_pairs)
#       
#       ## save output file name
#       file_name <- paste0(date, "_", df_name, "_", stat_name, "_L", stat_cutoff, "_top", plot_n, "_lineage.pdf")
#   }
#   
#   length <- d.plot %>%
#     ungroup() %>%
#     distinct(target_compound_label) %>%
#     nrow()
#     
#   # aspect_ratio <- 1
#   n_rows <- 1
#   n_cols <- 1
#   n_pages <- ceiling(length/(n_rows*n_cols))
#   
#   ## make stat label for labeling significance brackets
#   stat_label <- paste0(toupper(stat_name), " = ", "{p_val}")
# 
#   ## save output file name
#   file_name <- paste(date, df_name, stat_name, "L", stat_cutoff, "cell_lineage.pdf", sep = "_")
# 
#   pdf(file.path(out_dir, file_name), width = 6, height = 4)
#   for(i in 1:n_pages){
# 
#     q <- ggplot(d.plot, aes_string(x = x_var, y = y_var)) +
#       geom_jitter(aes(color = expr_outlier_flag), width = 0.3, size = 1) +
#       geom_boxplot(alpha = 0) +
#       scale_color_manual(values = c("low" = "blue", "neither" = "gray50", "high" = "red")) +
#       labs(x = x_title, y = y_title) +
#       theme_bw() +
#       theme(#aspect.ratio = aspect_ratio,
#             axis.text = element_text(color = "black"),
#             axis.ticks = element_line(color = "black"),
#             axis.text.x = element_text(angle = 45, hjust=1)) +
#       facet_wrap_paginate(~target_compound_label, scale = "free",
#                           nrow = n_rows, ncol = n_cols, page = i)
#     print(ggarrange(q, ncol = 1, align = "hv", legend = "right"))
#   }
#   dev.off()
#   
# }


```


```{r}

# write_rds(d.DeKegel_top5_pairs_drug_sens_expr_A2_lm_pos_top_outliers_wrst_sens_adjOutlierCutoff,
#           file.path(out_dir, "d.DeKegel_top5_pairs_drug_sens_expr_A2_lm_pos_top_outliers_wrst_sens_adjOutlierCutoff"))

d.DeKegel_top5_pairs_drug_sens_expr_A2_lm_pos_top_outliers_wrst_sens_adjOutlierCutoff %>%
  filter(target == "AKT1_AKT3" & compound_name == "GSK2110183") %>%
  ggplot(aes(x = lineage, y = A2_log_tpm)) +
      geom_jitter(aes(color = A2_expr_outlier_flag), width = 0.3, size = 1) +
      geom_boxplot(alpha = 0, outlier.shape = NA, coef = 0) +
      scale_color_manual(values = c("low" = "#0080C6", "neither" = "#7F7F7F", "high" = "#B84250")) +
      labs(x = "lineage", y = "log2_TPM") +
      theme_bw() +
      theme(aspect.ratio = 0.25,
            axis.text = element_text(color = "black"),
            axis.ticks = element_line(color = "black"),
            axis.text.x = element_text(angle = 45, hjust=1)) +
  facet_wrap(~A2_hgnc)
# ggsave(file.path(out_dir, paste0(date, "DeKegel_top5_pairs_drug_sens_expr_A2_lm_pos_top_outliers_wrst_sens_adjOutlierCutoff_AKT3_expr_lineage_outliersColored.pdf")), 
#        width = 8, height = 3, useDingbats = FALSE)

d.DeKegel_top5_pairs_drug_sens_expr_A2_lm_pos_top_outliers_wrst_sens_adjOutlierCutoff %>%
  filter(target == "AKT1_AKT3" & compound_name == "GSK2110183") %>%
  ggplot(aes(x = lineage, y = A2_log_tpm)) +
      geom_jitter(width = 0.3, size = 1) +
      geom_boxplot(alpha = 0, color = "gray55", outlier.shape = NA, coef = 0) +
      labs(x = "lineage", y = "log2_TPM") +
      theme_bw() +
      theme(aspect.ratio = 0.25,
            axis.text = element_text(color = "black"),
            axis.ticks = element_line(color = "black"),
            axis.text.x = element_text(angle = 45, hjust=1)) +
  facet_wrap(~A2_hgnc)
# ggsave(file.path(out_dir, paste0(date, "DeKegel_top5_pairs_drug_sens_expr_A2_lm_pos_top_outliers_wrst_sens_adjOutlierCutoff_AKT3_expr_lineage.pdf")), 
#        width = 6, height = 3, useDingbats = FALSE)


```



