---
title: "Synthetic lethal paralog expression vs. drug sensitivity"
author: "Phoebe Parrish"
date: "2023-09-20"
output: html_document
---

# Setup

```{r setup, include=FALSE}

# library(biomaRt)
# library(depmap)
# library(ExperimentHub)
library(tidyverse)
library(ggrepel)
library(ggforce)
library(ggsignif)
library(RColorBrewer)
library(ggpubr) # another way to do significance bars
library(ggprism) # another significance bar package...
library(scales) # scientific notation
library(ggtext) # for using superscript in labels
library(DescTools) # for robust scaling
library(kableExtra)

```


### Assign variables
```{r}

date <- Sys.Date()

in_dir <- file.path("..", "results/integrated_dfs")

out_dir <- file.path("..", "results/expression_output")

options(dplyr.summarise.inform = FALSE)

```


### Import files

```{r, message = FALSE}

d.DeKegel_top5_pairs_drug_sens_expr_annot <- read_rds(file.path(in_dir, "d.DeKegel_top5_pairs_drug_sens_expr_annot"))

d.drug_sens_all_genes_expr <- read_rds(file.path(in_dir, "d.drug_sens_all_genes_expr"))

```

#### Reformat DFs
```{r}

## keep only relevant columns
d.DeKegel_top5_pairs_drug_sens_expr <- d.DeKegel_top5_pairs_drug_sens_expr_annot %>%
  dplyr::select(-c(entrez_pair:prediction_rank, prediction_score:family_size,
                   broad_id, primary_disease, subtype_disease))

d.drug_sens_all_genes_expr <- d.drug_sens_all_genes_expr %>%
  rename(log_tpm = rna_expression)

## how many gene pairs are included in the paralog dataset?
d.DeKegel_top5_pairs_drug_sens_expr %>%
  distinct(sorted_gene_pair) %>%
  nrow()

```

### Import functions

```{r}

source("shared_functions.R")

```


# Analysis

## Single-gene

This analysis is intended to demonstrate expected results for targeted inhibitors and their target (single) genes. 

### PRISM Phase 1 drug sensitivity

#### Continuous/linear regression

```{r}

## running lm for each sorted_gene_pair/compound_id group

## use broom::glance to get rsquared and pval
d.drug_sens_all_genes_expr_lm_summary <- d.drug_sens_all_genes_expr %>%
  filter(!is.na(dependency)) %>%
  filter(!is.na(log_tpm)) %>%
  group_by(target, compound_id) %>%
  group_modify(~ broom::glance(lm(dependency ~ log_tpm, data = .x))) %>%
  dplyr::ungroup() %>%
  rename(r_squared = r.squared, p_val = p.value)
# d.drug_sens_all_genes_expr_lm_summary

## FDR-adjust p-vals for multiple testing and add back to summary DF
d.stats <- BH_adjust(d.drug_sens_all_genes_expr_lm_summary)
# d.stats

d.drug_sens_all_genes_expr_lm_summary <- d.stats %>%
  dplyr::select(-p_val) %>%
  right_join(d.drug_sens_all_genes_expr_lm_summary, by = c("target", "compound_id")) %>%
  dplyr::select(target, compound_id, r_squared, p_val, fdr)
# d.drug_sens_all_genes_expr_lm_summary

## use broom::tidy to get slope
d.drug_sens_all_genes_expr_lm_slope <- d.drug_sens_all_genes_expr %>%
  filter(!is.na(dependency)) %>%
  filter(!is.na(log_tpm)) %>%
  group_by(target, compound_id) %>%
  group_modify(~ broom::tidy(lm(dependency ~ log_tpm, data = .x))) %>%
  dplyr::ungroup() %>%
  ## extract slope of lm fit line from output
  filter(term == "log_tpm") %>%
  rename(slope = estimate) %>%
  dplyr::select(target, compound_id, slope) 
# d.drug_sens_all_genes_expr_lm_slope

## add rsquared, p-val, fdr, and slope info to main DF
d.drug_sens_all_genes_expr_lm <- d.drug_sens_all_genes_expr %>%
  left_join(d.drug_sens_all_genes_expr_lm_summary, by = c("target", "compound_id")) %>%
  left_join(d.drug_sens_all_genes_expr_lm_slope, by = c("target", "compound_id")) %>%
  filter(!is.na(dependency)) %>%
  filter(!is.na(log_tpm))

summarize_stats(d.drug_sens_all_genes_expr_lm)

```



```{r}

## number of pairs with negative vs. positive slope
d.drug_sens_all_genes_expr_lm %>%
  distinct(target, compound_id, .keep_all = TRUE) %>%
  mutate(slope_flag = ifelse(slope < 0, "negative", "positive")) %>%
  group_by(slope_flag) %>%
  summarize(n = n()) %>%
  print_kbl()

## get low expression = low LFC pairs (positive slope)
d.drug_sens_all_genes_expr_lm_pos_slope <-d.drug_sens_all_genes_expr_lm %>%
  filter(slope > 0)

## get high expression = low LFC pairs (negative slope)
d.drug_sens_all_genes_expr_lm_neg_slope <- d.drug_sens_all_genes_expr_lm %>%
  filter(slope < 0)

## summarize significant drug/gene combos for each new DF
summarize_stats(d.drug_sens_all_genes_expr_lm_pos_slope)
summarize_stats(d.drug_sens_all_genes_expr_lm_neg_slope)

```

##### Plot
###### example genes
```{r, message = FALSE}

d.top_neg_hits <- d.drug_sens_all_genes_expr_lm_neg_slope %>% 
  unite("target_compound_label", c("target", "compound_name"), sep = "\n", remove = FALSE) %>% 
  arrange(fdr) %>%
  filter(fdr < 0.05) %>%
  distinct(target_compound_label)
print_kbl(head(d.top_neg_hits, n = 20))

```


```{r}



make_lm_plot <- function(df, plot_list, x_val){
  d.plot <- df %>%
    unite("target_compound_label", c("target", "compound_name"), sep = "\n", remove = FALSE) %>%
    filter(target_compound_label %in% plot_list) %>%
    mutate("target_compound_label" = reorder(target_compound_label, fdr))
  
  d.label <- d.plot %>%
    group_by(target, compound_id) %>%
    mutate(group_label_x = min({{x_val}}),
           group_label_y = max(dependency)) %>%
    dplyr::ungroup() %>%
    dplyr::select(target_compound_label, r_squared, p_val, fdr, group_label_x, group_label_y) %>%
    distinct(target_compound_label, .keep_all = TRUE) %>%
    mutate(r_squared = round(r_squared, 3),
           p_val = scientific(p_val, 3),
           fdr = scientific(fdr, 3)) 
  # d.label
    
  plot <- ggplot(d.plot, aes(x = {{x_val}}, y = dependency)) +
    geom_hline(yintercept = 0) +
    geom_point(size = 1) +
    geom_smooth(method = lm, se = FALSE, color = "gray60") +
    geom_richtext(data = d.label,
                  aes(x = group_label_x, y = group_label_y,
                  ## this package can do HTML-formatting for text
                  label = paste0("r<sup>2</sup>=", r_squared, "<br>", 
                                 "FDR =", fdr)),
                  ## adjust label alignment and remove borders
                  hjust = 0, vjust = 1, label.color = NA, size = 3,
                  ## label background = transparent, text is not
                  fill = alpha(c("white"), 0.75),
                  label.margin = unit(c(0, 0, 0, 0), "lines"),
                  label.padding = unit(c(0, 0, 0, 0), "lines")) +
    labs(x = "log2(TPM)", y = "Cell_viability") +
    theme_bw() +
    theme(aspect.ratio = 1,
          axis.text = element_text(color = "black"),
          axis.ticks = element_line(color = "black")) +
    facet_wrap(~target_compound_label, scales = "free", ncol = 3)
  
  return(plot)
}

```


```{r, message = FALSE}

plot_list <- c("MDM2\nidasanutlin", "ERBB2\nGW-583340", "EGFR\nPD-153035")


selected_top_lm_hits <- make_lm_plot(d.drug_sens_all_genes_expr_lm_neg_slope, plot_list, log_tpm)
selected_top_lm_hits

```


```{r}

d.drug_sens_all_genes_expr_lm_neg_slope_topHits <- d.drug_sens_all_genes_expr_lm_neg_slope %>% 
  filter(fdr < 0.05) %>%
  arrange(fdr) %>%
  distinct(target, compound_id, .keep_all = TRUE) %>%
  dplyr::select(target, compound_name, slope, r_squared:fdr, moa)
print_kbl(head(d.drug_sens_all_genes_expr_lm_neg_slope_topHits))

d.drug_sens_all_genes_expr_lm_pos_slope_fdr_topHits <- d.drug_sens_all_genes_expr_lm_pos_slope %>% 
  filter(fdr < 0.05) %>%
  arrange(fdr) %>%
  distinct(target, compound_id, .keep_all = TRUE) %>%
  dplyr::select(target, compound_name, slope, r_squared:fdr, moa)
print_kbl(head(d.drug_sens_all_genes_expr_lm_pos_slope_fdr_topHits))

```


#### Z-scaled outliers

Since many target genes do not show a truly linear pattern of expression vs. cell viability, and in some cases the linear regression fit line seemed to be strongly influenced by outliers, we decided to take another approach to model these data. We instead used an outlier-based approach to group cell lines into low, high, and "neither" expression categories. 

```{r}

## robustly scale expression data
d.drug_sens_all_genes_expr_outlier <- d.drug_sens_all_genes_expr %>%
  filter(!is.na(log_tpm)) %>%
  filter(!is.na(dependency)) %>%
  group_by(target, compound_id) %>% 
  mutate(scaled_log_tpm = RobScale(log_tpm, center = TRUE, scale = TRUE)) %>%
  mutate(expr_outlier_flag = case_when(
    scaled_log_tpm < -2 ~ "low",
    scaled_log_tpm > 2 ~ "high", 
    TRUE ~ "neither")) %>%
  mutate(expr_low_flag = case_when(
    scaled_log_tpm < -2 ~ TRUE,
    TRUE ~ FALSE)) %>% ## if expression ≥ -2, classify as "not low"
  ungroup() %>%
  mutate(expr_outlier_flag = factor(expr_outlier_flag, levels = c("low", "neither", "high"))) 

## how many target/compound pairs have no expression outliers?
d.drug_sens_all_genes_expr_outlier %>%
  group_by(target, compound_id) %>%
  distinct(expr_low_flag) %>%
  summarize(n = n()) %>%
  filter(n < 2) %>%
  nrow() %>%
  print_kbl()

```


##### Split into positive and negative outliers

```{r}

d.drug_sens_all_genes_expr_outlier %>%
  group_by(target, compound_id, expr_outlier_flag) %>%
  summarize(n = n()) %>%
  group_by(expr_outlier_flag) %>%
  summarize(mean = mean(n),
            median = median(n)) %>%
  print_kbl()

## how many compounds/targets have 0 high or low outlier cell lines?
d.drug_sens_all_genes_expr_outlier %>%
  group_by(target, compound_id, expr_outlier_flag) %>%
  summarize(n = n()) %>%
  ungroup() %>%
  group_by(target, compound_id) %>%
  summarize(n = n()) %>%
  filter(n < 3) %>%
  nrow() %>%
  print_kbl()
  
```

```{r}

## low vs. neither (with high included but ignoring it) => 
## calculate diff in median LFC between relevant groups

d.drug_sens_all_genes_expr_outlier_low <- d.drug_sens_all_genes_expr_outlier %>%
  group_by(target, compound_id) %>%
  ## get rid of drug/targets that don't have any "low" cell lines
  filter(any(expr_outlier_flag == "low")) %>%
  ungroup() 

d.drug_sens_all_genes_expr_outlier_low_grp_med_diff <- d.drug_sens_all_genes_expr_outlier_low %>%
  group_by(target, compound_id, expr_outlier_flag) %>%
  ## calculate median dependency score for each outlier group
  summarize(grp_med_dependency = median(dependency)) %>%
  pivot_wider(names_from = expr_outlier_flag, values_from = grp_med_dependency) %>%
  ## calculate difference between low and neither groups
  mutate(grp_med_diff = low - neither) %>%
  dplyr::select(target, compound_id, grp_med_diff) 

d.drug_sens_all_genes_expr_outlier_low <- d.drug_sens_all_genes_expr_outlier_low %>%
  left_join(d.drug_sens_all_genes_expr_outlier_low_grp_med_diff, by = c("target", "compound_id"))

## high vs. neither (with low included but ignoring it) => 
## calculate diff in median LFC between relevant groups =>
d.drug_sens_all_genes_expr_outlier_high <- d.drug_sens_all_genes_expr_outlier %>%
  group_by(target, compound_id) %>%
  ## get rid of drug/targets that don't have any "high" cell lines
  filter(any(expr_outlier_flag == "high")) %>%
  ungroup() 

d.drug_sens_all_genes_expr_outlier_high_grp_med_diff <- d.drug_sens_all_genes_expr_outlier_high %>%
  group_by(target, compound_id, expr_outlier_flag) %>%
  ## calculate median dependency score for each outlier group
  summarize(grp_med_dependency = median(dependency)) %>%
  pivot_wider(names_from = expr_outlier_flag, values_from = grp_med_dependency) %>%
  ## calculate difference between low and neither groups
  mutate(grp_med_diff = high - neither) %>%
  dplyr::select(target, compound_id, grp_med_diff) 

d.drug_sens_all_genes_expr_outlier_high <- d.drug_sens_all_genes_expr_outlier_high %>%
  left_join(d.drug_sens_all_genes_expr_outlier_high_grp_med_diff, by = c("target", "compound_id"))

```



```{r}

## low outliers and drug sensitivity
## calculate p-value (adjusted for multiple within-group tests?)
d.drug_sens_all_genes_expr_outlier_low_wrst_pvals <- d.drug_sens_all_genes_expr_outlier_low %>%
  ## remove cell lines that are high outliers 
  filter(expr_outlier_flag != "high") %>%
  group_by(target, compound_id) %>%
  summarize(p_val = wilcox.test(dependency ~ expr_outlier_flag)$p.value)
# d.drug_sens_all_genes_expr_outlier_low_wrst_sens_pvals

## adjust for multiple testing overall
d.stats <- BH_adjust(d.drug_sens_all_genes_expr_outlier_low_wrst_pvals)
# d.stats

d.drug_sens_all_genes_expr_outlier_low_wrst <- d.drug_sens_all_genes_expr_outlier_low %>%
  left_join(d.stats, by = c("target", "compound_id")) 

d.drug_sens_all_genes_expr_outlier_low_wrst_sens <- d.drug_sens_all_genes_expr_outlier_low_wrst %>%
  filter(grp_med_diff < 0)

d.drug_sens_all_genes_expr_outlier_low_wrst_resist <- d.drug_sens_all_genes_expr_outlier_low_wrst %>%
  filter(grp_med_diff > 0)

# d.drug_sens_all_genes_expr_outlier_low_wrst_resist

## high outliers and drug sensitivity
d.drug_sens_all_genes_expr_outlier_high_wrst_sens_pvals <- d.drug_sens_all_genes_expr_outlier_high %>%
  ## remove cell lines that are low outliers 
  filter(expr_outlier_flag != "low") %>%
  group_by(target, compound_id) %>%
  summarize(p_val = wilcox.test(dependency ~ expr_outlier_flag)$p.value)
# d.drug_sens_all_genes_expr_outlier_high_wrst_sens_pvals

## adjust for multiple testing overall
d.stats <- BH_adjust(d.drug_sens_all_genes_expr_outlier_high_wrst_sens_pvals)
# d.stats

d.drug_sens_all_genes_expr_outlier_high_wrst <- d.drug_sens_all_genes_expr_outlier_high %>%
  left_join(d.stats, by = c("target", "compound_id"))
# d.drug_sens_all_genes_expr_outlier_high_wrst_sens

## high outliers and drug sensitivity
d.drug_sens_all_genes_expr_outlier_high_wrst_sens <- d.drug_sens_all_genes_expr_outlier_high_wrst %>%
  ## remove cell lines where high outliers don't have a higher median dependency score than the other groups
  filter(grp_med_diff < 0)

d.drug_sens_all_genes_expr_outlier_high_wrst_resist <- d.drug_sens_all_genes_expr_outlier_high_wrst %>%
  ## remove cell lines where low outliers don't have a lower median dependency score than the other groups
  filter(grp_med_diff > 0)

```

```{r}

## low outliers
summarize_stats(d.drug_sens_all_genes_expr_outlier_low_wrst_sens)
summarize_stats(d.drug_sens_all_genes_expr_outlier_low_wrst_resist)

## high outliers
summarize_stats(d.drug_sens_all_genes_expr_outlier_high_wrst_sens)
summarize_stats(d.drug_sens_all_genes_expr_outlier_high_wrst_resist)

```

###### Plot
```{r, message = FALSE}

make_outlier_plot <- function(df, plot_list, direction, outlier_var, expr_var){

  d.plot <- df %>%
    unite("target_compound_label", c("target", "compound_name"), sep = "\n", remove = FALSE) %>%
    filter(target_compound_label %in% plot_list) %>%
    mutate("target_compound_label" = reorder(target_compound_label, fdr))
  
  d.label <- d.plot %>%
    group_by(target, compound_id) %>%
    mutate(y.position = (max(dependency) + 0.1*(max(dependency) - min(dependency)))) %>%
    dplyr::ungroup() %>%
    dplyr::select(target_compound_label, fdr, y.position) %>%
    distinct(target_compound_label, .keep_all = TRUE) %>%
    mutate(group1 = direction,
           group2 = "neither",
           fdr = scientific(fdr, 3)) 
  # d.label
  
  p <-  ggplot(d.plot, aes(x = {{outlier_var}}, y = dependency)) +
    geom_hline(yintercept = 0) +
    geom_jitter(aes(color = {{outlier_var}}), width = 0.3, alpha = 0.6) +
    geom_boxplot(aes(fill = {{outlier_var}}), outlier.shape = NA, coef = 0) +
    stat_pvalue_manual(d.label, label = paste0("FDR = {fdr}"), tip.length = 0.03, size = 3, vjust = -0.5) +
    ## adds extra space on top & bottom so p-value label will fit
    scale_y_continuous(expand = expansion(mult = 0.15)) +
    # scale_color_manual(values = c("low" = "blue", "neither" = "gray50", "high" = "red")) +
    scale_color_manual(values = c("low" = "#0080C6", "neither" = "#7F7F7F", "high" = "#B84250")) +
    scale_fill_manual(values = c("low" = "#0080C6", "neither" = "#7F7F7F", "high" = "#B84250")) +
    labs(x = "Cell_viability",
         y = "Expression_group") +
    theme_bw() +
    theme(aspect.ratio = 0.75,
          axis.text = element_text(color = "black"),
          axis.ticks = element_line(color = "black"),
          legend.position = "none") +
    facet_wrap(~target_compound_label, scales = "free_y",
                        ncol = 1)
  
  q <- ggplot(d.plot, aes(x = target, y = {{expr_var}})) +
    geom_jitter(aes(color = {{outlier_var}}), width = 0.3) +
    geom_boxplot(alpha = 0.3, outlier.shape = NA, coef = 0) +
    # scale_color_manual(values = c("low" = "blue", "neither" = "gray50", "high" = "red")) +
    scale_color_manual(values = c("low" = "#0080C6", "neither" = "#7F7F7F", "high" = "#B84250")) +
    labs(x = "Target", y = "log2(TPM)") +
    theme_bw() +
    theme(aspect.ratio = 1,
          axis.text = element_text(color = "black"),
          axis.ticks = element_line(color = "black")) +
    facet_wrap(~target_compound_label, scale = "free", ncol = 1)
  
  ggarrange(p, q, ncol = 2, align = "hv", 
            common.legend = TRUE, legend = "right")

}
        
```


```{r, fig.height=8}

make_outlier_plot(d.drug_sens_all_genes_expr_outlier_low_wrst_resist,
                  c("MDM2\nidasanutlin", "ERBB2\nGW-583340", "EGFR\nPD-153035"), "low", expr_outlier_flag, log_tpm)

```

Compare this output to that from the linear model-based approach: 
```{r}

selected_top_lm_hits

```



##### Plot lineage

What does the expression of each target gene look like across cell lineages?

```{r}

make_lineage_plot <- function(df, plot_list, y_var, color_flag){

  d.plot <- df %>% 
    ungroup() %>%
    unite("target_compound_label", c("target", "compound_name"), sep = "\n", remove = FALSE) %>%
    filter(target_compound_label %in% plot_list) %>%
    mutate("target_compound_label" = reorder(target_compound_label, fdr),
           "target" = reorder(target, fdr)) 

  ggplot(d.plot, aes(x = lineage, y = {{y_var}})) +
    geom_jitter(aes(color = {{color_flag}}), width = 0.3, size = 1) +
    geom_boxplot(alpha = 0, outlier.shape = NA, coef = 0) +
    scale_color_manual(values = c("low" = "#0080C6", "neither" = "#7F7F7F", "high" = "#B84250")) +
    labs(x = "Cell_lineage", y = "log2(TPM)") +
    theme_bw() +
    theme(axis.text = element_text(color = "black"),
          axis.ticks = element_line(color = "black"),
          axis.text.x = element_text(angle = 45, hjust=1)) +
    facet_wrap(~target, scales = "free_y",
               ncol = 1)
}


```

```{r, fig.height=6}

make_lineage_plot(d.drug_sens_all_genes_expr_outlier_low_wrst_resist, 
                  c("MDM2\nidasanutlin", "ERBB2\nGW-583340", "EGFR\nPD-153035"), log_tpm, expr_outlier_flag)


```



## Paralogs

### PRISM Phase 1 drug sensitivity

```{r}

## add target col so single-gene functions will work
d.DeKegel_top5_pairs_drug_sens_expr <- d.DeKegel_top5_pairs_drug_sens_expr %>%
  mutate(target = sorted_gene_pair)
# d.DeKegel_top5_pairs_drug_sens_expr

```

#### Continuous/linear regression

```{r}

## A1 expr

## use broom::glance to get rsquared and pval
d.DeKegel_top5_pairs_drug_sens_A1_expr_lm_summary <- d.DeKegel_top5_pairs_drug_sens_expr %>%
  filter(!is.na(dependency)) %>%
  filter(!is.na(A1_log_tpm)) %>%
  group_by(target, compound_id) %>%
  group_modify(~ broom::glance(lm(dependency ~ A1_log_tpm, data = .x))) %>%
  dplyr::ungroup() %>%
  rename(r_squared = r.squared, p_val = p.value)
# d.DeKegel_top5_pairs_drug_sens_A1_expr_lm_summary

## FDR-adjust p-vals for multiple testing and add back to summary DF
d.stats <- BH_adjust(d.DeKegel_top5_pairs_drug_sens_A1_expr_lm_summary)
# d.stats

d.DeKegel_top5_pairs_drug_sens_A1_expr_lm_summary <- d.stats %>%
  dplyr::select(-p_val) %>%
  right_join(d.DeKegel_top5_pairs_drug_sens_A1_expr_lm_summary, by = c("target", "compound_id")) %>%
  dplyr::select(target, compound_id, r_squared, p_val, fdr)
# d.DeKegel_top5_pairs_drug_sens_A1_expr_lm_summary

## use broom::tidy to get slope
## note: group_modify expects a DF output
d.DeKegel_top5_pairs_drug_sens_A1_expr_lm_slope <- d.DeKegel_top5_pairs_drug_sens_expr %>%
  filter(!is.na(dependency)) %>%
  filter(!is.na(A1_log_tpm)) %>%
  group_by(target, compound_id) %>%
  group_modify(~ broom::tidy(lm(dependency ~ A1_log_tpm, data = .x))) %>%
  dplyr::ungroup() %>%
  ## extract slope of lm fit line from output
  filter(term == "A1_log_tpm") %>%
  rename(slope = estimate) %>%
  dplyr::select(target, compound_id, slope) 
# d.DeKegel_top5_pairs_drug_sens_A1_expr_lm_slope

## add rsquared, p-val, fdr, and slope info to main DF
d.DeKegel_top5_pairs_drug_sens_A1_expr_lm <- d.DeKegel_top5_pairs_drug_sens_expr %>%
  left_join(d.DeKegel_top5_pairs_drug_sens_A1_expr_lm_summary, by = c("target", "compound_id")) %>%
  left_join(d.DeKegel_top5_pairs_drug_sens_A1_expr_lm_slope, by = c("target", "compound_id")) %>%
  filter(!is.na(dependency)) %>%
  filter(!is.na(A1_log_tpm))

summarize_stats(d.DeKegel_top5_pairs_drug_sens_A1_expr_lm)

```

```{r}

## A2 expr

## use broom::glance to get rsquared and pval
d.DeKegel_top5_pairs_drug_sens_A2_expr_lm_summary <- d.DeKegel_top5_pairs_drug_sens_expr %>%
  filter(!is.na(dependency)) %>%
  filter(!is.na(A2_log_tpm)) %>%
  group_by(target, compound_id) %>%
  group_modify(~ broom::glance(lm(dependency ~ A2_log_tpm, data = .x))) %>%
  dplyr::ungroup() %>%
  rename(r_squared = r.squared, p_val = p.value)
# d.DeKegel_top5_pairs_drug_sens_A2_expr_lm_summary

## FDR-adjust p-vals for multiple testing and add back to summary DF
d.stats <- BH_adjust(d.DeKegel_top5_pairs_drug_sens_A2_expr_lm_summary)
# d.stats

d.DeKegel_top5_pairs_drug_sens_A2_expr_lm_summary <- d.stats %>%
  dplyr::select(-p_val) %>%
  right_join(d.DeKegel_top5_pairs_drug_sens_A2_expr_lm_summary, by = c("target", "compound_id")) %>%
  dplyr::select(target, compound_id, r_squared, p_val, fdr)
# d.DeKegel_top5_pairs_drug_sens_A2_expr_lm_summary

## use broom::tidy to get slope
## note: group_modify expects a DF output
d.DeKegel_top5_pairs_drug_sens_A2_expr_lm_slope <- d.DeKegel_top5_pairs_drug_sens_expr %>%
  filter(!is.na(dependency)) %>%
  filter(!is.na(A2_log_tpm)) %>%
  group_by(target, compound_id) %>%
  group_modify(~ broom::tidy(lm(dependency ~ A2_log_tpm, data = .x))) %>%
  dplyr::ungroup() %>%
  ## extract slope of lm fit line from output
  filter(term == "A2_log_tpm") %>%
  rename(slope = estimate) %>%
  dplyr::select(target, compound_id, slope) 
# d.DeKegel_top5_pairs_drug_sens_A2_expr_lm_slope

## add rsquared, p-val, fdr, and slope info to main DF
d.DeKegel_top5_pairs_drug_sens_A2_expr_lm <- d.DeKegel_top5_pairs_drug_sens_expr %>%
  left_join(d.DeKegel_top5_pairs_drug_sens_A2_expr_lm_summary, by = c("target", "compound_id")) %>%
  left_join(d.DeKegel_top5_pairs_drug_sens_A2_expr_lm_slope, by = c("target", "compound_id")) %>%
  filter(!is.na(dependency)) %>%
  filter(!is.na(A2_log_tpm))

summarize_stats(d.DeKegel_top5_pairs_drug_sens_A2_expr_lm)

```



```{r}

d.DeKegel_top5_pairs_drug_sens_A1_expr_lm %>%
  distinct(target, compound_id, .keep_all = TRUE) %>%
  mutate(slope_flag = ifelse(slope < 0, "negative", "positive")) %>%
  group_by(slope_flag) %>%
  summarize(n = n()) %>%
  print_kbl()

d.DeKegel_top5_pairs_drug_sens_A2_expr_lm %>%
  distinct(target, compound_id, .keep_all = TRUE) %>%
  mutate(slope_flag = ifelse(slope < 0, "negative", "positive")) %>%
  group_by(slope_flag) %>%
  summarize(n = n()) %>%
  print_kbl()

```


```{r}

## get low expression = low LFC pairs (positive slope)
d.DeKegel_top5_pairs_drug_sens_A1_expr_lm_pos_slope <-d.DeKegel_top5_pairs_drug_sens_A1_expr_lm %>%
  filter(slope > 0)
summarize_stats(d.DeKegel_top5_pairs_drug_sens_A1_expr_lm_pos_slope)

## get high expression = low LFC pairs (negative slope)
d.DeKegel_top5_pairs_drug_sens_A1_expr_lm_neg_slope <- d.DeKegel_top5_pairs_drug_sens_A1_expr_lm %>%
  filter(slope < 0)
summarize_stats(d.DeKegel_top5_pairs_drug_sens_A1_expr_lm_neg_slope)

## get low expression = low LFC pairs (positive slope)
d.DeKegel_top5_pairs_drug_sens_A2_expr_lm_pos_slope <-d.DeKegel_top5_pairs_drug_sens_A2_expr_lm %>%
  filter(slope > 0)
summarize_stats(d.DeKegel_top5_pairs_drug_sens_A2_expr_lm_pos_slope)

## get high expression = low LFC pairs (negative slope)
d.DeKegel_top5_pairs_drug_sens_A2_expr_lm_neg_slope <- d.DeKegel_top5_pairs_drug_sens_A2_expr_lm %>%
  filter(slope < 0)
summarize_stats(d.DeKegel_top5_pairs_drug_sens_A2_expr_lm_neg_slope)

```

##### Plot A1 top hits

```{r}

A1_top_hits <- d.DeKegel_top5_pairs_drug_sens_A1_expr_lm_pos_slope %>%
  dplyr::select(target, compound_name, r_squared, fdr) %>%
  distinct(target, compound_name, .keep_all = TRUE) %>%
  arrange(fdr) 

print_kbl(head(A1_top_hits, n = 20))

```


```{r, message = FALSE, fig.height = 8}

A1_top_hits_plot_list <- A1_top_hits %>%
  arrange(fdr) %>%
  unite("target_compound_label", c("target", "compound_name"), sep = "\n", remove = FALSE) %>%
  head(n = 9) %>%
  pull(target_compound_label)
# A1_top_hits_plot_list

make_lm_plot(d.DeKegel_top5_pairs_drug_sens_A1_expr_lm_pos_slope, A1_top_hits_plot_list, A1_log_tpm)
# d.DeKegel_top5_pairs_drug_sens_A1_expr_lm_pos_slope


```

##### Plot A2 top hits

```{r}

A2_top_hits <- d.DeKegel_top5_pairs_drug_sens_A2_expr_lm_pos_slope %>%
  dplyr::select(target, compound_name, r_squared, fdr) %>%
  distinct(target, compound_name, .keep_all = TRUE) %>%
  arrange(fdr) 

print_kbl(head(A2_top_hits, n = 20))

```


```{r, message = FALSE, fig.height = 8}

A2_top_hits_plot_list <- A2_top_hits %>%
  arrange(fdr) %>%
  unite("target_compound_label", c("target", "compound_name"), sep = "\n", remove = FALSE) %>%
  head(n = 9) %>%
  pull(target_compound_label)
# A1_top_hits_plot_list

make_lm_plot(d.DeKegel_top5_pairs_drug_sens_A2_expr_lm_pos_slope, A2_top_hits_plot_list, A2_log_tpm)
# d.DeKegel_top5_pairs_drug_sens_A1_expr_lm_pos_slope


```




##### Write output df
```{r}


# write_paralog_lm_output_df <- function(df, col_name, stat_cutoff){
#   
#   df_out <- df %>%
#     filter({{col_name}} < stat_cutoff) %>%
#     arrange({{col_name}}) %>%
#     distinct(target, compound_id, .keep_all = TRUE) %>%
#     dplyr::select(target, compound_name, r_squared:slope, moa)
#   
#   df_name <- deparse(substitute(df))
#   stat_name <- deparse(substitute(col_name))
#   
#   # df_name <- paste0(deparse(substitute(df)), "_",
#   #                   deparse(substitute(col_name)), "_",
#   #                   "L", stat_cutoff, ".txt")
#   # print(df_name)
#   # print(stat_name)
#   
#   # print(nrow(df_out))
#   
#   out_path <- file.path(out_dir, paste0(df_name, "_", stat_name, "_", "L", stat_cutoff, ".txt"))
#   # print(out_path)
#   
#   write_tsv(df_out, out_path)
#   
#   cat("\ndf written to", out_path)
# }

```


```{r}

# write_paralog_lm_output_df(d.DeKegel_top5_pairs_drug_sens_A1_expr_lm_pos_slope,
#                            fdr, 0.1)
# 
# write_paralog_lm_output_df(d.DeKegel_top5_pairs_drug_sens_A2_expr_lm_pos_slope,
#                            fdr, 0.1)
# 
# write_paralog_lm_output_df(d.DeKegel_top5_pairs_drug_sens_A1_expr_lm_neg_slope,
#                            fdr, 0.1)
# 
# write_paralog_lm_output_df(d.DeKegel_top5_pairs_drug_sens_A2_expr_lm_neg_slope,
#                            fdr, 0.1)

```


#### Outliers for top lm hits

```{r}

## A2 top hits
target_list <- c("AKT1_AKT2", "AKT2_AKT3", "AKT1_AKT3", "PTPN11_PTPN6", "CDK2_CDK5", "CDK1_CDK2", "MAP2K1_MAP2K2")
compound_list <- c("GSK2110183", "MK−2206", "GDC−0068", "AZD5363", "BVT_948", "bosutinib", "NU6027", "canertinib")

## robustly scale expression data
d.DeKegel_top5_pairs_drug_sens_expr_A2_lm_pos_top_outliers_unfiltered <- d.DeKegel_top5_pairs_drug_sens_expr %>%
  filter(target %in% target_list) %>%
  filter(grepl(str_c(compound_list, collapse = "|"), compound_name)) %>%
  filter(!is.na(A2_log_tpm) | !is.na(dependency)) %>%
  group_by(target, compound_id) %>%
  mutate(scaled_A2_log_tpm = RobScale(A2_log_tpm, center = TRUE, scale = TRUE)) %>%
  mutate(A2_expr_outlier_flag = case_when(
    scaled_A2_log_tpm < -1.5 ~ "low",
    scaled_A2_log_tpm > 1.5 ~ "high",
    TRUE ~ "neither")) %>%
  mutate(scaled_A1_log_tpm = RobScale(A1_log_tpm, center = TRUE, scale = TRUE)) %>%
  mutate(A1_expr_outlier_flag = case_when(
    scaled_A1_log_tpm < -1.5 ~ "low",
    scaled_A1_log_tpm > 1.5 ~ "high",
    TRUE ~ "neither")) %>%
  ungroup() %>%
  mutate(A2_expr_outlier_flag = factor(A2_expr_outlier_flag, levels = c("low", "neither", "high")))%>%
  mutate(A1_expr_outlier_flag = factor(A1_expr_outlier_flag, levels = c("low", "neither", "high")))
# d.DeKegel_top5_pairs_drug_sens_expr_A2_lm_pos_top_outliers_unfiltered

```

```{r}

d.DeKegel_top5_pairs_drug_sens_expr_A2_lm_pos_top_outliers <- d.DeKegel_top5_pairs_drug_sens_expr_A2_lm_pos_top_outliers_unfiltered %>%
  group_by(target, compound_id) %>%
  ## get rid of drug/targets that don't have any "low" cell lines
  filter(any(A2_expr_outlier_flag == "low")) %>%
  ungroup() 

d.DeKegel_top5_pairs_drug_sens_expr_A2_lm_pos_top_outliers_grp_med_diff <- d.DeKegel_top5_pairs_drug_sens_expr_A2_lm_pos_top_outliers %>%
  group_by(target, compound_id, A2_expr_outlier_flag) %>%
  ## calculate median dependency score for each outlier group
  summarize(grp_med_dependency = median(dependency)) %>%
  pivot_wider(names_from = A2_expr_outlier_flag, values_from = grp_med_dependency) %>%
  ## calculate difference between low and neither groups
  mutate(grp_med_diff = low - neither) %>%
  dplyr::select(target, compound_id, grp_med_diff) 

d.DeKegel_top5_pairs_drug_sens_expr_A2_lm_pos_top_outliers <- d.DeKegel_top5_pairs_drug_sens_expr_A2_lm_pos_top_outliers %>%
  left_join(d.DeKegel_top5_pairs_drug_sens_expr_A2_lm_pos_top_outliers_grp_med_diff, by = c("target", "compound_id"))

## low outliers and drug sensitivity
d.DeKegel_top5_pairs_drug_sens_expr_A2_lm_pos_top_outliers_wrst_pvals <- d.DeKegel_top5_pairs_drug_sens_expr_A2_lm_pos_top_outliers %>%
  filter(A2_expr_outlier_flag != "high") %>%
  group_by(target, compound_id) %>%
  summarize(p_val = wilcox.test(dependency ~ A2_expr_outlier_flag)$p.value)
# d.drug_sens_all_genes_expr_outlier_low_wrst_sens_pvals

## adjust for multiple testing overall
d.stats <- BH_adjust(d.DeKegel_top5_pairs_drug_sens_expr_A2_lm_pos_top_outliers_wrst_pvals)
# d.stats

d.DeKegel_top5_pairs_drug_sens_expr_A2_lm_pos_top_outliers_wrst <- d.DeKegel_top5_pairs_drug_sens_expr_A2_lm_pos_top_outliers %>%
  left_join(d.stats, by = c("target", "compound_id")) 

## filter for hits where low-expressing cell lines are more sensitive to the drug
## (low group has lower median than neither group)
d.DeKegel_top5_pairs_drug_sens_expr_A2_lm_pos_top_outliers_wrst_sens <- d.DeKegel_top5_pairs_drug_sens_expr_A2_lm_pos_top_outliers_wrst %>%
  filter(grp_med_diff < 0)

## filter for hits where low-expressing ell lines are more resistant to the drug
## (low group has higher median than neither group)
d.DeKegel_top5_pairs_drug_sens_expr_A2_lm_pos_top_outliers_wrst_resist <- d.DeKegel_top5_pairs_drug_sens_expr_A2_lm_pos_top_outliers_wrst %>%
  filter(grp_med_diff > 0)

summarize_stats(d.DeKegel_top5_pairs_drug_sens_expr_A2_lm_pos_top_outliers_wrst_sens)
summarize_stats(d.DeKegel_top5_pairs_drug_sens_expr_A2_lm_pos_top_outliers_wrst_resist)

```

##### Plot

```{r, fig.height=9}

# d.DeKegel_top5_pairs_drug_sens_expr_A2_lm_pos_top_outliers_wrst_sens

plot_list <- d.DeKegel_top5_pairs_drug_sens_expr_A2_lm_pos_top_outliers_wrst_sens %>%
  unite("target_compound_id", c("target", "compound_name"), sep = "\n", remove = FALSE) %>%
  distinct(target_compound_id) %>%
  pull(target_compound_id)

make_outlier_plot(d.DeKegel_top5_pairs_drug_sens_expr_A2_lm_pos_top_outliers_wrst_sens,
                  plot_list[1:4], "low", A2_expr_outlier_flag, A2_log_tpm)

make_outlier_plot(d.DeKegel_top5_pairs_drug_sens_expr_A2_lm_pos_top_outliers_wrst_sens,
                  plot_list[5:8], "low", A2_expr_outlier_flag, A2_log_tpm)

```

```{r, fig.height=7}

make_lineage_plot(d.DeKegel_top5_pairs_drug_sens_expr_A2_lm_pos_top_outliers_wrst_sens,
                  plot_list, A2_log_tpm, A2_expr_outlier_flag)

```

```{r}

write_rds(d.DeKegel_top5_pairs_drug_sens_expr_A2_lm_pos_top_outliers_wrst_sens, 
          file.path(out_dir, "d.top_outlier_pairs"))

```
